### 2022-09-09 manuscript/supplement.tex
start supplement file

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\begin{document}

\begin{center}
\textbf{Supplementary material for:}\\
\emph{Geographically aware estimates of remotely sensed water properties for Chesapeake Bay}
\end{center}

\begin{figure}[h!]
    \begin{center}
          \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/counts_obs_table}
    \end{center}    
\end{figure}

\clearpage

\end{document}

### 2022-09-09 figures/validation.py
point validation plot to only test data

@@ -18,20 +18,19 @@ rf_random = pickle.load(open(rf_random_path, "rb"))
 
 predictors = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
 
-X_predict_raw = pd.read_csv("data/data_w_fwi.csv").dropna()
-X_predict = call_data2.clean_data(
-    variable, var_col, predictors, test_size=0, data=X_predict_raw
-)
+X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
+y_test = pickle.load(open("data/y_test_" + variable + ".pkl", "rb"))
 
-predictions = rf_random.predict(X_predict[0])
-res = pd.DataFrame(X_predict[1], columns=["obs"])
+predictions = rf_random.predict(X_test)
+res = pd.DataFrame(y_test, columns=["obs"])
 res["predict"] = predictions.copy()
 
 plt.close()
 fig, ax = plt.subplots(figsize=(8, 4))
 
 g = sns.scatterplot(data=res, x="predict", y="obs", ax=ax, s=7, color=".15")
-sns.histplot(data=res, x="predict", y="obs", bins=100, pthresh=0.1,cmap="mako", ax=ax)
+# increasing bin makes coarser boxes, increasing pthresh makes less boxes
+sns.histplot(data=res, x="predict", y="obs", bins=100, pthresh=0.2,cmap="mako", ax=ax)
 plt.plot([0, 28], [0, 28], color="red")
 ax.set_xlim(-0.5, 32)
 ax.set_ylim(-0.5, 32)

### 2022-09-09 figures/obs_stats.py
stash code for obs fun fact

@@ -50,6 +50,11 @@ def _count_stats(dt, variable):
     dt_sub = utils.freq_count(
         dt_sub, ["daystamp", "source", "station_id"]
     )  # collapse to daily
+
+    # What are the top 3 programs and what percentage of obs do they account for?
+    utils.freq_count(dt_sub, ["source"]).head(3)
+    sum(utils.freq_count(dt_sub, ["source"]).iloc[0:3]["count"]) / sum(utils.freq_count(dt_sub, ["source"])["count"])
+
     n_measures = (
         utils.freq_count(dt_sub, ["daystamp", "source", "station_id"])
         .sort_values(["station_id", "daystamp"], ascending=False)

### 2022-09-09 figures/importance.py
cleanup panel titles

@@ -20,7 +20,7 @@ def imp_plot(variable):
     plt.figure(figsize=(5.7, 5))
 
     plt.barh(y=dset["attr"], width=dset["importance"], color="#1976D2")
-    plt.title(variable + " - feature importances", fontsize=16, fontweight="bold", pad=20)
+    plt.title(variable, fontsize=16, fontweight="bold", pad=20)
     plt.xlabel("Importance", fontsize=14, labelpad=20)
 
     # plt.show()

### 2022-09-12 manuscript/manuscript.tex
ms styling

\documentclass{article}
\usepackage{graphicx}
\usepackage{lipsum}
\usepackage[margin=1in]{geometry}
\usepackage{lineno}
\modulolinenumbers[2]
\usepackage[authoryear]{natbib}

\title{Geographically aware estimates of remotely sensed water properties for Chesapeake Bay}
\author{Jemma Stachelek$^{1}$, Sofia Avendaño$^{1,}$$^{2}$, Jon Schwenk$^{1}$  \\
        \small $^{1}$Los Alamos National Laboratory, Division of Earth and Environmental Sciences, Los Alamos, NM 87545, USA \\
        \small $^{2}$Nicholas School of the Environment, Duke University, Durham, NC 27708, USA \\
}
\date{}

\begin{document}
\maketitle

\begin{abstract}
    \noindent Remotely sensed water properties are important for a variety of applications, including validation of Earth Systems models (ESMs). However, the usefulness of operational forecasting products and directly-observing satellite-based sensors for validation of next-generation ESMs is limited due to their temporal availability and spatial resolution of $<$ 1 year and $>$ 10 km2 respectively. To address this validation data gap, we developed a data-driven model to produce high-resolution ($<$ 1 km2) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs. Our model fits daily MODIS Aqua reflectance data to surface observations ($<$ 1 m depth) from 2000-2021 in Chesapeake Bay, USA. Our model has similar error statistics as prior efforts of this type for salinity (RMSE: 2.2) and temperature (RMSE: 1.9 C). However, unlike prior efforts our model is set up as a pipeline meaning that it has the advantage of producing predictions of water properties in future time periods as additional MODIS data is collected. In addition, our study is unique in that the predictions produced by our model are “geographically-aware” insofar as they capture geographic variation in the influence of flow and surface water exchange in upstream coastal watersheds. 
    \end{abstract} \hspace{10pt}

\linenumbers

\section{Introduction}

In partially mixed estuaries such as the Chesapeake Bay (hereafter, “the Bay”), productivity, organic matter production, and phytoplankton community composition are largely determined by stratification. Because stratification is primarily driven by salinity and temperature, extensive measurement networks exist to monitor these water properties. This data is critical for many uses such as establishing point-based trends and relationships (e.g. river flow or sea level rise vs salinity). However, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \citep{hoodChesapeakeBayProgram2021}. Applications such as these often turn to alternatives to direct observations such as output from directly-observing satellite-based sensors or operational now/fore-casting products. Unfortunately, these too are of limited use for applications such as validation of next-generation ESMs, where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 10 $km^2$). To address this validation data gap, we developed a data-driven model to produce high-resolution ($<$ 1 $km^2$) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.

\section{Methods}
This is the second section

\section{Results}

\begin{figure}[ht!]
    \begin{center}
          \includegraphics[width=0.95\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
    \end{center}    
\end{figure}

\begin{figure}[ht!]
    \begin{center}
          \includegraphics[width=0.95\textwidth,keepaspectratio]{figures/_importance_all}
    \end{center}    
\end{figure}

\section{Discussion}

\bibliographystyle{apalike}
\bibliography{manuscript/geowq}

\end{document}

### 2022-09-12 manuscript/manuscript.tex
test layout

@@
 \documentclass{article}
+\usepackage[breaklinks]{hyperref}
 \usepackage{graphicx}
@@
     \noindent Remotely sensed water properties are important for a variety of applications, including validation of Earth Systems models (ESMs).
-However, the usefulness of operational forecasting products and directly-observing satellite-based sensors for validation of next-generation ESMs is limited due to their temporal availability and spatial resolution of $<$ 1 year and $>$ 10 km2 respectively.
-To address this validation data gap, we developed a data-driven model to produce high-resolution ($<$ 1 km2) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.
+However, the usefulness of operational forecasting products and directly-observing satellite-based sensors for validation of next-generation ESMs is limited due to their temporal availability and spatial resolution of $<$ 1 year and $>$ 30 $km^2$ respectively.
+To address this validation data gap, we developed a data-driven model to produce high-resolution ($<$ 1 $km^2$) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.
 Our model fits daily MODIS Aqua reflectance data to surface observations ($<$ 1 m depth) from 2000-2021 in Chesapeake Bay, USA.
@@
 
-In partially mixed estuaries such as the Chesapeake Bay (hereafter, “the Bay”), productivity, organic matter production, and phytoplankton community composition are largely determined by stratification.
-Because stratification is primarily driven by salinity and temperature, extensive measurement networks exist to monitor these water properties.
-This data is critical for many uses such as establishing point-based trends and relationships (e.g.
-river flow or sea level rise vs salinity).
+In partially mixed estuaries such as the Chesapeake Bay (hereafter, “the Bay”), productivity, organic matter production, and phytoplankton community composition are largely determined by stratification \citep{xuClimateForcingSalinity2012}.
+Because stratification is primarily controlled by salinity and temperature, extensive measurement networks exist to monitor these water properties (Chesapeake Bay Program Water Quality Database, \url{https://www.chesapeakebay.net/what/downloads/cbp_water_quality_database_1984_present}).
+These data are critical for many uses such as establishing point-based trends and relationships (e.g.
+river flow or sea level rise vs salinity or hypoxia, \citet{hagyHypoxiaChesapeakeBay2004}.
 However, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \citep{hoodChesapeakeBayProgram2021}.
-Applications such as these often turn to alternatives to direct observations such as output from directly-observing satellite-based sensors or operational now/fore-casting products.
-Unfortunately, these too are of limited use for applications such as validation of next-generation ESMs, where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 10 $km^2$).
-To address this validation data gap, we developed a data-driven model to produce high-resolution ($<$ 1 $km^2$) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.
+For example, habitat suitability calculations are typically done on a gridded basis for the entirety of a particular estuary rather than at selected point locations.
+Applications such as these often turn to gridded alternatives to direct observations such as output from directly-observing satellite-based sensors \citep{foreCombinedActivePassive2016} or operational now/fore-casting products \citep{lanerolle2011second}.
+Unfortunately, these too are of limited use for applications such as validation of next-generation ESMs, where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
+
+To address this validation data gap, we developed a data-driven model to produce high-resolution (< 1 $km^2$) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.
+The primary input to our model is remotely sensed reflectance data.
+In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \citep{wangDevelopmentMODISData2018}.
+We supplemented MODIS data with gridded estimates of tributary inflows which we used to enforce physical realism (i.e.
+geographic “awareness”) especially in areas of the Bay with little to no observations.
 
@@
     \begin{center}
-          \includegraphics[width=0.95\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
+          \caption{Comparison between Random Forest prediction results and a CBOFS snapshot for Sept, 4, 2022.}
     \end{center}    
@@
           \includegraphics[width=0.95\textwidth,keepaspectratio]{figures/_importance_all}
+          \caption{Random Forest importance plot for salinity, temperature, and turbidity models.}
     \end{center}    

### 2022-09-12 src/rf_icom_utils.py
elimination not extraction

@@ -33,7 +33,7 @@ def run_rfe(X_train,
             variable,
             overwrite=False):
     """
-    Run recursive feature extraction using RFECV
+    Run recursive feature elimination using RFECV
     Inputs: 
     X_train => array of feature values (training set)
     y_train => array of label values (training set)

### 2022-09-12 manuscript/manuscript.tex
ms edits

@@
 \usepackage{graphicx}
+\usepackage{placeins}
 \usepackage{lipsum}
@@
 \usepackage[authoryear]{natbib}
+\usepackage{draftwatermark}
+\SetWatermarkScale{5}
 
@@
 These data are critical for many uses such as establishing point-based trends and relationships (e.g.
-river flow or sea level rise vs salinity or hypoxia, \citet{hagyHypoxiaChesapeakeBay2004}.
-However, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \citep{hoodChesapeakeBayProgram2021}.
+river flow or sea level rise vs salinity or hypoxia), \citet{hagyHypoxiaChesapeakeBay2004}.
+Of particular importance is monitoring for increased nutrient loads as result of expected increases in precipitation, tributary discharge, and river flows as a result of climate change \citep{najjarPotentialClimatechangeImpacts2010, irbyCompetingImpactsClimate2018}.
+
+Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \citep{hoodChesapeakeBayProgram2021}.
 For example, habitat suitability calculations are typically done on a gridded basis for the entirety of a particular estuary rather than at selected point locations.
 Applications such as these often turn to gridded alternatives to direct observations such as output from directly-observing satellite-based sensors \citep{foreCombinedActivePassive2016} or operational now/fore-casting products \citep{lanerolle2011second}.
-Unfortunately, these too are of limited use for applications such as validation of next-generation ESMs, where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
+Unfortunately, these too are of limited use for applications such as validation of next-generation Earth Systems Models (ESMs, \citet{golazDOEE3SMModel2022}), where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
 
@@
 geographic “awareness”) especially in areas of the Bay with little to no observations.
+Our model represents a further advantage over prior efforts at data-driven modeling of salinity in the Bay (e.g.
+\citet{vogelAssessingSatelliteSea2016}) because it is calibrated against not only observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
 
 \section{Methods}
-This is the second section
+
+\subsection{Study site}
+
+Chesapeake Bay is located in the Mid-Atlantic region of the United States and serves as the outlet of 9 major tributary systems.
+Of these, the Susquehanna River provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, Xu et al.
+2012).
+The Bay is fairly shallow ($<$ 30m, \citet{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed (Xu et al.
+2012).
+
+\subsection{Model overview}
+
+Prior data-driven models (e.g.
+\citet{urquhartGeospatialInterpolationMODISderived2013, vogelAssessingSatelliteSea2016}) of Bay salinity used Generalized Linear Models (GLM), Neural Networks (NN), Generalized Additive Models (GAMs), and Random Forests (RF).
+We chose to use the Random Forest algorithm in part because \citet{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) but also because of its ease of implementation and interpretation.
+We fit RF models using the Python package scikit-learn \citep{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model.
+We subsequently “tuned” each model using k-fold cross validation to optimize the size and splitting criteria of each tree.
+
+\subsection{Data description}
+
+We compiled over two decades of observational data (2000-2021) on salinity, temperature, and turbidity.
+Although this data came from a wide variety of sources, almost $80\%$ of the total unique observations came from three programs NDBC/CBIBS (\url{https://buoybay.noaa.gov/}), Maryland Continuous Monitoring Program (MDNR), and the Virginia Estuarine and Coastal Observing System (\url{http://vecos.vims.edu}).
+Because Chesapeake Bay has a mean optical depth of 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered collected observations to those occurring in the top 1m of the water column.
+We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours).
+Finally, we implemented spatial and temporal averaging of observational data to a daily mean within each MODIS pixel.
+We supplemented observational data with a grid of “freshwater influence” (FWI) estimates where each grid cell represents the average of flow weighted distance of that cell from the 5 major tributaries (Susquehanna, James, Pautexent, Choptank, and Potomac).
+Rather than straight-line distances we used “in-water” distances calculated using the scikit-image Python package \citep{vanderwaltScikitimageImageProcessing2014}.
+We combined the FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands values for each pixel spanning from 405 to 877nm wavelength \citep{vermoteericMOD09GAMODISTerra2015}.
+
+\subsection{Model validation}
+
+We validated RF model predictions both quantitatively against a hold-out test set of observational data (one third of the total data) and qualitatively against data from the Chesapeake Bay Operational Forecast System (CBOFS, \citet{lanerolle2011second}).
+CBOFS provides data on Bay-wide water level, temperature, and salinity every 6 hours.
+The model itself uses National Oceanic and Atmospheric Admistration (NOAA) water level, salinity observations, and US Geological Survey (USGS) river discharge data to calibrate an implementation of the Rutgers University Regional Ocean Modeling System (ROMS, \citet{shchepetkinRegionalOceanicModeling2005}).
 
@@
 
+\FloatBarrier
 \bibliographystyle{apalike}

### 2022-09-12 figures/rf_stats_table.py
format and include rmse table

@@ -10,17 +10,17 @@ utils.tabulate_to_latex(
     tabulate(
         dt.values, headers=[x for x in dt.columns], tablefmt="latex"
     ),
-    "figures/rf_stats_table_0.tex",
+    "figures/rf_stats_table_0_.tex",
     "Random Forest tuning results.",
     2,
 )
 
-dt = pd.read_csv("data/rmse_rf.csv")
+dt = pd.read_csv("data/rmse_rf.csv").round(2)
 utils.tabulate_to_latex(
     tabulate(
         dt.values, headers=[x for x in dt.columns], tablefmt="latex"
     ),
-    "figures/rf_stats_table_1.tex",
-    "Random Forest fitting results.",
-    1,
+    "figures/rf_stats_table_1_.tex",
+    "Random Forest fitting results. Values shown are root mean square error.",
+    0,
 )

### 2022-09-13 manuscript/manuscript.tex
format and include rmse table
text edits

@@
 Chesapeake Bay is located in the Mid-Atlantic region of the United States and serves as the outlet of 9 major tributary systems.
-Of these, the Susquehanna River provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, Xu et al.
-2012).
-The Bay is fairly shallow ($<$ 30m, \citet{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed (Xu et al.
-2012).
+Of these, the Susquehanna River provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \citet{xuClimateForcingSalinity2012}).
+The Bay is fairly shallow ($<$ 30m, \citet{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \citep{xuClimateForcingSalinity2012}.
 
@@
 We chose to use the Random Forest algorithm in part because \citet{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) but also because of its ease of implementation and interpretation.
-We fit RF models using the Python package scikit-learn \citep{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model.
+We fit RF models using the scikit-learn Python package \citep{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model.
 We subsequently “tuned” each model using k-fold cross validation to optimize the size and splitting criteria of each tree.
+We evaluated the “importance” of a particular variable as the average impurity decrease of tree splits based on that variable \citep{pedregosaScikitlearnMachineLearning2011}.
 
@@
 
-We compiled over two decades of observational data (2000-2021) on salinity, temperature, and turbidity.
-Although this data came from a wide variety of sources, almost $80\%$ of the total unique observations came from three programs NDBC/CBIBS (\url{https://buoybay.noaa.gov/}), Maryland Continuous Monitoring Program (MDNR), and the Virginia Estuarine and Coastal Observing System (\url{http://vecos.vims.edu}).
+We compiled over two decades of observational data (2000-2021) on salinity (n=6k), temperature (n=10k), and turbidity (n=5k, Table S1).
+Although this data came from a wide variety of sources (n=20, Table S1), almost $80\%$ of the total unique observations came from three programs NDBC/CBIBS (\url{https://buoybay.noaa.gov/}), Maryland Continuous Monitoring Program (MDNR), and the Virginia Estuarine and Coastal Observing System (\url{http://vecos.vims.edu}).
 Because Chesapeake Bay has a mean optical depth of 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered collected observations to those occurring in the top 1m of the water column.
@@
 CBOFS provides data on Bay-wide water level, temperature, and salinity every 6 hours.
-The model itself uses National Oceanic and Atmospheric Admistration (NOAA) water level, salinity observations, and US Geological Survey (USGS) river discharge data to calibrate an implementation of the Rutgers University Regional Ocean Modeling System (ROMS, \citet{shchepetkinRegionalOceanicModeling2005}).
+The model itself uses National Oceanic and Atmospheric Administration (NOAA) water level, salinity observations, and US Geological Survey (USGS) river discharge data to calibrate an implementation of the Rutgers University Regional Ocean Modeling System (ROMS, \citet{shchepetkinRegionalOceanicModeling2005}).
 
@@
           \caption{Comparison between Random Forest prediction results and a CBOFS snapshot for Sept, 4, 2022.}
+    \end{center}    
+\end{figure}
+
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/rf_stats_table_1_}          
     \end{center}    

### 2022-09-13 manuscript/supplement.tex
pipeline-ify ms and supp
add validation plot to supplement
cleanup best_params table

@@
     \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{../figures/counts_obs_table}
+          \includegraphics[width=0.85\textwidth,keepaspectratio]{figures/counts_obs_table}
+    \end{center}    
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+          \includegraphics[width=0.7\textwidth,keepaspectratio]{figures/_validation}
+          \caption{RF predicted salinity versus test set (observed) salinity. Hatching indicates the area containing 80\% of the data points.}
+    \end{center}    
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+          \includegraphics[width=1\textwidth,keepaspectratio]{figures/rf_stats_table_0_}
     \end{center}    

### 2022-09-13 figures/rf_stats_table.py

@@ -6,13 +6,16 @@ sys.path.append(".")
 from src import utils
 
 dt = pd.read_csv("data/best_params_rf.csv")
+dt = dt.drop(columns=["bootstrap", "min_samples_split"])
+variables = dt.pop("variable")
+dt.insert(0, "variable", variables)
 utils.tabulate_to_latex(
     tabulate(
         dt.values, headers=[x for x in dt.columns], tablefmt="latex"
     ),
     "figures/rf_stats_table_0_.tex",
     "Random Forest tuning results.",
-    2,
+    1,
 )
 
 dt = pd.read_csv("data/rmse_rf.csv").round(2)

### 2022-09-13 figures/validation.py
enable validation plots for all variables

@@ -7,43 +7,61 @@ from sklearn.linear_model import LinearRegression
 
 sys.path.append(".")
 from src import rf_icom_call_data2 as call_data2
+from src import utils
+
+def _get_predictions(variable):
+    rf_random_path = "data/rf_random_" + variable + ".pkl"
+    rf_random = pickle.load(open(rf_random_path, "rb"))
+
+    predictors = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
+
+    X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
+    y_test = pickle.load(open("data/y_test_" + variable + ".pkl", "rb"))
+
+    predictions = rf_random.predict(X_test)
+    res = pd.DataFrame(y_test, columns=["obs"])
+    res["predict"] = predictions.copy()
+    return res
+
+variables = ["SSS (psu)", "turbidity (NTU)", "SST (C)"]
+variables_str = ["salinity (psu)", "turbidity (NTU)", "temperature (C)"]
+variables_str_short = [utils.clean_var_name(v) for v in variables_str]
 
 
 # scatter plot of measured vs predicted
-variable = "salinity"
-var_col = "SSS (psu)"
+res = [_get_predictions(variable) for variable in variables_str_short]
 
-rf_random_path = "data/rf_random_" + variable + ".pkl"
-rf_random = pickle.load(open(rf_random_path, "rb"))
+def _plot(ax, dt, xy_min=None, xy_max=None, bins=100, pthresh=0.2, hatching=True):
+    # dt = res[1]
+    if xy_max is None:
+        xy_max = max(max(dt["obs"]), max(dt["predict"]))
+    if xy_min is None:
+        xy_min = min(min(dt["obs"]), min(dt["predict"]))
 
-predictors = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
+    g = sns.scatterplot(data=dt, x="predict", y="obs", ax=ax, s=7, color=".15")
+    # increasing bin makes coarser boxes, increasing pthresh makes less boxes
+    if hatching:
+        sns.histplot(data=dt, x="predict", y="obs", bins=bins, pthresh=pthresh,cmap="mako", ax=ax)
[…]

### 2022-09-13 figures/validation.py
formatting with yapf

@@ -9,11 +9,10 @@ sys.path.append(".")
 from src import rf_icom_call_data2 as call_data2
 from src import utils
 
+
 def _get_predictions(variable):
     rf_random_path = "data/rf_random_" + variable + ".pkl"
     rf_random = pickle.load(open(rf_random_path, "rb"))
-
-    predictors = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
 
     X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
     y_test = pickle.load(open("data/y_test_" + variable + ".pkl", "rb"))
@@ -23,15 +22,22 @@ def _get_predictions(variable):
     res["predict"] = predictions.copy()
     return res
 
+
 variables = ["SSS (psu)", "turbidity (NTU)", "SST (C)"]
 variables_str = ["salinity (psu)", "turbidity (NTU)", "temperature (C)"]
 variables_str_short = [utils.clean_var_name(v) for v in variables_str]
 
-
 # scatter plot of measured vs predicted
 res = [_get_predictions(variable) for variable in variables_str_short]
 
-def _plot(ax, dt, xy_min=None, xy_max=None, bins=100, pthresh=0.2, hatching=True):
+
+def _plot(ax,
+          dt,
+          xy_min=None,
+          xy_max=None,
+          bins=100,
+          pthresh=0.2,
+          hatching=True):
     # dt = res[1]
     if xy_max is None:
         xy_max = max(max(dt["obs"]), max(dt["predict"]))
@@ -41,21 +47,29 @@ def _plot(ax, dt, xy_min=None, xy_max=None, bins=100, pthresh=0.2, hatching=True
     g = sns.scatterplot(data=dt, x="predict", y="obs", ax=ax, s=7, color=".15")
     # increasing bin makes coarser boxes, increasing pthresh makes less boxes
     if hatching:
-        sns.histplot(data=dt, x="predict", y="obs", bins=bins, pthresh=pthresh,cmap="mako", ax=ax)
-    plt.plot([0, xy_max-4], [0, xy_max-4], color="red")
+        sns.histplot(data=dt,
+                     x="predict",
+                     y="obs",
+                     bins=bins,
+                     pthresh=pthresh,
+                     cmap="mako",
+                     ax=ax)
[…]

### 2022-09-13 src/fit_sine.py
cleanup

@@ -1,3 +1,5 @@
+# https://scipy-cookbook.readthedocs.io/items/FittingData.html
+
 import numpy as np
 import pandas as pd
 from scipy import optimize
@@ -5,48 +7,32 @@ import matplotlib.pyplot as plt
 
 
 def fit_sine(data):
-
-    ##Read in Data
-    #    data=pd.read_csv('temperature.csv')
-    #    data.replace([np.inf, -np.inf], np.nan, inplace=True)
+    # data = pd.read_csv('temperature.csv')
+    # data.replace([np.inf, -np.inf], np.nan, inplace=True)
     # data = pd.read_csv("aggregated_w_bandvals.csv")
+
     data['datetime'] = pd.to_datetime(data['datetime'])
     data['datetime'] = data['datetime'].astype('datetime64')
     data['datetime'] = data['datetime'].dt.strftime('%j')
     data['datetime'] = pd.to_numeric(data["datetime"], downcast="float")
     date = data['datetime']
-    #    diff_time=time_data[1:,]-time_data[:-1,]
-    ##Get features
-    temp = data['SST (C)'].values
 
     temp_doy = []
     for i in np.arange(1, 367):
         temp_mean = np.mean(data['SST (C)'][data['datetime'] == i])
         temp_doy.append(temp_mean)
-    plt.scatter(np.arange(1, 367), temp_doy)
-    #    temp_doy=np.vstack((temp_doy, temp_doy))
     date_doy = np.arange(1, 367)
-    #    date_doy=np.arange(0,366*2)
 
-    ##https://scipy-cookbook.readthedocs.io/items/FittingData.html
-
-    # Fit the first set
-    #fitfunc = lambda p, x: p[0]*np.sin(2*np.pi/p[1]*x+p[2])+p[3] # Target function
     fitfunc = lambda p, x: p[0] * np.sin(2 * np.pi / p[1] * x + p[2]) + p[3]
     errfunc = lambda p, x, y: fitfunc(p, x
                                       ) - y  # Distance to the target function
-    #    p0 = [1.21877004e+01, 3.15741655e+16, 2.95421427e+01, 1.65708992e+01] # Initial guess for the parameters
     p0 = [12.23414294, 368.06372131, 29.41529669, 16.38386361]
-    #p1, success = optimize.leastsq(errfunc, p0[:], args=(date, temp))
[…]

### 2022-09-13 src/fit_sine.py
towards backing out non-sine adjusted temperature

@@ -1,15 +1,17 @@
 # https://scipy-cookbook.readthedocs.io/items/FittingData.html
 
+import pickle
 import numpy as np
 import pandas as pd
 from scipy import optimize
 import matplotlib.pyplot as plt
 
+fitfunc = lambda p, x: p[0] * np.sin(2 * np.pi / p[1] * x + p[2]) + p[3]    
 
 def fit_sine(data):
     # data = pd.read_csv('temperature.csv')
     # data.replace([np.inf, -np.inf], np.nan, inplace=True)
-    # data = pd.read_csv("aggregated_w_bandvals.csv")
+    # data = pd.read_csv("data/aggregated_w_bandvals.csv")
 
     data['datetime'] = pd.to_datetime(data['datetime'])
     data['datetime'] = data['datetime'].astype('datetime64')
@@ -22,14 +24,15 @@ def fit_sine(data):
         temp_mean = np.mean(data['SST (C)'][data['datetime'] == i])
         temp_doy.append(temp_mean)
     date_doy = np.arange(1, 367)
-
-    fitfunc = lambda p, x: p[0] * np.sin(2 * np.pi / p[1] * x + p[2]) + p[3]
+    
     errfunc = lambda p, x, y: fitfunc(p, x
                                       ) - y  # Distance to the target function
     p0 = [12.23414294, 368.06372131, 29.41529669, 16.38386361]
-    p1, success = optimize.leastsq(errfunc, p0[:], args=(date_doy, temp_doy))
-    # print(p1)
+    p1, success = optimize.leastsq(errfunc, p0[:], args=(date_doy, temp_doy))    
     # print(success)
+    # print(p1)    
+    pickle.dump(p1, open("data/temperature_sine_coef.pkl", "wb"))
+
     fitted_sine = fitfunc(p1, date)
     
     plt.figure(figsize=(20, 10))

### 2022-09-13 src/rf_icom_call_data2.py

@@ -35,7 +35,7 @@ def clean_data(variable, var_col, predictors, test_size=0.5, data=None):
         "cost", "latitude", "longitude"
     ]
     X_train, y_train, X_test, y_test, feature_names = call_data2.clean_data(
-        "salinity", "SSS (psu)", predictors
+        "temperature", "SST (C)", predictors
     )
     ```
     """    
@@ -83,7 +83,6 @@ def clean_data(variable, var_col, predictors, test_size=0.5, data=None):
     data["datetime"] = pd.to_numeric(data["datetime"], downcast="float")
 
     if variable == "temperature":
-        print(data["SST (C)"])
         fitted_sine = fit_sine.fit_sine(data)
         # print('fitted_sine')
         # print(fitted_sine)
@@ -97,10 +96,6 @@ def clean_data(variable, var_col, predictors, test_size=0.5, data=None):
     ## Convert to Day of Year
     data["datetime"] = utils.datetime_to_doy(data["datetime"])
 
-    #    print(len(data))
-    #    data=data.drop(data['turbidity (NTU)']<0)
-    #    print(len(data))
-
     ## Get features
     y = data[var_col].values
     X = data.drop(var_col, axis=1)
@@ -108,6 +103,7 @@ def clean_data(variable, var_col, predictors, test_size=0.5, data=None):
     lon_list = [lon for lon in X.longitude.values]
     lat_list = [lon for lon in X.latitude.values]
 
+    print(X.shape)
     X = X.values
     feature_names = [k for k in data.keys() if k in predictors]
 

### 2022-09-13 scripts/01_rf_fit.py
save prerfe test data

@@ -36,6 +36,10 @@ def main():
     X_train, y_train, X_test, y_test, feature_names = call_data2.clean_data(
         variable, var_col, predictors, test_size=0.33, data=data)
 
+    pickle.dump(X_test, open("data/X_test_prerfe_" + variable + ".pkl", "wb"))
+    pickle.dump(y_test, open("data/y_test_prerfe_" + variable + ".pkl", "wb"))
+    pickle.dump(feature_names, open("data/feature_names_" + variable + ".pkl", "wb"))
+
     ##Feature Selection vs Hyperparameter Tuning: https://stats.stackexchange.com/questions/264533/how-should-feature-selection-and-hyperparameter-optimization-be-ordered-in-the-m
     ##I added n_estimators=1000, and a max_depth =20 for the hyperparameters used in feature selection
 

### 2022-09-13 manuscript/manuscript.tex
text edits

@@
 Although this data came from a wide variety of sources (n=20, Table S1), almost $80\%$ of the total unique observations came from three programs NDBC/CBIBS (\url{https://buoybay.noaa.gov/}), Maryland Continuous Monitoring Program (MDNR), and the Virginia Estuarine and Coastal Observing System (\url{http://vecos.vims.edu}).
-Because Chesapeake Bay has a mean optical depth of 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered collected observations to those occurring in the top 1m of the water column.
+Because Chesapeake Bay has a mean optical depth of 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column.
 We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours).
+Prior to model fitting, we removed the annual cycle from temperature observations using least squares optimization routines \citep{virtanenSciPyFundamentalAlgorithms2020a}.
 Finally, we implemented spatial and temporal averaging of observational data to a daily mean within each MODIS pixel.
+
 We supplemented observational data with a grid of “freshwater influence” (FWI) estimates where each grid cell represents the average of flow weighted distance of that cell from the 5 major tributaries (Susquehanna, James, Pautexent, Choptank, and Potomac).
@@
 \section{Results}
+
+Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2, and 1.8 respectively (Table 1, Figure S1).
+The “size” (i.e.
+n\_estimators) of the final RF models were of a similar magnitude for each variable but the tree was much “deeper” (i.e.
+max\_depth) for temperature compared to the other variables (Table S2).
+Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance towards each tributary (Figure 1).
 

### 2022-09-14 src/rf_icom_call_data2.py
umm, the date field is already in doy, this is setting it to all 1s

@@ -94,7 +94,8 @@ def clean_data(variable, var_col, predictors, test_size=0.5, data=None):
     #    plt.show()
 
     ## Convert to Day of Year
-    data["datetime"] = utils.datetime_to_doy(data["datetime"])
+    # data["datetime"] = utils.datetime_to_doy(data["datetime"])
+    # ummm, the data is already in doy?
 
     ## Get features
     y = data[var_col].values

### 2022-09-15 figures/validation.py
paneled validation plot for all vars

@@ -6,11 +6,12 @@ import matplotlib.pyplot as plt
 from sklearn.linear_model import LinearRegression
 
 sys.path.append(".")
-from src import rf_icom_call_data2 as call_data2
 from src import utils
+from src import fit_sine
 
 
 def _get_predictions(variable):
+    # variable = "temperature"
     rf_random_path = "data/rf_random_" + variable + ".pkl"
     rf_random = pickle.load(open(rf_random_path, "rb"))
 
@@ -20,6 +21,14 @@ def _get_predictions(variable):
     predictions = rf_random.predict(X_test)
     res = pd.DataFrame(y_test, columns=["obs"])
     res["predict"] = predictions.copy()
+
+    if variable == "temperature":
+        # back out non-sine adjusted values
+        p1 = pickle.load(open("data/temperature_sine_coef.pkl", "rb"))
+        offset = fit_sine.fitfunc(p1, X_test[0:, 0])
+        res["predict"] = res["predict"].copy() + offset
+        res["obs"] = res["obs"].copy() + offset
+
     return res
 
 
@@ -33,16 +42,22 @@ res = [_get_predictions(variable) for variable in variables_str_short]
 
 def _plot(ax,
           dt,
+          title,
           xy_min=None,
           xy_max=None,
           bins=100,
           pthresh=0.2,
-          hatching=True):
+          hatching=True,
+          text_anchor=(3, 27),
+          text_space=2,
+          ylab=False):
     # dt = res[1]
     if xy_max is None:
         xy_max = max(max(dt["obs"]), max(dt["predict"]))
+        print(xy_max)
     if xy_min is None:
         xy_min = min(min(dt["obs"]), min(dt["predict"]))
+        print(xy_min)
 
     g = sns.scatterplot(data=dt, x="predict", y="obs", ax=ax, s=7, color=".15")
     # increasing bin makes coarser boxes, increasing pthresh makes less boxes
@@ -52,30 +67,38 @@ def _plot(ax,
                      y="obs",
                      bins=bins,
                      pthresh=pthresh,
-                     cmap="mako",
+                     color='black',
+                     hue_order="black",
+                     pmax=0,
[…]

### 2022-09-15 src/fit_sine.py

@@ -6,7 +6,7 @@ import pandas as pd
 from scipy import optimize
 import matplotlib.pyplot as plt
 
-fitfunc = lambda p, x: p[0] * np.sin(2 * np.pi / p[1] * x + p[2]) + p[3]    
+fitfunc = lambda p, x: p[0] * np.sin(2 * np.pi / p[1] * x + p[2]) + p[3]
 
 def fit_sine(data):
     # data = pd.read_csv('temperature.csv')

### 2022-09-15 figures/validation.py
formatting with yapf

@@ -85,20 +85,35 @@ def _plot(ax,
     reg = LinearRegression().fit(X, y)
 
     eq = "y = " + str(round(reg.coef_[0], 2)) + "x - " + str(
-        round(abs(reg.intercept_), 2))    
+        round(abs(reg.intercept_), 2))
     ax.text(text_anchor[0], text_anchor[1], eq, size=8)
-    ax.text(text_anchor[0], text_anchor[1]-text_space, "$R^2$ = " + str(round(reg.score(X, y), 2)), size=8)
+    ax.text(text_anchor[0],
+            text_anchor[1] - text_space,
+            "$R^2$ = " + str(round(reg.score(X, y), 2)),
+            size=8)
     ax.set_title(title)
-    
+
     return ax
 
 
 plt.close()
 fig, axes = plt.subplots(figsize=(8, 4), ncols=3)
 _plot(axes[0], res[0], "Salinity", -0.5, 32, ylab=True)
-_plot(axes[1], res[1], "Turbidity", xy_max=75, xy_min=0, hatching=False, text_anchor=(37, 10), text_space=4.5)
-_plot(axes[2], res[2], "Temperature", xy_min=-2, xy_max=35, text_anchor=(3, 30), text_space=2)
+_plot(axes[1],
+      res[1],
+      "Turbidity",
+      xy_max=75,
+      xy_min=0,
+      hatching=False,
+      text_anchor=(37, 10),
+      text_space=4.5)
+_plot(axes[2],
+      res[2],
+      "Temperature",
+      xy_min=-2,
+      xy_max=35,
+      text_anchor=(3, 30),
+      text_space=2)
 # plt.show()
-
 
 plt.savefig("figures/_validation.pdf")

### 2022-09-15 figures/validation.py
save r2

@@ -86,34 +86,36 @@ def _plot(ax,
 
     eq = "y = " + str(round(reg.coef_[0], 2)) + "x - " + str(
         round(abs(reg.intercept_), 2))
+    r2 = str(round(reg.score(X, y), 2))
     ax.text(text_anchor[0], text_anchor[1], eq, size=8)
     ax.text(text_anchor[0],
             text_anchor[1] - text_space,
-            "$R^2$ = " + str(round(reg.score(X, y), 2)),
+            "$R^2$ = " + r2,
             size=8)
     ax.set_title(title)
 
-    return ax
+    return ax, r2
 
 
 plt.close()
+r2 = []
 fig, axes = plt.subplots(figsize=(8, 4), ncols=3)
-_plot(axes[0], res[0], "Salinity", -0.5, 32, ylab=True)
-_plot(axes[1],
+r2.append(_plot(axes[0], res[0], "Salinity", -0.5, 32, ylab=True)[1])
+r2.append(_plot(axes[1],
       res[1],
       "Turbidity",
       xy_max=75,
       xy_min=0,
       hatching=False,
       text_anchor=(37, 10),
-      text_space=4.5)
-_plot(axes[2],
+      text_space=4.5)[1])
+r2.append(_plot(axes[2],
       res[2],
       "Temperature",
       xy_min=-2,
       xy_max=35,
       text_anchor=(3, 30),
-      text_space=2)
+      text_space=2)[1])
 # plt.show()
 
 plt.savefig("figures/_validation.pdf")

### 2022-09-15 figures/validation.py
fix validation plot aspect ratio

@@ -93,13 +93,19 @@ def _plot(ax,
             "$R^2$ = " + r2,
             size=8)
     ax.set_title(title)
+    ax.set_aspect("equal")
 
     return ax, r2
 
+# # sine-adjusted temperature plot
+# plt.close()
+# fig, ax = plt.subplots(figsize=(8, 4))
+# _plot(ax, res[2], title="asdf", xy_max=10, xy_min=-10)
+# plt.show()
 
 plt.close()
 r2 = []
-fig, axes = plt.subplots(figsize=(8, 4), ncols=3)
+fig, axes = plt.subplots(figsize=(8, 4.5), ncols=3)
 r2.append(_plot(axes[0], res[0], "Salinity", -0.5, 32, ylab=True)[1])
 r2.append(_plot(axes[1],
       res[1],

### 2022-09-15 figures/importance.py
cleanup

@@ -17,11 +17,11 @@ def imp_plot(variable):
     dset = dset.sort_values(by="importance", ascending=False)
     dset.loc[dset["attr"] == "cost", "attr"] = "fwi"
 
-    plt.figure(figsize=(5.7, 5))
+    plt.figure(figsize=(3, 5))
 
     plt.barh(y=dset["attr"], width=dset["importance"], color="#1976D2")
-    plt.title(variable, fontsize=16, fontweight="bold", pad=20)
-    plt.xlabel("Importance", fontsize=14, labelpad=20)
+    plt.title(variable, fontsize=16, pad=20)
+    plt.xlabel("Importance", fontsize=14, labelpad=10)
 
     # plt.show()
     plt.savefig("figures/" + variable + "_importance.pdf", bbox_inches='tight')

### 2022-09-15 figures/rf_stats_table.py
add r2 to rf_stats table

@@ -19,6 +19,11 @@ utils.tabulate_to_latex(
 )
 
 dt = pd.read_csv("data/rmse_rf.csv").round(2)
+dt = dt.T.reset_index().rename(columns={"index":"variable", 0:"rmse"})
+col_names = dt["variable"].copy()
+dt = dt.merge(pd.read_csv("data/r2.csv"), on="variable").T[1:3].reset_index()
+dt.columns = ["metric"] + [x for x in col_names]
+
 utils.tabulate_to_latex(
     tabulate(
         dt.values, headers=[x for x in dt.columns], tablefmt="latex"

### 2022-09-15 figures/validation.py

@@ -106,22 +106,26 @@ def _plot(ax,
 plt.close()
 r2 = []
 fig, axes = plt.subplots(figsize=(8, 4.5), ncols=3)
-r2.append(_plot(axes[0], res[0], "Salinity", -0.5, 32, ylab=True)[1])
+r2.append(_plot(axes[0], res[0], "salinity", -0.5, 32, ylab=True, text_space=2.5)[1])
 r2.append(_plot(axes[1],
       res[1],
-      "Turbidity",
+      "turbidity",
       xy_max=75,
       xy_min=0,
       hatching=False,
       text_anchor=(37, 10),
-      text_space=4.5)[1])
+      text_space=5)[1])
 r2.append(_plot(axes[2],
       res[2],
-      "Temperature",
+      "temperature",
       xy_min=-2,
       xy_max=35,
       text_anchor=(3, 30),
-      text_space=2)[1])
+      text_space=2.5)[1])
 # plt.show()
+
+r2_df = pd.DataFrame(r2, columns=["r2"])
+r2_df["variable"] = variables_str_short
+r2_df.to_csv("data/r2.csv", index=False)
 
 plt.savefig("figures/_validation.pdf")

### 2022-09-15 manuscript/manuscript.tex
move validation plot from supp to ms
text edits

@@
 
-Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2, and 1.8 respectively (Table 1, Figure S1).
+Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure S1).
+Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1).
+One area where the temperature model had poor performance was at the extremes, it was unable to recover very low values ($<$ 0) and very high values ($>$ 30, Figure 1).
+
 The “size” (i.e.
@@
 max\_depth) for temperature compared to the other variables (Table S2).
-Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance towards each tributary (Figure 1).
+
+Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 2).
+
+The date of observation was the most important feature in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 3).
+Across all models, the most important MODIS band was band 8.
+
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
+          \caption{Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.}
+    \end{center}    
+\end{figure}
 
@@
     \begin{center}
-          \includegraphics[width=0.95\textwidth,keepaspectratio]{figures/_importance_all}
+          \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_importance_all}
           \caption{Random Forest importance plot for salinity, temperature, and turbidity models.}

### 2022-09-16 figures/maps.py
functionalize getting prediction images

@@ -1,10 +1,10 @@
 import os
 import sys
 import rioxarray
+import subprocess
 import numpy as np
 import pandas as pd
 import xarray as xr
-from osgeo import gdal
 import geopandas as gpd
 import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
@@ -78,7 +78,7 @@ cbar.ax.set_title("obs count (n)", y=-0.08)
 # plt.show()
 plt.savefig("figures/_freqcount_hex.pdf")
 
-# --- input/output map
+# --- rf vs cbofs map
 bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
     epsg=4326)
 
@@ -92,16 +92,25 @@ img_cbofs = img_cbofs["band_data"].sel(band=1)
 # plt.show()
 
 # get rf prediction image
-# gdalwarp -te -77.3425000000000011 36.1675000000000040 -74.7974999999999994 39.6325000000000074 -ts 509 693 -overwrite 2022-09-04.tif 2022-09-04_downsample.tif
+def get_rf_prediction(date):
+    # date = "2022-09-04"
 
-date = "2022-09-04"
-img_rf = xr.open_dataset("data/prediction/" + date + "_downsample.tif",
-                         engine="rasterio")
-img_rf = img_rf.rio.clip(bay_gdf_hires.geometry)
-img_rf = img_rf["band_data"].sel(band=1)
-img_rf.rio.to_raster("data/prediction/" + date + "_downsample_clip.tif")
-# img_rf.plot.imshow()
-# plt.show()
+    bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
+    epsg=4326)
+
+    call_string = "gdalwarp -te -77.3425000000000011 36.1675000000000040 -74.7974999999999994 39.6325000000000074 -ts 509 693 -overwrite data/prediction/" + date + ".tif data/prediction/" + date + "_downsample.tif"
+    subprocess.call(call_string)
+
+    img_rf = xr.open_dataset("data/prediction/" + date + "_downsample.tif",
+                            engine="rasterio")
+    img_rf = img_rf.rio.clip(bay_gdf_hires.geometry)
+    img_rf = img_rf["band_data"].sel(band=1)
+    img_rf.rio.to_raster("data/prediction/" + date + "_downsample_clip.tif")
+    # img_rf.plot.imshow()
[…]

### 2022-09-16 figures/maps.py
create seasonality maps

@@ -1,17 +1,41 @@
 import os
 import sys
 import rioxarray
-import subprocess
 import numpy as np
 import pandas as pd
 import xarray as xr
 import geopandas as gpd
 import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
+from matplotlib import gridspec
 import h3pandas  # h3.geo_to_h3_aggregate
 
 sys.path.append(".")
 from src import utils
+
+def panel_add(i, axs, title, geo_grid, diff=False, j=None, vmax=27, height_frac=0.5):
+    if j is not None:        
+        ax= plt.subplot(axs[i,j], xlabel="", projection = ccrs.PlateCarree())
+        ax.coastlines(resolution="10m", color="black", linewidth=1)
+    else:
+        ax = axs[i]
+        ax.coastlines(resolution="10m", color="black", linewidth=1)
+    if diff:
+        geo_grid.plot.imshow(ax=ax,
+                             center=0,
+                             cbar_kwargs={
+                                 "shrink": 0.5,
+                                 'label': ''
+                             })
+    else:
+        geo_grid.plot.imshow(ax=ax,
+                             vmin=0,
+                             vmax=vmax,
+                             cbar_kwargs={
+                                 "shrink": 0.5,
+                                 'label': ''
+                             })  # , vmax=np.nanmax(img_cbofs.to_numpy()))    
+    ax.set_title(title, size="small", y=height_frac, x=0.9, rotation="vertical")
 
 
 # --- loc_id frequency count histograms and hex map
@@ -91,26 +115,7 @@ img_cbofs = img_cbofs["band_data"].sel(band=1)
 # img_cbofs.plot.imshow()
 # plt.show()
 
-# get rf prediction image
-def get_rf_prediction(date):
-    # date = "2022-09-04"
-
-    bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
-    epsg=4326)
-
-    call_string = "gdalwarp -te -77.3425000000000011 36.1675000000000040 -74.7974999999999994 39.6325000000000074 -ts 509 693 -overwrite data/prediction/" + date + ".tif data/prediction/" + date + "_downsample.tif"
-    subprocess.call(call_string)
-
[…]

### 2022-09-16 src/utils.py

@@ -1,6 +1,12 @@
+import os
 import re
+import subprocess
 import numpy as np
 import pandas as pd
+import xarray as xr
+import geopandas as gpd
+import matplotlib.pyplot as plt
+
 
 
 def select_var(dt, var_col):
@@ -95,3 +101,23 @@ def tabulate_to_latex(tabulate_output, fname, caption='nasdf', table_num=1):
 def modisaqua_path(date, band="sur_refl_b08"):
     # date='2018-01-01'
     return "data/MODIS-Aqua/{date}_{band}.tif".format(date=date, band=band)
+
+def get_rf_prediction(date):
+    # date = "2022-09-04"
+
+    bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
+    epsg=4326)
+
+    path_downsample = "data/prediction/" + date + "_downsample.tif"
+    if not os.path.exists(path_downsample):
+        call_string = "gdalwarp -te -77.3425000000000011 36.1675000000000040 -74.7974999999999994 39.6325000000000074 -ts 509 693 -overwrite data/prediction/" + date + ".tif data/prediction/" + date + "_downsample.tif"
+        subprocess.call(call_string)
+
+    img_rf = xr.open_dataset("data/prediction/" + date + "_downsample.tif",
+                            engine="rasterio")
+    img_rf = img_rf.rio.clip(bay_gdf_hires.geometry)
+    img_rf = img_rf["band_data"].sel(band=1)
+    img_rf.rio.to_raster("data/prediction/" + date + "_downsample_clip.tif")
+    # img_rf.plot.imshow()
+    # plt.show()
+    return img_rf

### 2022-09-16 figures/maps.py
formatting with yapf

@@ -13,9 +13,17 @@ import h3pandas  # h3.geo_to_h3_aggregate
 sys.path.append(".")
 from src import utils
 
-def panel_add(i, axs, title, geo_grid, diff=False, j=None, vmax=27, height_frac=0.5):
-    if j is not None:        
-        ax= plt.subplot(axs[i,j], xlabel="", projection = ccrs.PlateCarree())
+
+def panel_add(i,
+              axs,
+              title,
+              geo_grid,
+              diff=False,
+              j=None,
+              vmax=27,
+              height_frac=0.5):
+    if j is not None:
+        ax = plt.subplot(axs[i, j], xlabel="", projection=ccrs.PlateCarree())
         ax.coastlines(resolution="10m", color="black", linewidth=1)
     else:
         ax = axs[i]
@@ -34,8 +42,12 @@ def panel_add(i, axs, title, geo_grid, diff=False, j=None, vmax=27, height_frac=
                              cbar_kwargs={
                                  "shrink": 0.5,
                                  'label': ''
-                             })  # , vmax=np.nanmax(img_cbofs.to_numpy()))    
-    ax.set_title(title, size="small", y=height_frac, x=0.9, rotation="vertical")
+                             })  # , vmax=np.nanmax(img_cbofs.to_numpy()))
+    ax.set_title(title,
+                 size="small",
+                 y=height_frac,
+                 x=0.9,
+                 rotation="vertical")
 
 
 # --- loc_id frequency count histograms and hex map
@@ -165,11 +177,15 @@ imgs = [utils.get_rf_prediction(date) for date in dates]
 nrow = 2
 ncol = 2
 
-fig = plt.figure(figsize=(ncol+3, nrow+3))
-axs = gridspec.GridSpec(nrow, ncol,
-         wspace=0.0, hspace=0.0, 
-         top=1.-0.5/(nrow+1), bottom=0.5/(nrow+1), 
-         left=0.1666, right=0.7) # 0.83333
+fig = plt.figure(figsize=(ncol + 3, nrow + 3))
+axs = gridspec.GridSpec(nrow,
+                        ncol,
+                        wspace=0.0,
+                        hspace=0.0,
+                        top=1. - 0.5 / (nrow + 1),
+                        bottom=0.5 / (nrow + 1),
+                        left=0.1666,
+                        right=0.7)  # 0.83333
 
 panel_add(0, axs, dates[0], imgs[0], j=0)
 panel_add(0, axs, dates[1], imgs[1], j=1)

### 2022-09-16 src/utils.py

@@ -8,7 +8,6 @@ import geopandas as gpd
 import matplotlib.pyplot as plt
 
 
-
 def select_var(dt, var_col):
     if isinstance(dt, str):
         dt = pd.read_csv(dt)
@@ -26,14 +25,9 @@ def select_var(dt, var_col):
 
 def freq_count(dt, grp_cols):
     dt = dt.copy()
-    dt = (
-        dt.groupby(grp_cols)
-        .size()
-        .reset_index()
-        .rename(columns={0: "count"})
-        .sort_values("count", ascending=False)
-        .reset_index(drop=True)
-    )
+    dt = (dt.groupby(grp_cols).size().reset_index().rename(columns={
+        0: "count"
+    }).sort_values("count", ascending=False).reset_index(drop=True))
     return dt
 
 
@@ -45,21 +39,17 @@ def load_md(variable, fpath_prestring="data/best_params_"):
     # load best_params_[variable].md
     bp_colnames = _unwrap(
         str(
-            pd.read_table(fpath_prestring + variable + ".md", nrows=1, header=None)[0][
-                0
-            ]
-        ).strip()
-    )[0]
+            pd.read_table(fpath_prestring + variable + ".md",
+                          nrows=1,
+                          header=None)[0][0]).strip())[0]
     if "max_leaf_nodes" in bp_colnames:
         bp_colnames.remove("max_leaf_nodes")
     bp = pd.DataFrame(
         _unwrap(
             str(
-                pd.read_table(
-                    fpath_prestring + variable + ".md", skiprows=2, header=None
-                )[0][0]
-            ).strip()
-        ),
+                pd.read_table(fpath_prestring + variable + ".md",
+                              skiprows=2,
+                              header=None)[0][0]).strip()),
         columns=bp_colnames,
     )
     return bp
@@ -78,6 +68,7 @@ def datetime_to_doy(dt):
     dt = dt.dt.strftime("%j")
     dt = pd.to_numeric(dt, downcast="float")
     return dt
+
 
 def tabulate_to_latex(tabulate_output, fname, caption='nasdf', table_num=1):
     with open(fname, "w") as f:
@@ -98,15 +89,17 @@ def tabulate_to_latex(tabulate_output, fname, caption='nasdf', table_num=1):
         f.write("\\end{table}" + "\r\n")
         f.write("\\end{document}" + "\r\n")
 
+
[…]

### 2022-09-16 figures/obs_stats.py
formally include the obs_distribution table

@@ -95,6 +95,13 @@ dt_agg = pd.read_csv(
     os.environ["ICOM_DATA"] + "/Modeling Data/Processed Data p1/aggregated.csv",
     parse_dates=["datetime"],
 )
+
+# per call_data2.py,
+# for turbidity get rid of negative numbers
+dt_agg.loc[dt_agg["turbidity (NTU)"] < 0, "turbidity (NTU)"] = None
+# for salinity set negative numbers to 0
+dt_agg.loc[dt_agg["SSS (psu)"] < 0, "SSS (psu)"] = 0
+
 obs_distribution = (
     pd.concat(
         [
@@ -114,3 +121,11 @@ with open("figures/obs_distribution.md", "w") as f:
         tabulate(obs_distribution.values, headers=[x for x in obs_distribution.columns])
     )
 # obs_distribution.to_clipboard()
+utils.tabulate_to_latex(
+    tabulate(
+        obs_distribution.values, headers=[x for x in obs_distribution.columns], tablefmt="latex"
+    ),
+    "figures/obs_distribution.tex",
+    "Distribution of values in the observational dataset.",
+    1,
+)

### 2022-09-16 figures/rf_stats_table.py

@@ -15,7 +15,7 @@ utils.tabulate_to_latex(
     ),
     "figures/rf_stats_table_0_.tex",
     "Random Forest tuning results.",
-    1,
+    2,
 )
 
 dt = pd.read_csv("data/rmse_rf.csv").round(2)

### 2022-09-16 manuscript/manuscript.tex
results edits

@@
 \SetWatermarkScale{5}
+\usepackage{float}
 
@@
 Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1).
-One area where the temperature model had poor performance was at the extremes, it was unable to recover very low values ($<$ 0) and very high values ($>$ 30, Figure 1).
-
-The “size” (i.e.
-n\_estimators) of the final RF models were of a similar magnitude for each variable but the tree was much “deeper” (i.e.
-max\_depth) for temperature compared to the other variables (Table S2).
-
-Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 2).
-
-The date of observation was the most important feature in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 3).
-Across all models, the most important MODIS band was band 8.
-
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
-          \caption{Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.}
-    \end{center}    
-\end{figure}
-
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
-          \caption{Comparison between Random Forest prediction results and a CBOFS snapshot for Sept, 4, 2022.}
-    \end{center}    
-\end{figure}
+Despite excellent overall performance, our temperature and salinity models had poor performance at the extremes.
+They were unable to recover very low values ($<$ 0 C) and very high values ($>$ 30 C, $>$ 23 salinity, Figure 1).
+These areas of poor performance can be explained by their relative rarity in the observational dataset (Table S2).
 
@@
     \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
+          \caption{Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.}
+    \end{center}    
+\end{figure}
+
+The date of observation was the most important feature in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 2).
+Across all models, the most important MODIS band was band 8.
+The “size” (i.e.
+n\_estimators) of the final RF models were of a similar magnitude for each variable but the tree was much “deeper” (i.e.
+max\_depth) for temperature compared to the other variables (Table S3).
+
+\begin{figure}[ht!]
+    \begin{center}
           \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_importance_all}
@@
 \end{figure}
+
+Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 3).
+RF model predictions were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 4).
+The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 2).
+
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.4\textwidth,keepaspectratio]{figures/_seasonality}
+          \caption{Seasonal Random Forest salinity prediction results.}
+    \end{center}    
+\end{figure}
+
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
+          \caption{Comparison between Random Forest salinity prediction results and a CBOFS snapshot for Sept, 4, 2022.}
+    \end{center}    
+\end{figure}
+
+\FloatBarrier
 

### 2022-09-16 manuscript/manuscript.tex
text edits

@@
 
-We supplemented observational data with a grid of “freshwater influence” (FWI) estimates where each grid cell represents the average of flow weighted distance of that cell from the 5 major tributaries (Susquehanna, James, Pautexent, Choptank, and Potomac).
+We supplemented observational data with a grid of “freshwater influence” (FWI) estimates where each grid cell represents the average of annual-flow weighted distance of that cell from the 5 major tributaries (Susquehanna, James, Pautexent, Choptank, and Potomac).
+Annual flow magnitudes were calculated from National Water Information System instantaneous data (sites: 01488110, 01491000, 01576000, 01654000, 01594440, 02037500, \citet{nwis2021usgs}).
 Rather than straight-line distances we used “in-water” distances calculated using the scikit-image Python package \citep{vanderwaltScikitimageImageProcessing2014}.
-We combined the FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands values for each pixel spanning from 405 to 877nm wavelength \citep{vermoteericMOD09GAMODISTerra2015}.
+We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands values for each pixel spanning from 405 to 877nm wavelength \citep{vermoteericMOD09GAMODISTerra2015}.
 

### 2022-09-16 manuscript/manuscript.tex
text edits

@@
 
-\FloatBarrier
+\section{Discussion}
 
-\section{Discussion}
+Our data-driven yet geographically aware approach was able to accurately reproduce spatial and temporal patterns of Bay salinity and temperature.
+In addition to being accurate, our approach provides the ability to generate rapid bay-wide prediction surfaces in seconds and it is easily updatable insofar as saved models can be reloaded and have their training “updated”.
+Our models are also lightweight with minimal internal complexity and extensively tested against a uniquely large observational dataset.
+
+\subsection{Spatial patterning}
+
+\subsection{Temporal patterning}
+
+\subsection{Future research}
 

### 2022-09-16 manuscript/manuscript.tex
text edits

@@
 
+One shortcoming of our work which may be a fruitful area of future research is the lack of realistic small-scale features evident in the physics based output of the CBOFS (Figure 4).
+These small-scale features such as eddies and other circulation features may be resolved with the use of convolutional neural networks (CNNs).
+A key difference between our approach and that of CNNs is that our approach fits each observation essentially in isolation and does not attempt to reproduce the “neighborhood” surrounding observations as is the case with CNNs (citation).
+
+Another potential area of future research is resolving water properties at depths beyond the surface.
+A key challenge in this area is lack of observational data at multiple locations.
+Depth arrayed measurements are typically only associated with buoy deployments of which there are few in the Bay.
+Perhaps these can be supplemented by data from physics-based models such as CBOFS.
+The results of such as an effort can be regarded more so as an emulator than a standalone model.
+
 \FloatBarrier

### 2022-09-16 scripts/animation.py
animate cbofs data

@@ -5,12 +5,10 @@ import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
 import matplotlib.animation as anim
 
-flist = glob.glob("data/prediction/*.tif")
-
 
 def read_xr(date):
     # date = "2018-01-01"
-    fpath = "data/prediction/" + date + ".tif"
+    fpath = "data/save_cbofs/salt_" + date + ".tif"
     date_pd = pd.date_range(date, date, freq='D', inclusive='left')
     date_xr = xr.DataArray(date_pd, [("time", date_pd)])
 
@@ -19,15 +17,23 @@ def read_xr(date):
     return dt
 
 
+def f_to_date(x):
+    # x = flist[0]
+    return x.split("_")[2].replace(".tif", "")
+    # return x[0:4] + "-" + x[4:6] + "-" + x[6:8]
+
+
+flist = glob.glob("data/save_cbofs/*.tif")
+dates = [f_to_date(f) for f in flist]
+
 #  --- animation
+xr_list = [read_xr(date) for date in dates]
+test = xr.concat(xr_list, dim="time")
 
-test = xr.concat([read_xr("2018-01-01"), read_xr("2018-01-02")], dim="time")
-
-variable = test.band_data.sel(time=slice('2018', '2018'))
+variable = test.band_data.sel(time=slice('2021', '2022'))
 
 fig = plt.figure()
 ax = plt.axes(projection=ccrs.PlateCarree())
-
 image = variable.isel(time=0).squeeze().plot.imshow(
     ax=ax, transform=ccrs.PlateCarree(), animated=True)
 
@@ -42,6 +48,7 @@ def update(t):
 
 animation = anim.FuncAnimation(fig,
                                update,
+                               interval=700,
                                frames=variable.time.values,
                                blit=False)
 # animation.save("test.gif")

### 2022-09-19 figures/maps.py
towards prediction pipeline for all vars

@@ -127,7 +127,7 @@ img_cbofs = img_cbofs["band_data"].sel(band=1)
 # img_cbofs.plot.imshow()
 # plt.show()
 
-img_rf = utils.get_rf_prediction("2022-09-04")
+img_rf = utils.get_rf_prediction("2022-09-04", "salinity")
 
 # gdal_calc.py -a data/prediction/2022-09-04_downsample_clip.tif -b data/cbofs/salt_20220904.tif --calc="a - b" --outfile c.tif
 
@@ -172,7 +172,7 @@ bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
     epsg=4326)
 
 dates = ["2021-12-04", "2022-03-04", "2022-06-04", "2022-09-04"]
-imgs = [utils.get_rf_prediction(date) for date in dates]
+imgs = [utils.get_rf_prediction(date, "salinity") for date in dates]
 
 nrow = 2
 ncol = 2

### 2022-09-19 figures/timeseries.py

@@ -15,7 +15,7 @@ flist = list(
     itertools.compress(flist,
                        [not strng.__contains__("2018") for strng in flist]))
 
-dates = [x.split("\\")[1].replace(".tif", "") for x in flist]
+dates = [x.split("\\")[1].replace(".tif", "").replace("_salinity", "") for x in flist]
 
 pixel_centers_all = gpd.read_file("data/pixel_centers_4326.shp").rename(
     columns={"pix_idx": "pix_id"})

### 2022-09-19 scripts/04_rf_predict.py

@@ -23,27 +23,28 @@ def _get_coords(x):
     return (longitude, latitude)
 
 
-variable = "salinity"
-var_col = "SSS (psu)"
-best_params = utils.load_md(variable)
-
-rf_random_path = "data/rf_random_" + variable + ".pkl"
-rf_random = pickle.load(open(rf_random_path, "rb"))
-
-predictors = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
-# predictors = [
-#     'datetime', 'sur_refl_b08', 'sur_refl_b09', 'sur_refl_b10', 'sur_refl_b11',
-#     'sur_refl_b12', 'sur_refl_b13', 'sur_refl_b14', 'sur_refl_b15',
-#     'sur_refl_b16', 'latitude', 'longitude', "cost"
-# ]
-
-
 def main():
     parser = argparse.ArgumentParser()
     parser.add_argument("--date", type=str)
+    parser.add_argument("--variable", type=str)
+    parser.add_argument("--var_col", type=str)
     args = vars(parser.parse_args())
     date = args["date"]
+    variable = args["variable"]
+    var_col = args["var_col"]
     # date = "2022-09-04"
+    # variable = "salinity"
+    # var_col = "SSS (psu)"
+
+    rf_random_path = "data/rf_random_" + variable + ".pkl"
+    rf_random = pickle.load(open(rf_random_path, "rb"))
+
+    predictors = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
+    # predictors = [
+    #     'datetime', 'sur_refl_b08', 'sur_refl_b09', 'sur_refl_b10', 'sur_refl_b11',
+    #     'sur_refl_b12', 'sur_refl_b13', 'sur_refl_b14', 'sur_refl_b15',
+    #     'sur_refl_b16', 'latitude', 'longitude', "cost"
+    # ]
 
     # --- prep prediction data
     dt = pd.read_csv("data/prediction/modis-" + date.replace("-", "_") +
@@ -109,7 +110,7 @@ def main():
     #     geo_grid.predict != geo_grid.predict.rio.nodata).plot()
     # plt.show()
 
-    out_path = "data/prediction/" + date + ".tif"
+    out_path = "data/prediction/" + date + "_" + variable + ".tif"
     geo_grid["predict"].rio.to_raster(out_path)
 
     return out_path

### 2022-09-19 src/rf_icom_call_data2.py

@@ -109,7 +109,10 @@ def clean_data(variable, var_col, predictors, test_size=0.5, data=None):
     feature_names = [k for k in data.keys() if k in predictors]
 
     plt.figure(figsize=(20, 10))
-    data.hist()
+    try:
+        data.hist()
+    except:        
+        data.drop(columns=["datetime"]).hist()
     plt.tight_layout()
     plt.savefig("figures/" + variable + "_data_hist.png")
 

### 2022-09-19 src/utils.py

@@ -95,22 +95,26 @@ def modisaqua_path(date, band="sur_refl_b08"):
     return "data/MODIS-Aqua/{date}_{band}.tif".format(date=date, band=band)
 
 
-def get_rf_prediction(date):
+def get_rf_prediction(date, variable):
     # date = "2022-09-04"
+    # variable = "salinity"
 
     bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
         epsg=4326)
 
-    path_downsample = "data/prediction/" + date + "_downsample.tif"
+    path_base = "data/prediction/" + date + "_" + variable + ".tif"
+    path_downsample = "data/prediction/" + date + "_" + variable + "_downsample.tif"
+    path_downsample_clip = "data/prediction/" + date + "_" + variable + "_downsample_clip.tif"
+
     if not os.path.exists(path_downsample):
-        call_string = "gdalwarp -te -77.3425000000000011 36.1675000000000040 -74.7974999999999994 39.6325000000000074 -ts 509 693 -overwrite data/prediction/" + date + ".tif data/prediction/" + date + "_downsample.tif"
+        call_string = "gdalwarp -te -77.3425000000000011 36.1675000000000040 -74.7974999999999994 39.6325000000000074 -ts 509 693 -overwrite " + path_base + " " + path_downsample
         subprocess.call(call_string)
 
-    img_rf = xr.open_dataset("data/prediction/" + date + "_downsample.tif",
+    img_rf = xr.open_dataset(path_downsample,
                              engine="rasterio")
     img_rf = img_rf.rio.clip(bay_gdf_hires.geometry)
     img_rf = img_rf["band_data"].sel(band=1)
-    img_rf.rio.to_raster("data/prediction/" + date + "_downsample_clip.tif")
+    img_rf.rio.to_raster(path_downsample_clip)
     # img_rf.plot.imshow()
     # plt.show()
     return img_rf

### 2022-09-19 scripts/04_rf_predict.py
prediction surfaces for temp and turb

@@ -12,7 +12,7 @@ from geocube.api.core import make_geocube
 from geocube.rasterize import rasterize_points_griddata
 
 sys.path.append(".")
-from src import utils
+from src import fit_sine
 from src import rf_icom_call_data2 as call_data2
 
 
@@ -82,7 +82,17 @@ def main():
     # --- predict prediction data
     predictions = rf_random.predict(dt_clean_X)
     res = pd.DataFrame(dt_clean_y, columns=["obs"])
-    res["predict"] = predictions.copy()
+
+    if variable == "temperature":
+        # back out non-sine adjusted values    
+        # pickle.load(open("data/feature_names_temperature.pkl", "rb"))
+        p1 = pickle.load(open("data/temperature_sine_coef.pkl", "rb"))
+        offset = fit_sine.fitfunc(p1, dt_clean_X[0:, 0])
+        res["predict"] = predictions.copy() + offset
+        res["obs"] = res["obs"] + offset
+    else:
+        res["predict"] = predictions.copy()
+
     res["longitude"] = lon_list
     res["latitude"] = lat_list
     res = gpd.GeoDataFrame(data=res,

### 2022-09-19 src/rf_icom_call_data2.py

@@ -45,10 +45,13 @@ def clean_data(variable, var_col, predictors, test_size=0.5, data=None):
 
     data = utils.select_var(data, var_col)
 
-    ## For turbidity get rid of negative numbers
     print(len(data))
-    if var_col == "turbidity (NTU)":
-        data = data[data["turbidity (NTU)"] > 0]
+    
+    ## For turbidity get rid of negative numbers
+    if test_size > 0: # aka we are not in prediction mode    
+        if var_col == "turbidity (NTU)":
+            # any(data["turbidity (NTU)"] > 0)
+            data = data[data["turbidity (NTU)"] > 0]
 
     ## For salinity, set negative numbers to 0
     if var_col == "SSS (psu)":        

### 2022-09-19 figures/maps.py
towards seasonality map pipeline

@@ -21,6 +21,8 @@ def panel_add(i,
               diff=False,
               j=None,
               vmax=27,
+              vmin=0,
+              ticks=None,
               height_frac=0.5):
     if j is not None:
         ax = plt.subplot(axs[i, j], xlabel="", projection=ccrs.PlateCarree())
@@ -37,11 +39,12 @@ def panel_add(i,
                              })
     else:
         geo_grid.plot.imshow(ax=ax,
-                             vmin=0,
+                             vmin=vmin,
                              vmax=vmax,
                              cbar_kwargs={
                                  "shrink": 0.5,
-                                 'label': ''
+                                 'label': '',
+                                 'ticks': ticks,
                              })  # , vmax=np.nanmax(img_cbofs.to_numpy()))
     ax.set_title(title,
                  size="small",
@@ -168,34 +171,42 @@ panel_add(2, axs, "RF-CBOFS", test, diff=True, height_frac=0.7)
 plt.savefig("figures/_rf-vs-cbofs.pdf")
 
 # --- seasonality maps
-bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
-    epsg=4326)
+def _seasonality_map(variable, vmin=(0,0,0,0), vmax=(27,27,27,27), ticks=None):
+    # variable  = "turbidity"
+    # vmin=(3,3,21,21)
+    # vmax=(12, 12, 29, 29)
+    dates = ["2021-12-04", "2022-03-04", "2022-06-04", "2022-09-04"]
+    imgs = [utils.get_rf_prediction(date, variable) for date in dates]
 
-dates = ["2021-12-04", "2022-03-04", "2022-06-04", "2022-09-04"]
-imgs = [utils.get_rf_prediction(date, "salinity") for date in dates]
+    nrow = 2
+    ncol = 2
 
-nrow = 2
-ncol = 2
+    fig = plt.figure(figsize=(ncol + 3, nrow + 3))
+    axs = gridspec.GridSpec(nrow,
+                            ncol,
+                            wspace=0.0,
+                            hspace=0.0,
+                            top=1. - 0.5 / (nrow + 1),
+                            bottom=0.5 / (nrow + 1),
[…]

### 2022-09-19 src/utils.py

@@ -96,8 +96,8 @@ def modisaqua_path(date, band="sur_refl_b08"):
 
 
 def get_rf_prediction(date, variable):
-    # date = "2022-09-04"
-    # variable = "salinity"
+    # date = "2021-12-04"
+    # variable = "temperature"
 
     bay_gdf_hires = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
         epsg=4326)

### 2022-09-19 figures/maps.py
placeholder for overall panel title

@@ -203,6 +203,7 @@ def _seasonality_map(variable, vmin=(0,0,0,0), vmax=(27,27,27,27), ticks=None):
     # fig.subplots_adjust(wspace=0, hspace=0)
     # plt.show()
     out_path = "figures/_seasonality_" + variable + ".pdf"
+    # fig.suptitle(variable, y=0.87, x=0.42)
     plt.savefig(out_path)
 
     return out_path

### 2022-09-19 manuscript/manuscript.tex
typo
multi-multi-panel prediction maps

@@
 
-Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure S1).
+Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 1).
 Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1).
 Despite excellent overall performance, our temperature and salinity models had poor performance at the extremes.
-They were unable to recover very low values ($<$ 0 C) and very high values ($>$ 30 C, $>$ 23 salinity, Figure 1).
-These areas of poor performance can be explained by their relative rarity in the observational dataset (Table S2).
+They were unable to recover very low values ($<$ 0 C) or very high values ($>$ 30 C, $>$ 23 salinity, Figure 1).
+One explanation for these areas of poor performance are their relative rarity in the observational dataset (Table S2).
 
@@
 Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 3).
-RF model predictions were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 4).
-The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 2).
+A realistic spatial trend can also be seen in the RF temperature predictions whereby maximum water temperatures were located in the upper tributaries while minimum water temperatures were highest at the Bay mouth (Figure 3).
+
+RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 4).
+The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 3).
 
@@
     \begin{center}
-          \includegraphics[width=0.4\textwidth,keepaspectratio]{figures/_seasonality}
-          \caption{Seasonal Random Forest salinity prediction results.}
+          \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_seasonality}
+          \caption{Seasonal Random Forest prediction results for temperature and salinity.}
     \end{center}    
@@
 Perhaps these can be supplemented by data from physics-based models such as CBOFS.
-The results of such as an effort can be regarded more so as an emulator than a standalone model.
+The results of such an effort can be regarded more so as an emulator than a standalone model.
 

### 2022-09-19 figures/maps.py

@@ -203,7 +203,7 @@ def _seasonality_map(variable, vmin=(0,0,0,0), vmax=(27,27,27,27), ticks=None):
     # fig.subplots_adjust(wspace=0, hspace=0)
     # plt.show()
     out_path = "figures/_seasonality_" + variable + ".pdf"
-    # fig.suptitle(variable, y=0.87, x=0.42)
+    fig.suptitle(variable, y=0.87, x=0.42)
     plt.savefig(out_path)
 
     return out_path

### 2022-09-19 manuscript/manuscript.tex
text edits

@@
 Annual flow magnitudes were calculated from National Water Information System instantaneous data (sites: 01488110, 01491000, 01576000, 01654000, 01594440, 02037500, \citet{nwis2021usgs}).
-Rather than straight-line distances we used “in-water” distances calculated using the scikit-image Python package \citep{vanderwaltScikitimageImageProcessing2014}.
+Rather than straight-line distances we used “in-water” distances calculated using the scikit-image Python package \citep{littleKrigingEstuariesCrow1997,vanderwaltScikitimageImageProcessing2014}.
 We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands values for each pixel spanning from 405 to 877nm wavelength \citep{vermoteericMOD09GAMODISTerra2015}.
@@
 
-Tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 1).
-Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1).
+Our tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 1).
+Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1, Figure S1).
 Despite excellent overall performance, our temperature and salinity models had poor performance at the extremes.
-They were unable to recover very low values ($<$ 0 C) or very high values ($>$ 30 C, $>$ 23 salinity, Figure 1).
+They were unable to recover very low temperature values ($<$ 0 C) or very high temperature and salinity values ($>$ 30 C, $>$ 23 salinity, Figure 1).
 One explanation for these areas of poor performance are their relative rarity in the observational dataset (Table S2).
+The salinity model however did faithfully capture areas of low salinity (Figure 1).
+These low salinity values can be regarded as low uncertainty (high confidence) estimates given the extent to which the upper tributaries are well represented in the observational dataset (Figure S2).
 
@@
 RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 4).
+CBOFS surface water salinity estimates are known to be saltier than observations \citep{lanerolle2011second, vogelAssessingSatelliteSea2016} partly explaining our saltier data-driven results.
 The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 3).
@@
     \begin{center}
-          \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_seasonality}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_seasonality}
           \caption{Seasonal Random Forest prediction results for temperature and salinity.}
@@
 
-Our data-driven yet geographically aware approach was able to accurately reproduce spatial and temporal patterns of Bay salinity and temperature.
+Our geographically aware approach was able to accurately reproduce spatial and temporal patterns of Bay salinity and temperature.
+Model errors were generally lower than prior data-driven efforts and as good or better than physics-based models \citep{vogelAssessingSatelliteSea2016}.
 In addition to being accurate, our approach provides the ability to generate rapid bay-wide prediction surfaces in seconds and it is easily updatable insofar as saved models can be reloaded and have their training “updated”.
-Our models are also lightweight with minimal internal complexity and extensively tested against a uniquely large observational dataset.
+Our models are lightweight with minimal internal complexity and have been extensively tested against a uniquely large observational dataset.
 
@@
 
+Temperature and salinity models produced realistic predictions insofar as they were characterized by low salinity in the upper tributaries and high minimum water temperatures at the Bay mouth.
+This indicates that the model captures both freshwater discharge from tributaries as well as the moderating influence of ocean water in areas near the Bay mouth \citep{dingSpatiotemporalPatternsWater2015}.
+
+Often it is difficult to assess the inherent uncertainty of data-driven efforts such as ours.
+However, in our case, we assembled such a large observational dataset, that it makes such assessments possible.
+This is a key difference between our approach and prior efforts which have mostly been trained on high salinity buoy data or in-situ cruises from the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013}.
+Predicted areas of low salinity in particular likely have low associated uncertainties because the upper tributaries were represented with high density in our observational dataset (Figure S2).
+Because CBOFS salinity data was fresher than our RF estimates in the upper tributaries, this suggests that CBOFS data for these areas are true underestimates and not a deficiency of the RF model.
+
 \subsection{Temporal patterning}
+
+The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons.
+However, because the salinity model was not able to reproduce the highest salinity values Fall and Winter estimates may be biased low.
+For the temperature model, we were unable to recover low temperature observations less than 0 deg C.
+These results are consistent with prior efforts such as \citet{geigerSatellitederivedCoastalOcean2013} where winter had the highest seasonal error.
+One explanation for these high winter error rates is the fact that winter in the Bay is generally characterized by strong vertical mixing due to high winds \citep{sonWaterPropertiesChesapeake2012, xuClimateForcingSalinity2012}.
 
@@
 Depth arrayed measurements are typically only associated with buoy deployments of which there are few in the Bay.
-Perhaps these can be supplemented by data from physics-based models such as CBOFS.
+Perhaps these can be supplemented by data from physics-based models such as CBOFS which are fully three-dimensional.
 The results of such an effort can be regarded more so as an emulator than a standalone model.
 
+\section{Conclusion}
+
+We have demonstrated an approach to produce realistic estimates of water properties in Chesapeake Bay using remotely sensed data.
+Much of this realism comes from our FWI routines which enforce physical consistency in upstream-downstream spatial patterning.
+Estimates produced by our model have a high potential to inform calibration and evaluation of Earth Systems Models as well as in other applications such as habitat suitability and sea level rise models.
+
 \FloatBarrier
+
+\section{Acknowledgements}
+
+J.S.
+built models, analyzed data, and wrote the paper.
+S.A.
+designed the initial model architecture and variable importance routines.
+J.Sc.
+conceived the study, collected observational data, and edited the paper.
+This work was supported by the Integrated Coastal Modeling (ICoM) project through the Department of Energy, Office of Science,
+Biological and Environmental Research, Earth and Environment Systems Sciences Division.
+
 \bibliographystyle{apalike}

### 2022-09-20 manuscript/manuscript.tex
text edits

@@
 Although this data came from a wide variety of sources (n=20, Table S1), almost $80\%$ of the total unique observations came from three programs NDBC/CBIBS (\url{https://buoybay.noaa.gov/}), Maryland Continuous Monitoring Program (MDNR), and the Virginia Estuarine and Coastal Observing System (\url{http://vecos.vims.edu}).
+The temporal range of the observations includes both times of minimum and maximum freshwater input to the Bay (\url{https://www.usgs.gov/media/images/estimated-annual-mean-streamflow-entering-chesapeake-bay}).
+
 Because Chesapeake Bay has a mean optical depth of 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column.
@@
 Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1, Figure S1).
-Despite excellent overall performance, our temperature and salinity models had poor performance at the extremes.
+Despite excellent overall performance for the temperature and salinity models, they had poor performance at the extremes.
 They were unable to recover very low temperature values ($<$ 0 C) or very high temperature and salinity values ($>$ 30 C, $>$ 23 salinity, Figure 1).
@@
     \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_seasonality}
+          \includegraphics[width=0.7\textwidth,keepaspectratio]{figures/_seasonality}
           \caption{Seasonal Random Forest prediction results for temperature and salinity.}
@@
 
-Our geographically aware approach was able to accurately reproduce spatial and temporal patterns of Bay salinity and temperature.
+Our geographically aware approach was able to accurately reproduce realistic spatial and temporal patterns of Bay salinity and temperature.
 Model errors were generally lower than prior data-driven efforts and as good or better than physics-based models \citep{vogelAssessingSatelliteSea2016}.
@@
 Our models are lightweight with minimal internal complexity and have been extensively tested against a uniquely large observational dataset.
+
+One reason why the turbidity model may have had poor performance is because our observational data only contains data in Nephelometric Turbidity Units (NTUs) which are not standardized units.
+As a result, they cannot be reliably compared across different instruments.
+A better turbidity model would likely result from using total suspended matter (TSM) data measured in a standardized unit such as mg/l \citep{ondrusekDevelopmentNewOptical2012}.
 
@@
 However, in our case, we assembled such a large observational dataset, that it makes such assessments possible.
-This is a key difference between our approach and prior efforts which have mostly been trained on high salinity buoy data or in-situ cruises from the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013}.
+This is a key difference between our approach and prior efforts which have mostly been trained on high salinity buoy data or in-situ cruises from the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013, ondrusekDevelopmentNewOptical2012}.
 Predicted areas of low salinity in particular likely have low associated uncertainties because the upper tributaries were represented with high density in our observational dataset (Figure S2).
 Because CBOFS salinity data was fresher than our RF estimates in the upper tributaries, this suggests that CBOFS data for these areas are true underestimates and not a deficiency of the RF model.
+
+In contrast to the upper tributaries, an area of maximum uncertainty is likely the mainstem and eastern Bay across from the mouth of the Potomac (Tangier Sound).
+This appears to be the upper limit of the marine plume emanating from the Bay mouth.
+The hydrodynamics here seem to be complex given that the CBOFS has difficulty resolving spatial patterning in this area.
+Improving our data-driven modeling approach would be difficult here because this area has few observational data (Figure S2).
 

### 2022-09-20 manuscript/manuscript.tex
text edits

@@
 \begin{abstract}
-    \noindent Remotely sensed water properties are important for a variety of applications, including validation of Earth Systems models (ESMs).
-However, the usefulness of operational forecasting products and directly-observing satellite-based sensors for validation of next-generation ESMs is limited due to their temporal availability and spatial resolution of $<$ 1 year and $>$ 30 $km^2$ respectively.
+    \noindent Remotely sensed water properties are important for a variety of applications, including validation of Earth Systems models (ESMs), habitat suitability models, and sea level rise projections.
+For the validation of next-generation ESMs in particular, the usefulness of operational forecasting products and directly-observing satellite-based sensors for validation of next-generation ESMs is limited due to their temporal availability and spatial resolution of $<$ 1 year and $>$ 30 $km^2$ respectively.
 To address this validation data gap, we developed a data-driven model to produce high-resolution ($<$ 1 $km^2$) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.
 Our model fits daily MODIS Aqua reflectance data to surface observations ($<$ 1 m depth) from 2000-2021 in Chesapeake Bay, USA.
-Our model has similar error statistics as prior efforts of this type for salinity (RMSE: 2.2) and temperature (RMSE: 1.9 C).
+The resulting models have similar error statistics as prior efforts of this type for salinity (RMSE: 2.2) and temperature (RMSE: 1.9 C).
 However, unlike prior efforts our model is set up as a pipeline meaning that it has the advantage of producing predictions of water properties in future time periods as additional MODIS data is collected.
@@
 
-To address this validation data gap, we developed a data-driven model to produce high-resolution (< 1 $km^2$) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.
+To address this validation data gap, we developed a data-driven model to produce high-resolution (< 1 $km^2$) estimates of temperature, salinity, and turbidity at daily time-steps (over decadal time scales) as required by next-generation ESMs.
 The primary input to our model is remotely sensed reflectance data.
@@
 Our model represents a further advantage over prior efforts at data-driven modeling of salinity in the Bay (e.g.
-\citet{vogelAssessingSatelliteSea2016}) because it is calibrated against not only observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
+\citet{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
 
@@
 Chesapeake Bay is located in the Mid-Atlantic region of the United States and serves as the outlet of 9 major tributary systems.
-Of these, the Susquehanna River provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \citet{xuClimateForcingSalinity2012}).
+Of these, the Susquehanna River (at the Northern extent of the Bay) provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \citet{xuClimateForcingSalinity2012}).
 The Bay is fairly shallow ($<$ 30m, \citet{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \citep{xuClimateForcingSalinity2012}.
@@
 Prior data-driven models (e.g.
-\citet{urquhartGeospatialInterpolationMODISderived2013, vogelAssessingSatelliteSea2016}) of Bay salinity used Generalized Linear Models (GLM), Neural Networks (NN), Generalized Additive Models (GAMs), and Random Forests (RF).
-We chose to use the Random Forest algorithm in part because \citet{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) but also because of its ease of implementation and interpretation.
+\citet{urquhartGeospatialInterpolationMODISderived2013, vogelAssessingSatelliteSea2016}) of Bay water properties used Generalized Linear Models (GLM), Neural Networks (NN), Generalized Additive Models (GAMs), and Random Forests (RF).
+We chose to use the Random Forest algorithm in part because \citet{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) and also because of its ease of implementation and interpretation.
 We fit RF models using the scikit-learn Python package \citep{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model.
-We subsequently “tuned” each model using k-fold cross validation to optimize the size and splitting criteria of each tree.
-We evaluated the “importance” of a particular variable as the average impurity decrease of tree splits based on that variable \citep{pedregosaScikitlearnMachineLearning2011}.
+We subsequently “tuned” each model using k-fold cross validation to optimize the size, depth, and splitting criteria of each tree.
+We evaluated the “importance” of particular variables as the average impurity decrease of tree splits based on that variable \citep{pedregosaScikitlearnMachineLearning2011}.
 
@@
 
+\begin{sloppypar}
+
 We compiled over two decades of observational data (2000-2021) on salinity (n=6k), temperature (n=10k), and turbidity (n=5k, Table S1).
 Although this data came from a wide variety of sources (n=20, Table S1), almost $80\%$ of the total unique observations came from three programs NDBC/CBIBS (\url{https://buoybay.noaa.gov/}), Maryland Continuous Monitoring Program (MDNR), and the Virginia Estuarine and Coastal Observing System (\url{http://vecos.vims.edu}).
-The temporal range of the observations includes both times of minimum and maximum freshwater input to the Bay (\url{https://www.usgs.gov/media/images/estimated-annual-mean-streamflow-entering-chesapeake-bay}).
+The temporal range of the observations (2000-2021) includes the full range of freshwater input magnitudes to the Bay spanning both times of minimum and maximum freshwater input (\url{https://www.usgs.gov/media/images/estimated-annual-mean-streamflow-entering-chesapeake-bay}).
 
-Because Chesapeake Bay has a mean optical depth of 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column.
+\end{sloppypar}
+
+Because Chesapeake Bay has a mean optical depth of approximately 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column.
 We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours).
 Prior to model fitting, we removed the annual cycle from temperature observations using least squares optimization routines \citep{virtanenSciPyFundamentalAlgorithms2020a}.
-Finally, we implemented spatial and temporal averaging of observational data to a daily mean within each MODIS pixel.
+This annual cycle was added back to model estimates for evaluation purposes and in the production of prediction surfaces.
+As a final observational data processing step, we implemented spatial and temporal averaging of observational data to a daily mean within each MODIS pixel.
 
@@
 Our tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 1).
-Out of sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1, Figure S1).
-Despite excellent overall performance for the temperature and salinity models, they had poor performance at the extremes.
+Out-of-sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1, Figure S1).
+Despite excellent overall performance for the temperature and salinity models, there was some indication of poor performance at the extremes.
 They were unable to recover very low temperature values ($<$ 0 C) or very high temperature and salinity values ($>$ 30 C, $>$ 23 salinity, Figure 1).
-One explanation for these areas of poor performance are their relative rarity in the observational dataset (Table S2).
+One explanation for these areas of poor performance is that such extremes were rare in the observational dataset (Table S2).
 The salinity model however did faithfully capture areas of low salinity (Figure 1).
@@
 
-The date of observation was the most important feature in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 2).
-Across all models, the most important MODIS band was band 8.
+The date of observation was the most important feature identified in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 2).
+Across all models, the most important MODIS band was band 8, which is the shortest wavelength band of the MODIS Aqua product at the extreme of the visible light range.
 The “size” (i.e.
@@
 
-One reason why the turbidity model may have had poor performance is because our observational data only contains data in Nephelometric Turbidity Units (NTUs) which are not standardized units.
+Although salinity and temperature models were accurate, our turbidity model showed relatively poor performance.
+One reason for this may have been because our observational data only contains data in Nephelometric Turbidity Units (NTUs) which are not standardized units.
 As a result, they cannot be reliably compared across different instruments.
@@
 
-Often it is difficult to assess the inherent uncertainty of data-driven efforts such as ours.
-However, in our case, we assembled such a large observational dataset, that it makes such assessments possible.
-This is a key difference between our approach and prior efforts which have mostly been trained on high salinity buoy data or in-situ cruises from the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013, ondrusekDevelopmentNewOptical2012}.
-Predicted areas of low salinity in particular likely have low associated uncertainties because the upper tributaries were represented with high density in our observational dataset (Figure S2).
-Because CBOFS salinity data was fresher than our RF estimates in the upper tributaries, this suggests that CBOFS data for these areas are true underestimates and not a deficiency of the RF model.
+Often it is difficult to assess the inherent spatial uncertainty of data-driven efforts such as ours.
+However, we assembled such a large observational dataset, that it makes such assessments possible in the present study.
+This is a key difference between our approach and prior efforts which have mostly been trained on limited (and relatively high-salinity) buoy data or on data from in-situ cruises located in the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013, ondrusekDevelopmentNewOptical2012}.
+Predicted areas of low salinity coming from our model likely have particularly low associated uncertainties because the upper tributaries were well-represented in our observational dataset (Figure S2).
+Because CBOFS salinity data was fresher than our RF estimates in the upper tributaries, this suggests that CBOFS data for these areas are likely true underestimates and not a deficiency of the RF model predictions.
 
 In contrast to the upper tributaries, an area of maximum uncertainty is likely the mainstem and eastern Bay across from the mouth of the Potomac (Tangier Sound).
-This appears to be the upper limit of the marine plume emanating from the Bay mouth.
+This appears to be the collision point of the marine plume emanating from the Bay mouth with the discharge of the Potomac River.
 The hydrodynamics here seem to be complex given that the CBOFS has difficulty resolving spatial patterning in this area.
-Improving our data-driven modeling approach would be difficult here because this area has few observational data (Figure S2).
+This was an area where we found limited observational data (Figure S2).
+Improvements to data-driven modeling in this area may be possible with the inclusion of more observational data.
 
@@
 These results are consistent with prior efforts such as \citet{geigerSatellitederivedCoastalOcean2013} where winter had the highest seasonal error.
-One explanation for these high winter error rates is the fact that winter in the Bay is generally characterized by strong vertical mixing due to high winds \citep{sonWaterPropertiesChesapeake2012, xuClimateForcingSalinity2012}.
+One explanation for these high winter error rates is the fact that winter in the Bay is generally characterized by strong vertical mixing due to high winds \citep{sonWaterPropertiesChesapeake2012, xuClimateForcingSalinity2012} making “true” patterns more dynamic and more difficult to resolve.
 
@@
 One shortcoming of our work which may be a fruitful area of future research is the lack of realistic small-scale features evident in the physics based output of the CBOFS (Figure 4).
-These small-scale features such as eddies and other circulation features may be resolved with the use of convolutional neural networks (CNNs).
-A key difference between our approach and that of CNNs is that our approach fits each observation essentially in isolation and does not attempt to reproduce the “neighborhood” surrounding observations as is the case with CNNs (citation).
+It may be possible to resolve small-scale features such as eddies and other circulation features with the use of convolutional neural networks (CNNs).
+A key difference between our approach and that of CNNs is that our approach essentially fits each observation in isolation and does not exploit the “spatial locality” (i.e.
+the “neighborhood”) surrounding observations as is the case with CNNs \citep{goodfellow2016convolutional}.
 
@@
 A key challenge in this area is lack of observational data at multiple locations.
-Depth arrayed measurements are typically only associated with buoy deployments of which there are few in the Bay.
-Perhaps these can be supplemented by data from physics-based models such as CBOFS which are fully three-dimensional.
+Depth arrayed measurements are typically only associated with buoy deployments of which there are few in the Bay \cite{vogelAssessingSatelliteSea2016}.
+It may be possible to supplement these buoy observations with data from physics-based models such as CBOFS which are fully three-dimensional.
 The results of such an effort can be regarded more so as an emulator than a standalone model.

### 2022-09-23 manuscript/manuscript.tex
use official acknowledgements text

@@
 conceived the study, collected observational data, and edited the paper.
-This work was supported by the Integrated Coastal Modeling (ICoM) project through the Department of Energy, Office of Science,
-Biological and Environmental Research, Earth and Environment Systems Sciences Division.
+This work was supported by the Earth System Model Development and Regional and Global Modeling and Analysis program areas of the U.S.
+Department of Energy, Office of Science, Office of Biological and Environmental Research as part of the multi-program, collaborative Integrated Coastal Modeling (ICoM) project.
 

### 2022-09-26 manuscript/manuscript.tex
text edits

@@
 The primary input to our model is remotely sensed reflectance data.
-In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \citep{wangDevelopmentMODISData2018}.
+In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in the Bay as well as in coastal Louisiana \citep{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}.
 We supplemented MODIS data with gridded estimates of tributary inflows which we used to enforce physical realism (i.e.
@@
 Because Chesapeake Bay has a mean optical depth of approximately 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column.
-We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours).
+We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours GMT).
 Prior to model fitting, we removed the annual cycle from temperature observations using least squares optimization routines \citep{virtanenSciPyFundamentalAlgorithms2020a}.

### 2022-09-26 figures/validation.py
reorder to match importance plot

@@ -32,8 +32,8 @@ def _get_predictions(variable):
     return res
 
 
-variables = ["SSS (psu)", "turbidity (NTU)", "SST (C)"]
-variables_str = ["salinity (psu)", "turbidity (NTU)", "temperature (C)"]
+variables = ["SSS (psu)", "SST (C)", "turbidity (NTU)"]
+variables_str = ["salinity (psu)", "temperature (C)", "turbidity (NTU)"]
 variables_str_short = [utils.clean_var_name(v) for v in variables_str]
 
 # scatter plot of measured vs predicted
@@ -107,16 +107,16 @@ plt.close()
 r2 = []
 fig, axes = plt.subplots(figsize=(8, 4.5), ncols=3)
 r2.append(_plot(axes[0], res[0], "salinity", -0.5, 32, ylab=True, text_space=2.5)[1])
-r2.append(_plot(axes[1],
-      res[1],
+r2.append(_plot(axes[2],
+      res[2],
       "turbidity",
       xy_max=75,
       xy_min=0,
       hatching=False,
       text_anchor=(37, 10),
       text_space=5)[1])
-r2.append(_plot(axes[2],
-      res[2],
+r2.append(_plot(axes[1],
+      res[1],
       "temperature",
       xy_min=-2,
       xy_max=35,

### 2022-09-27 figures/maps.py
better variable ordering

@@ -76,7 +76,7 @@ dt_raw = pd.read_csv(
     os.environ["ICOM_DATA"] +
     "/Modeling Data/Processed Data p1/aggregated_w_bandvals.csv")
 id_vars = ["loc_id", "latitude", "longitude"]
-variables = ["SSS (psu)", "turbidity (NTU)", "SST (C)"]
+variables = ["SSS (psu)", "SST (C)", "turbidity (NTU)"]
 dt_filtered = dt_raw[id_vars + variables]
 dt_melt = pd.melt(dt_filtered, id_vars=id_vars, value_vars=variables)
 

### 2022-09-27 figures/obs_stats.py

@@ -8,7 +8,7 @@ import matplotlib.pyplot as plt
 sys.path.append(".")
 from src import utils
 
-variables = ["SSS (psu)", "turbidity (NTU)", "SST (C)"]
+variables = ["SSS (psu)", "SST (C)", "turbidity (NTU)"]
 
 # --- raw data histograms by variable
 dt_raw = pd.read_csv(

### 2022-09-27 figures/cbibs.py
start cbibs map

@@ -1,0 +1,52 @@
+import os
+import json
+import requests
+import itertools
+import pandas as pd
+import geopandas as gpd
+import cartopy.crs as ccrs
+import matplotlib.pyplot as plt
+
+outfile = "data/cbibs.csv"
+
+## CBIBS
+if not os.path.exists(outfile):
+    base_url = 'https://mw.buoybay.noaa.gov/api/v1'
+    apikey = 'f159959c117f473477edbdf3245cc2a4831ac61f'
+    start = '2000-09-08T01:00:00z'
+    end = '2021-12-09T23:59:59z'
+    var = 'Position'
+
+    query_url = '{}/json/query?key={}&sd={}&ed={}&var={}'.format(base_url,apikey,start,end,var)
+
+    pull = json.loads(requests.get(query_url).text)
+
+    def _extract(dt):
+        # dt = pull["stations"][0]
+        if dt["stationLongName"].__contains__("CBIBS") and dt["active"]:
+            coords = {k: v for k, v in dt.items() if k.endswith('tude')}
+            name = {k: v for k, v in dt.items() if k.endswith('ShortName')}
+            res = name | coords
+        else:
+            res = None
+        return res
+
+    stations = [_extract(dt) for dt in pull["stations"]]
+    is_cbibs = [stations[i] is not None for i in range(0, len(stations))]
+    stations = list(itertools.compress(stations, is_cbibs))
+
+    stations = pd.DataFrame(stations)
+    stations.to_csv(outfile, index=False)
+
+stations = pd.read_csv(outfile)
+stations = gpd.GeoDataFrame(stations,
+                               geometry=gpd.points_from_xy(
+                                   stations["longitude"], stations["latitude"]))
+
+
+ax = plt.axes(projection=ccrs.PlateCarree())
+ax.coastlines(resolution="10m", color="black", linewidth=1)
+stations.plot(ax=ax)
+
+
+plt.show()

### 2022-09-27 figures/cbibs.py
cleanup cbibs map

@@ -43,10 +43,14 @@ stations = gpd.GeoDataFrame(stations,
                                geometry=gpd.points_from_xy(
                                    stations["longitude"], stations["latitude"]))
 
+b = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(
+    epsg=4326)
+b_bounds = [x for x in b.bounds.iloc[0]]
+extent = (b_bounds[0], b_bounds[2], b_bounds[1], b_bounds[3])
 
 ax = plt.axes(projection=ccrs.PlateCarree())
+ax.set_extent(extent, ccrs.PlateCarree())
 ax.coastlines(resolution="10m", color="black", linewidth=1)
 stations.plot(ax=ax)
-
-
-plt.show()
+stations.apply(lambda x: ax.annotate(text=x['stationShortName'], xy=x.geometry.coords[0], xytext=(3, 3), textcoords="offset points"), axis=1)
+plt.savefig("figures/cbibs.pdf")

### 2022-09-27 src/fit_sine.py
formatting with yapf

@@ -8,9 +8,8 @@ import matplotlib.pyplot as plt
 
 fitfunc = lambda p, x: p[0] * np.sin(2 * np.pi / p[1] * x + p[2]) + p[3]
 
+
 def fit_sine(data):
-    # data = pd.read_csv('temperature.csv')
-    # data.replace([np.inf, -np.inf], np.nan, inplace=True)
     # data = pd.read_csv("data/aggregated_w_bandvals.csv")
 
     data['datetime'] = pd.to_datetime(data['datetime'])
@@ -24,17 +23,17 @@ def fit_sine(data):
         temp_mean = np.mean(data['SST (C)'][data['datetime'] == i])
         temp_doy.append(temp_mean)
     date_doy = np.arange(1, 367)
-    
+
     errfunc = lambda p, x, y: fitfunc(p, x
                                       ) - y  # Distance to the target function
     p0 = [12.23414294, 368.06372131, 29.41529669, 16.38386361]
-    p1, success = optimize.leastsq(errfunc, p0[:], args=(date_doy, temp_doy))    
+    p1, success = optimize.leastsq(errfunc, p0[:], args=(date_doy, temp_doy))
     # print(success)
-    # print(p1)    
+    # print(p1)
     pickle.dump(p1, open("data/temperature_sine_coef.pkl", "wb"))
 
     fitted_sine = fitfunc(p1, date)
-    
+
     plt.figure(figsize=(20, 10))
     plt.plot(date_doy, temp_doy, "ro")
     plt.plot(date_doy, fitfunc(p1, date_doy), "bo")

### 2022-09-27 src/fit_sine.py
towards cleaned up temperature sine plot

@@ -34,9 +34,17 @@ def fit_sine(data):
 
     fitted_sine = fitfunc(p1, date)
 
-    plt.figure(figsize=(20, 10))
-    plt.plot(date_doy, temp_doy, "ro")
-    plt.plot(date_doy, fitfunc(p1, date_doy), "bo")
+    plt.rc('axes', labelsize=12)
+    ax = plt.subplot(111)
+    ax.plot(date_doy, temp_doy, "ro")
+    ax.plot(date_doy, fitfunc(p1, date_doy), "bo")    
+    plt.xlabel("DOY")    
+    ax.spines.right.set_visible(False)
+    ax.spines.left.set_visible(False)
+    ax.spines.top.set_visible(False)
+    ax.get_yaxis().set_ticks([])
+    ax.xaxis.set_ticks_position('bottom')
+    # plt.show()
     plt.savefig('figures/fitted_sine.png')
 
     return fitted_sine

### 2022-09-28 src/fit_sine.py
more formatting

@@ -34,17 +34,22 @@ def fit_sine(data):
 
     fitted_sine = fitfunc(p1, date)
 
-    plt.rc('axes', labelsize=12)
-    ax = plt.subplot(111)
+    plt.rc('axes', labelsize=14)
+
+    fig = plt.figure(figsize=[3.4/1.4, 2.8/1.4])
+    ax = fig.add_subplot(111)
     ax.plot(date_doy, temp_doy, "ro")
-    ax.plot(date_doy, fitfunc(p1, date_doy), "bo")    
-    plt.xlabel("DOY")    
+    ax.plot(date_doy, fitfunc(p1, date_doy), "bo", markersize=2)    
+    plt.xlabel("DOY", fontweight="bold")    
     ax.spines.right.set_visible(False)
     ax.spines.left.set_visible(False)
     ax.spines.top.set_visible(False)
     ax.get_yaxis().set_ticks([])
     ax.xaxis.set_ticks_position('bottom')
     # plt.show()
+    plt.tight_layout()
+    labels = ax.get_xticklabels()
+    [label.set_fontweight('bold') for label in labels]
     plt.savefig('figures/fitted_sine.png')
 
     return fitted_sine

### 2022-09-28 scripts/00_calculate_waterdistance.py
move fwi functions to src file

@@ -1,6 +1,7 @@
 # https://gis.stackexchange.com/a/78699/32531
 
 import os
+import sys
 import glob
 import random
 import pickle
@@ -12,25 +13,9 @@ from tqdm import tqdm
 import geopandas as gpd
 from joblib import Parallel, delayed
 from geocube.api.core import make_geocube
-from skimage.graph import route_through_array
 
-
-def get_idx_coords(lon, lat):
-    # lon = stations.iloc[[0]]["longitude"][0]
-    # lat = stations.iloc[[0]]["latitude"][0]
-    dt_where = dt.sel(x=lon, y=lat, method="nearest")
-    dt_where = dt.where(
-        (dt.x == float(dt_where["x"])) & (dt.y == float(dt_where["y"])), drop=True
-    ).squeeze()
-    return (int(dt_where["x_idx"]), int(dt_where["y_idx"]))
-
-
-def get_geo_coords(x, y):
-    # y = 1199
-    # x = 310
-    dt_where = dt.where((dt.x_idx == x) & (dt.y_idx == y), drop=True)
-    return (float(dt_where["x"]), float(dt_where["y"]))
-
+sys.path.append("src")
+from src import fwi
 
 def get_idx_water():
     """
@@ -43,41 +28,6 @@ def get_idx_water():
     return dt_where_water
 
 
-def get_distance(stop_idx_y, stop_idx_x, i):
-    try:
-        indices, weight = route_through_array(
-            cost_surface_array,
-            (start_idx_y, start_idx_x),
-            (stop_idx_x, stop_idx_y),
-            geometric=True,
-            fully_connected=True,
-        )
-        indices = np.array(indices).T
-        weight = int(round(weight, 0))
-
-        path = np.zeros_like(cost_surface_array)
-        path[np.array([indices[0][-1]]), np.array([indices[1][-1]])] = weight
-        # path[indices[0], indices[1]] = weight
-
-        res = xr.DataArray(path, coords=[dt.y.values, dt.x.values], dims=["y", "x"])
-        res["spatial_ref"] = dt["spatial_ref"]
-        res = res.where(res > 0)
-
-        df = res.to_dataframe("cost").reset_index()
-        df = df[~pd.isna(df["cost"])]
-        gdf = gpd.GeoDataFrame(
[…]

### 2022-09-28 src/fwi.py

@@ -1,0 +1,67 @@
+import numpy as np
+import xarray as xr
+import pandas as pd
+import geopandas as gpd
+import matplotlib.pyplot as plt
+import skimage
+from skimage.graph import route_through_array
+
+
+def get_idx_coords(lon, lat, dt):
+    # lon = stations.iloc[[0]]["longitude"][0]
+    # lat = stations.iloc[[0]]["latitude"][0]
+    dt_where = dt.sel(x=lon, y=lat, method="nearest")
+    dt_where = dt.where(
+        (dt.x == float(dt_where["x"])) & (dt.y == float(dt_where["y"])),
+        drop=True).squeeze()
+    return (int(dt_where["x_idx"]), int(dt_where["y_idx"]))
+
+
+def get_geo_coords(x, y, dt):
+    # y = 1199
+    # x = 310
+    dt_where = dt.where((dt.x_idx == x) & (dt.y_idx == y), drop=True)
+    return (float(dt_where["x"]), float(dt_where["y"]))
+
+
+def get_distance(stop_idx_y, stop_idx_x, i, start_idx_y, start_idx_x,
+                 cost_surface_array, dt):
+    try:
+        indices, weight = route_through_array(
+            cost_surface_array,
+            (start_idx_y, start_idx_x),
+            (stop_idx_x, stop_idx_y),
+            geometric=True,
+            fully_connected=True,
+        )
+        indices = np.array(indices).T
+        weight = int(round(weight, 0))
+
+        path = np.zeros_like(cost_surface_array)
+        # set ONLY end point to weight
+        path[np.array([indices[0][-1]]), np.array([indices[1][-1]])] = weight
+        # path[indices[0], indices[1]] = weight
+
+        res = xr.DataArray(path,
+                           coords=[dt.y.values, dt.x.values],
+                           dims=["y", "x"])
+        res["spatial_ref"] = dt["spatial_ref"]
+        res = res.where(res > 0)
+
+        df = res.to_dataframe("cost").reset_index()
+        df = df[~pd.isna(df["cost"])]
+        gdf = gpd.GeoDataFrame({
+            "cost": df.cost,
+            "i": i
+        },
+                               geometry=gpd.points_from_xy(df.x, df.y))
+
+    except:
+        (lon, lat) = get_geo_coords(stop_idx_x, stop_idx_y, dt)
+        gdf = gpd.GeoDataFrame({
+            "cost": 999,
+            "i": i
+        },
+                               geometry=gpd.points_from_xy(lon, lat))
+
+    return gdf

### 2022-09-28 figures/fwi.py
towards fwi path figure, #6

@@ -1,0 +1,38 @@
+import sys
+import pickle
+import xarray as xr
+import pandas as pd
+import geopandas as gpd
+
+sys.path.append("src")
+from src import fwi
+
+stations = gpd.read_file("stations.gpkg", driver="GPKG")
+fpath = "data/cost.tif"
+dt = xr.open_dataset(fpath, engine="rasterio").sel(band=1)
+dt = dt.assign_coords(
+    x_idx=dt["x"] * 0 + [x for x in range(0, 1038)],
+    y_idx=dt["y"] * 0 + [x for x in range(0, 1448)],
+)
+cost_surface_array = dt.to_array().values[0, :, :]
+
+(start_idx_x,
+ start_idx_y) = fwi.get_idx_coords(stations.iloc[[0]]["longitude"][0],
+                                   stations.iloc[[0]]["latitude"][0], dt)
+
+end_points = pickle.load(open("data/end_points.pkl", "rb"))
+
+i = 171133
+# i=45050
+test = fwi.get_distance(end_points[i][1],
+                 end_points[i][0],
+                 i,
+                 start_idx_y,
+                 start_idx_x,
+                 cost_surface_array,
+                 dt,
+                 return_full_path=True)
+
+test.to_file("test.gpkg", driver="GPKG")
+
+discharge = pd.read_csv("data/discharge_median.csv")

### 2022-09-28 src/fwi.py

@@ -24,8 +24,14 @@ def get_geo_coords(x, y, dt):
     return (float(dt_where["x"]), float(dt_where["y"]))
 
 
-def get_distance(stop_idx_y, stop_idx_x, i, start_idx_y, start_idx_x,
-                 cost_surface_array, dt):
+def get_distance(stop_idx_y,
+                 stop_idx_x,
+                 i,
+                 start_idx_y,
+                 start_idx_x,
+                 cost_surface_array,
+                 dt,
+                 return_full_path=False):
     try:
         indices, weight = route_through_array(
             cost_surface_array,
@@ -55,6 +61,25 @@ def get_distance(stop_idx_y, stop_idx_x, i, start_idx_y, start_idx_x,
             "i": i
         },
                                geometry=gpd.points_from_xy(df.x, df.y))
+
+        if return_full_path:
+            full_path = path
+            for i in range(0, indices.shape[1]):
+                full_path[np.array([indices[0][i]]),
+                          np.array([indices[1][i]])] = weight
+
+            res_full = xr.DataArray(full_path,
+                                    coords=[dt.y.values, dt.x.values],
+                                    dims=["y", "x"])
+            res_full["spatial_ref"] = dt["spatial_ref"]
+            res_full = res_full.where(res_full > 0)
+            df = res_full.to_dataframe("cost").reset_index()
+            df = df[~pd.isna(df["cost"])]
+            gdf = gpd.GeoDataFrame({
+                "cost": df.cost,
+                "i": i
+            },
+                                   geometry=gpd.points_from_xy(df.x, df.y))
 
     except:
         (lon, lat) = get_geo_coords(stop_idx_x, stop_idx_y, dt)

### 2022-09-28 figures/fwi.py
actual plotting of fwi

@@ -1,11 +1,16 @@
 import sys
 import pickle
+import numpy as np
 import xarray as xr
 import pandas as pd
 import geopandas as gpd
+import cartopy.crs as ccrs
+import matplotlib.pyplot as plt
 
 sys.path.append("src")
 from src import fwi
+
+discharge = pd.read_csv("data/discharge_median.csv")
 
 stations = gpd.read_file("stations.gpkg", driver="GPKG")
 fpath = "data/cost.tif"
@@ -16,23 +21,43 @@ dt = dt.assign_coords(
 )
 cost_surface_array = dt.to_array().values[0, :, :]
 
-(start_idx_x,
- start_idx_y) = fwi.get_idx_coords(stations.iloc[[0]]["longitude"][0],
-                                   stations.iloc[[0]]["latitude"][0], dt)
 
-end_points = pickle.load(open("data/end_points.pkl", "rb"))
+def _get_path(j):
+    # j = 1
+    (start_idx_x, start_idx_y) = fwi.get_idx_coords(
+        stations.iloc[[j]]["longitude"].reset_index(drop=True)[0],
+        stations.iloc[[j]]["latitude"].reset_index(drop=True)[0], dt)
 
-i = 171133
-# i=45050
-test = fwi.get_distance(end_points[i][1],
-                 end_points[i][0],
-                 i,
-                 start_idx_y,
-                 start_idx_x,
-                 cost_surface_array,
-                 dt,
-                 return_full_path=True)
+    end_points = pickle.load(open("data/end_points.pkl", "rb"))
 
-test.to_file("test.gpkg", driver="GPKG")
+    i = 171133
+    path = fwi.get_distance(end_points[i][1],
+                            end_points[i][0],
+                            i,
+                            start_idx_y,
+                            start_idx_x,
+                            cost_surface_array,
+                            dt,
+                            return_full_path=True)
+    path["tributary"] = stations.iloc[[j]]["name"].reset_index(drop=True)[0]
+    return path
 
-discharge = pd.read_csv("data/discharge_median.csv")
+
+res = {
+    stations.iloc[[j]]["name"].reset_index(drop=True)[0]: _get_path(j)
+    for j in range(0, stations.shape[0])
+}
+
+b = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(epsg=4326)
+b_bounds = [x for x in b.bounds.iloc[0]]
[…]

### 2022-09-28 figures/fwi.py
fight with styling

@@ -6,11 +6,10 @@ import pandas as pd
 import geopandas as gpd
 import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
+from shapely.geometry import LineString
 
 sys.path.append("src")
 from src import fwi
-
-discharge = pd.read_csv("data/discharge_median.csv")
 
 stations = gpd.read_file("stations.gpkg", driver="GPKG")
 fpath = "data/cost.tif"
@@ -52,12 +51,34 @@ b = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(epsg=4326)
 b_bounds = [x for x in b.bounds.iloc[0]]
 extent = (b_bounds[0], b_bounds[2], b_bounds[1], b_bounds[3])
 
+test_lines = pd.concat([
+    gpd.GeoDataFrame(
+        {"tributary": key},
+        geometry=[LineString(res[key]['geometry'].reset_index(drop=True))],
+        index=[0]) for key in res.keys()
+])
 test = pd.concat([res[key] for key in res.keys()])
+discharge = pd.read_csv("data/discharge_median.csv")
 test = test.merge(discharge, left_on="tributary", right_on="site_str")
-test["discharge_va"] = np.log(test["discharge_va"] * 2)
+test["weight"] = np.log(test["discharge_va"] * 4)
+
+a = test["weight"]
+alphas = np.interp(a, (a.min(), a.max()), (0.7, +1))
 
 ax = plt.axes(projection=ccrs.PlateCarree())
 ax.set_extent(extent, ccrs.PlateCarree())
 ax.coastlines(resolution="10m", color="black", linewidth=1)
-test.plot(ax=ax, column="tributary", legend=True, markersize="discharge_va")
+test.plot(ax=ax, column="tributary", markersize="weight", legend=True)
+plt.show()
+
+ax = plt.axes(projection=ccrs.PlateCarree())
+ax.coastlines(resolution="10m", color="black", linewidth=1, alpha=1)
+test[test["tributary"] == "Choptank"].plot(ax=ax,                                           
+                                           markersize="weight",
+                                           alpha=0.1,
+                                           color="blue")
+test[test["tributary"] == "Potomac"].plot(ax=ax,                                           
+                                           markersize="weight",
+                                           alpha=1,
+                                           color="blue")
 plt.show()

### 2022-09-29 figures/fwi.py
make end_points pixel a fxn parameter

@@ -21,7 +21,7 @@ dt = dt.assign_coords(
 cost_surface_array = dt.to_array().values[0, :, :]
 
 
-def _get_path(j):
+def _get_path(j, i=171133):
     # j = 1
     (start_idx_x, start_idx_y) = fwi.get_idx_coords(
         stations.iloc[[j]]["longitude"].reset_index(drop=True)[0],
@@ -29,7 +29,6 @@ def _get_path(j):
 
     end_points = pickle.load(open("data/end_points.pkl", "rb"))
 
-    i = 171133
     path = fwi.get_distance(end_points[i][1],
                             end_points[i][0],
                             i,

### 2022-09-29 figures/fwi.py
settle on scaled lines for fwi figure

@@ -11,7 +11,6 @@ from shapely.geometry import LineString
 sys.path.append("src")
 from src import fwi
 
-stations = gpd.read_file("stations.gpkg", driver="GPKG")
 fpath = "data/cost.tif"
 dt = xr.open_dataset(fpath, engine="rasterio").sel(band=1)
 dt = dt.assign_coords(
@@ -41,6 +40,7 @@ def _get_path(j, i=171133):
     return path
 
 
+stations = gpd.read_file("stations.gpkg", driver="GPKG")
 res = {
     stations.iloc[[j]]["name"].reset_index(drop=True)[0]: _get_path(j)
     for j in range(0, stations.shape[0])
@@ -59,25 +59,29 @@ test_lines = pd.concat([
 test = pd.concat([res[key] for key in res.keys()])
 discharge = pd.read_csv("data/discharge_median.csv")
 test = test.merge(discharge, left_on="tributary", right_on="site_str")
-test["weight"] = np.log(test["discharge_va"] * 4)
 
-a = test["weight"]
-alphas = np.interp(a, (a.min(), a.max()), (0.7, +1))
+test["weight"] = test["discharge_va"] * test["cost"]
+test["weight"] = np.interp(test["weight"], (test["weight"].min(), test["weight"].max()), (0.1, +8))
+np.unique(test["weight"])
 
-ax = plt.axes(projection=ccrs.PlateCarree())
+fig = plt.figure(figsize=[3.4/1.4, 2.8/1.4])
+ax = fig.add_subplot(111, projection=ccrs.PlateCarree())
 ax.set_extent(extent, ccrs.PlateCarree())
 ax.coastlines(resolution="10m", color="black", linewidth=1)
-test.plot(ax=ax, column="tributary", markersize="weight", legend=True)
-plt.show()
+test.plot(ax=ax, column="tributary", markersize="weight", alpha=0.6)
+plt.savefig("figures/_fwi.pdf", bbox_inches='tight')
 
-ax = plt.axes(projection=ccrs.PlateCarree())
-ax.coastlines(resolution="10m", color="black", linewidth=1, alpha=1)
-test[test["tributary"] == "Choptank"].plot(ax=ax,                                           
-                                           markersize="weight",
-                                           alpha=0.1,
-                                           color="blue")
-test[test["tributary"] == "Potomac"].plot(ax=ax,                                           
-                                           markersize="weight",
-                                           alpha=1,
[…]

### 2022-09-29 scripts/00_calculate_waterdistance.py
calculate james waterdistance

@@ -47,9 +47,9 @@ def get_distance_grp(rng):
 
 stations = gpd.GeoDataFrame(
     {
-        "name": ["Choptank", "Susquehanna", "Pautexent", "Potomac"],
-        "longitude": [-75.900, -76.069, -76.692, -77.110],
-        "latitude": [38.811, 39.525, 38.664, 38.38],
+        "name": ["Choptank", "Susquehanna", "Pautexent", "Potomac", "James"],
+        "longitude": [-75.900, -76.069, -76.692, -77.110, -77.2767],
+        "latitude": [38.811, 39.525, 38.664, 38.38, 37.3238],
     }
 )
 stations = gpd.GeoDataFrame(
@@ -69,6 +69,22 @@ if not os.path.exists("data/end_points.pkl"):
     end_points = get_idx_water()
     pickle.dump(end_points, open("data/end_points.pkl", "wb"))
 end_points = pickle.load(open("data/end_points.pkl", "rb"))
+
+# --- james
+if not os.path.exists("data/waterdistance/james.gpkg"):
+    (start_idx_x, start_idx_y) = fwi.get_idx_coords(
+        stations.iloc[[4]].reset_index()["longitude"][0],
+        stations.iloc[[4]].reset_index()["latitude"][0],
+        dt
+    )
+    l = np.array_split(np.array(range(0, end_points.shape[0])), 6)    
+    [get_distance_grp(s) for s in l]
+
+    flist = glob.glob("test_*.gpkg")
+    gdfs = [gpd.read_file(f) for f in flist]
+    pd.concat(gdfs).to_file("james.gpkg")
+    [os.remove(f) for f in flist]
+
 
 # --- choptank
 if not os.path.exists("data/waterdistance/choptank.gpkg"):

### 2022-09-29 figures/maps.py
bettern hex styling

@@ -82,6 +82,16 @@ dt_melt = pd.melt(dt_filtered, id_vars=id_vars, value_vars=variables)
 
 dt_grps = [get_pnt_counts(dt_melt, variable) for variable in variables]
 
+mins = []
+maxs = []
+gdf_aggs = []
+for i in range(0, len(dt_grps)):
+    gdf = dt_grps[i]
+    gdf_agg = gdf.h3.geo_to_h3_aggregate(resolution=5)
+    mins.append(min(gdf_agg.reset_index()["count"]))
+    maxs.append(max(gdf_agg.reset_index()["count"]))
+    gdf_aggs.append(gdf_agg)
+
 fig, axs = plt.subplots(
     ncols=3,
     nrows=1,
@@ -89,21 +99,11 @@ fig, axs = plt.subplots(
     subplot_kw={"projection": ccrs.PlateCarree()},
 )
 
-mins = []
-maxs = []
-gdf_aggs = []
-for i in range(0, len(dt_grps)):
-    gdf = dt_grps[i]
-    gdf_agg = gdf.h3.geo_to_h3_aggregate(6)
-    mins.append(min(gdf_agg.reset_index()["count"]))
-    maxs.append(max(gdf_agg.reset_index()["count"]))
-    gdf_aggs.append(gdf_agg)
-
 for i in range(0, len(axs)):
     ax = axs[i]
     ax.set_title(variables[i])
     gdf_agg = gdf_aggs[i]
-    gdf_agg.plot("count", ax=ax, legend=False, vmax=max(maxs))
+    gdf_agg.plot("count", ax=ax, legend=False, vmax=max(maxs), alpha=0.8)
     ax.coastlines(resolution="10m", color="black", linewidth=1)
 
 scales = np.linspace(1, max(maxs), 7)

### 2022-10-03 figures/fwi.py
styling and subplot spacing

@@ -1,10 +1,12 @@
 import sys
 import pickle
+import matplotlib
 import numpy as np
 import xarray as xr
 import pandas as pd
 import geopandas as gpd
 import cartopy.crs as ccrs
+from matplotlib import gridspec
 import matplotlib.pyplot as plt
 from xrspatial.classify import quantile
 from shapely.geometry import LineString
@@ -150,16 +152,27 @@ def panel_add(i,
               format=None):
     if j is not None:
         ax = axs[i, j]
+        ax = plt.subplot(ax, xlabel="", projection=ccrs.PlateCarree())
+        ax.set_extent(extent, ccrs.PlateCarree())
         ax.coastlines(resolution="10m", color="black", linewidth=1)
     else:
         ax = axs[i]
+        ax = plt.subplot(ax, xlabel="", projection=ccrs.PlateCarree())
+        ax.set_extent(extent, ccrs.PlateCarree())
         ax.coastlines(resolution="10m", color="black", linewidth=1)
+
+    # https://matplotlib.org/stable/gallery/color/named_colors.html
+    cmap = matplotlib.colors.ListedColormap([
+        "royalblue", "cornflowerblue", "lightsteelblue", "powderblue",
+        "aliceblue"
+    ])
 
     im = geo_grid.plot.imshow(ax=ax,
                               vmin=vmin,
                               vmax=vmax,
+                              cmap=cmap,
                               cbar_kwargs={
-                                  "shrink": 0.5,
+                                  "shrink": 0.7,
                                   'label': '',
                                   'ticks': ticks
                               })
@@ -173,25 +186,16 @@ def panel_add(i,
                  x=0.9,
                  rotation="vertical")
 
-
-fig, axes = plt.subplots(
-    ncols=2,
-    nrows=2,
-    constrained_layout=True,
-    subplot_kw={
-        "projection": ccrs.PlateCarree(),
-        "xlabel": ""
-    },
-)
+    return ax
 
 
 def line_add(i, j, i_ends, axes):
     ax = axes[i, j]
+    ax = plt.subplot(ax, xlabel="", projection=ccrs.PlateCarree())
[…]

### 2022-10-03 figures/fwi.py
panel spacing

@@ -216,7 +216,7 @@ def line_add(i, j, i_ends, axes):
 ncol = 2
 nrow = 2
 fig = plt.figure(figsize=(ncol + 3, nrow + 3))
-axes = gridspec.GridSpec(nrow, ncol, wspace=0.0, hspace=0.0, top=1, right=0.9)
+axes = gridspec.GridSpec(nrow, ncol, wspace=0.0, hspace=0.05, top=1, right=0.9)
 
 line_add(0, 0, 4, axes)
 line_add(1, 0, 0, axes)
@@ -233,11 +233,11 @@ cbar_labels = [
 panel_add(0, axes, "", grids[0][0], j=1, vmax=5, format=cbar_labels[0])
 panel_add(1, axes, "", grids[1][0], j=1, vmax=5, format=cbar_labels[1])
 
-# plt.subplots_adjust(right=0)
+# plt.show()
+plt.savefig("figures/_fwi_panels.pdf", bbox_inches='tight')
 
-plt.show()
 
-# ---
+# --- towards overall waterdistance plot
 
 ncol = 1
 nrow = 1

### 2022-10-03 figures/importance.py
sort in decending order

@@ -14,7 +14,7 @@ def imp_plot(variable):
     dset = pd.DataFrame()
     dset["attr"] = feature_names
     dset["importance"] = rfecv.estimator_.feature_importances_
-    dset = dset.sort_values(by="importance", ascending=False)
+    dset = dset.sort_values(by="importance", ascending=True)
     dset.loc[dset["attr"] == "cost", "attr"] = "fwi"
 
     plt.figure(figsize=(3, 5))

### 2022-10-03 figures/timeseries.py
only use salinity surfaces

@@ -14,6 +14,9 @@ flist = list(
 flist = list(
     itertools.compress(flist,
                        [not strng.__contains__("2018") for strng in flist]))
+flist = list(
+    itertools.compress(flist,
+                       [strng.__contains__("salinity") for strng in flist]))
 
 dates = [x.split("\\")[1].replace(".tif", "").replace("_salinity", "") for x in flist]
 
@@ -34,6 +37,7 @@ def _extract(i):
 
 res_raw = [_extract(i) for i in range(0, len(dates))]
 res = pd.concat(res_raw).reset_index(drop=True)
+res["date"] = [x.replace("_salinity", "") for x in res["date"]]
 res["date"] = pd.to_datetime(res["date"])
 res["month"] = [x.strftime("%m") for x in res["date"]]
 x_dates = res['date'].dt.strftime('%m').sort_values().unique()

### 2022-10-03 figures/study_site.py
towards study site figure

@@ -1,0 +1,18 @@
+import geopandas as gpd
+import cartopy.crs as ccrs
+import matplotlib.pyplot as plt
+from matplotlib import gridspec
+
+b = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(epsg=4326)
+b_bounds = [x for x in b.bounds.iloc[0]]
+extent = (b_bounds[0], b_bounds[2], b_bounds[1], b_bounds[3])
+
+ncol = 1
+nrow = 1
+fig = plt.figure(figsize=(ncol + 3, nrow + 3))
+axes = gridspec.GridSpec(nrow, ncol, wspace=0.0, hspace=0.0, top=1, right=0.9)
+ax = plt.subplot(axes[0], xlabel="", projection=ccrs.PlateCarree())
+ax.set_extent(extent, ccrs.PlateCarree())
+ax.coastlines(resolution="10m", color="black", linewidth=1)
+
+plt.show()

### 2022-10-03 figures/study_site.py
pretty-fi

@@ -1,7 +1,60 @@
+# https://stackoverflow.com/a/35705477/3362993
+# https://scitools.org.uk/cartopy/docs/v0.13/examples/tick_labels.html
+# https://coolum001.github.io/cartopylayout.html
+
+import numpy as np
 import geopandas as gpd
 import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
 from matplotlib import gridspec
+from cartopy.mpl.ticker import LongitudeFormatter, LatitudeFormatter
+
+
+def scale_bar(ax, length=None, location=(0.5, 0.05), linewidth=3):
+    """
+    ax is the axes to draw the scalebar on.
+    length is the length of the scalebar in km.
+    location is center of the scalebar in axis coordinates.
+    (ie. 0.5 is the middle of the plot)
+    linewidth is the thickness of the scalebar.
+    """
+    llx0, llx1, lly0, lly1 = ax.get_extent(ccrs.PlateCarree())
+    # Make tmc horizontally centred on the middle of the map,
+    # vertically at scale bar location
+    sbllx = (llx1 + llx0) / 2
+    sblly = lly0 + (lly1 - lly0) * location[1]
+    tmc = ccrs.TransverseMercator(sbllx, sblly)
+    tmc_text = ccrs.TransverseMercator(sbllx, sblly-0.03)
+    # Get the extent of the plotted area in coordinates in metres
+    x0, x1, y0, y1 = ax.get_extent(tmc)
+    # Turn the specified scalebar location into coordinates in metres
+    sbx = x0 + (x1 - x0) * location[0]
+    sby = y0 + (y1 - y0) * location[1]
+
+    #Calculate a scale bar length if none has been given
+    if not length:
+        length = (x1 - x0) / 5000  #in km
+        ndim = int(np.floor(np.log10(length)))  #number of digits in number
+        length = round(length, -ndim)  #round to 1sf
+
+        #Returns numbers starting with the list
+        def scale_number(x):
+            if str(x)[0] in ['1', '2', '5']: return int(x)
+            else: return scale_number(x - 10**ndim)
+
+        length = scale_number(length)
+
+    #Generate the x coordinate for the ends of the scalebar
+    bar_xs = [sbx - length * 500, sbx + length * 500]
[…]

### 2022-10-03 figures/study_site.py
add labels

@@ -3,6 +3,7 @@
 # https://coolum001.github.io/cartopylayout.html
 
 import numpy as np
+import pandas as pd
 import geopandas as gpd
 import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
@@ -24,7 +25,7 @@ def scale_bar(ax, length=None, location=(0.5, 0.05), linewidth=3):
     sbllx = (llx1 + llx0) / 2
     sblly = lly0 + (lly1 - lly0) * location[1]
     tmc = ccrs.TransverseMercator(sbllx, sblly)
-    tmc_text = ccrs.TransverseMercator(sbllx, sblly-0.03)
+    tmc_text = ccrs.TransverseMercator(sbllx, sblly - 0.03)
     # Get the extent of the plotted area in coordinates in metres
     x0, x1, y0, y1 = ax.get_extent(tmc)
     # Turn the specified scalebar location into coordinates in metres
@@ -56,6 +57,17 @@ def scale_bar(ax, length=None, location=(0.5, 0.05), linewidth=3):
             verticalalignment='top')
 
 
+stations = gpd.read_file("stations.gpkg", driver="GPKG")
+stations_jitter = pd.DataFrame({
+    "name": ["Choptank", "Susquehanna", "Pautexent", "Potomac", "James"],
+    "longitude": [-75.800, -76.469, -76.692, -77.110, -77.1767],
+    "latitude": [38.851, 39.525, 38.724, 38.2, 37.3838],
+})
+stations_jitter = gpd.GeoDataFrame(stations_jitter,
+                                   geometry=gpd.points_from_xy(
+                                       stations_jitter["longitude"],
+                                       stations_jitter["latitude"]))
+
 b = gpd.read_file("data/Boundaries/chk_water_only.shp").to_crs(epsg=4326)
 b_bounds = [x for x in b.bounds.iloc[0]]
 extent = (b_bounds[0], b_bounds[2], b_bounds[1], b_bounds[3])
@@ -66,7 +78,22 @@ fig = plt.figure(figsize=(ncol + 3, nrow + 3))
 axes = gridspec.GridSpec(nrow, ncol, wspace=0.0, hspace=0.0, top=1, right=0.9)
 ax = plt.subplot(axes[0], xlabel="", projection=ccrs.PlateCarree())
 ax.set_extent(extent, ccrs.PlateCarree())
+# ---
+ax.text(-75.98,
+        38.04,
[…]

### 2022-10-04 figures/obs_stats.py
log xform correct panel

@@ -32,7 +32,7 @@ dt_melt = pd.melt(
 ]
 g = sns.FacetGrid(dt_melt, col="variable", sharex=False)
 g.map(sns.histplot, "value")
-g.axes[0][1].set_xscale("log")
+g.axes[0][2].set_xscale("log")
 # plt.show()
 plt.savefig("figures/obs_varhist.pdf")
 

### 2022-10-04 figures/validation.py
verify that posthoc offset is the same as fitted_sine

@@ -1,5 +1,6 @@
 import sys
 import pickle
+import numpy as np
 import pandas as pd
 import seaborn as sns
 import matplotlib.pyplot as plt
@@ -17,17 +18,21 @@ def _get_predictions(variable):
 
     X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
     y_test = pickle.load(open("data/y_test_" + variable + ".pkl", "rb"))
+    imp_params = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
+    fitted_sine = X_test[:,int(np.where([x == "fitted_sine" for x in imp_params])[0])]    
 
     predictions = rf_random.predict(X_test)
-    res = pd.DataFrame(y_test, columns=["obs"])
+    res = pd.DataFrame(y_test + fitted_sine, columns=["obs"])
     res["predict"] = predictions.copy()
 
     if variable == "temperature":
-        # back out non-sine adjusted values
-        p1 = pickle.load(open("data/temperature_sine_coef.pkl", "rb"))
-        offset = fit_sine.fitfunc(p1, X_test[0:, 0])
-        res["predict"] = res["predict"].copy() + offset
-        res["obs"] = res["obs"].copy() + offset
+        # back out non-sine adjusted values        
+        # p1 = pickle.load(open("data/temperature_sine_coef.pkl", "rb"))
+        # offset = fit_sine.fitfunc(p1, X_test[0:, 0])
+        # offset should be the same as fitted_sine
+        # plt.plot(offset, fitted_sine)
+        # plt.show()
+        res["predict"] = res["predict"].copy() + fitted_sine
 
     return res
 

### 2022-10-04 src/fit_sine.py
show more detailed fitted sine plot

@@ -52,4 +52,12 @@ def fit_sine(data):
     [label.set_fontweight('bold') for label in labels]
     plt.savefig('figures/fitted_sine.png')
 
+    fig = plt.figure()
+    ax = fig.add_subplot(111)
+    ax.plot(date_doy, temp_doy, "ro")
+    ax.plot(date_doy, fitfunc(p1, date_doy), "bo", markersize=2)
+    plt.xlabel("DOY", fontweight="bold")
+    plt.tight_layout()
+    plt.savefig('figures/fitted_sine_detail.png')
+
     return fitted_sine

### 2022-10-04 figures/validation.py
only do fitted_sine with temperature var

@@ -18,21 +18,22 @@ def _get_predictions(variable):
 
     X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
     y_test = pickle.load(open("data/y_test_" + variable + ".pkl", "rb"))
-    imp_params = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
-    fitted_sine = X_test[:,int(np.where([x == "fitted_sine" for x in imp_params])[0])]    
 
     predictions = rf_random.predict(X_test)
-    res = pd.DataFrame(y_test + fitted_sine, columns=["obs"])
+    res = pd.DataFrame(y_test, columns=["obs"])
     res["predict"] = predictions.copy()
 
     if variable == "temperature":
+        imp_params = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
+        fitted_sine = X_test[:,int(np.where([x == "fitted_sine" for x in imp_params])[0])]
+        res["predict"] = res["predict"].copy() + fitted_sine
+        res["obs"] = res["obs"].copy() + fitted_sine
         # back out non-sine adjusted values        
         # p1 = pickle.load(open("data/temperature_sine_coef.pkl", "rb"))
         # offset = fit_sine.fitfunc(p1, X_test[0:, 0])
         # offset should be the same as fitted_sine
         # plt.plot(offset, fitted_sine)
-        # plt.show()
-        res["predict"] = res["predict"].copy() + fitted_sine
+        # plt.show()        
 
     return res
 

### 2022-10-04 scripts/01_rf_fit.py
towards debugging negative temperatures

@@ -53,6 +53,9 @@ def main():
                                     variable,
                                     overwrite=True)
 
+    # TODO: if temperature, save minimum temperature, add it to the series to make everything positive
+    breakpoint()
+
     pickle.dump(X_train, open("data/X_train_" + variable + ".pkl", "wb"))
     pickle.dump(X_test, open("data/X_test_" + variable + ".pkl", "wb"))
     pickle.dump(y_train, open("data/y_train_" + variable + ".pkl", "wb"))

### 2022-10-05 manuscript/manuscript.tex
cite darwin
sync ms text

@@
     \noindent Remotely sensed water properties are important for a variety of applications, including validation of Earth Systems models (ESMs), habitat suitability models, and sea level rise projections.
-For the validation of next-generation ESMs in particular, the usefulness of operational forecasting products and directly-observing satellite-based sensors for validation of next-generation ESMs is limited due to their temporal availability and spatial resolution of $<$ 1 year and $>$ 30 $km^2$ respectively.
+For the validation of next-generation, high or multi-resolution (30-60 km) ESMs in particular, the usefulness of operational forecasting products and directly-observing satellite-based sensors for validation is limited due to their temporal availability and spatial resolution of $<$ 1 year (in some cases) and $>$ 30 $km^2$ respectively.
 To address this validation data gap, we developed a data-driven model to produce high-resolution ($<$ 1 $km^2$) estimates of temperature, salinity, and turbidity over decadal time scales as required by next-generation ESMs.
@@
 The resulting models have similar error statistics as prior efforts of this type for salinity (RMSE: 2.2) and temperature (RMSE: 1.9 C).
-However, unlike prior efforts our model is set up as a pipeline meaning that it has the advantage of producing predictions of water properties in future time periods as additional MODIS data is collected.
-In addition, our study is unique in that the predictions produced by our model are “geographically-aware” insofar as they capture geographic variation in the influence of flow and surface water exchange in upstream coastal watersheds.
+However, unlike prior efforts our model is designed as a pipeline meaning that it has the advantage of producing predictions of water properties in future time periods as additional MODIS data becomes available.
+We also include novel  “geographically-aware” predictive features insofar as they capture geographic variation in the influence of flow and surface water exchange in upstream coastal watersheds.
     \end{abstract} \hspace{10pt}
@@
 In partially mixed estuaries such as the Chesapeake Bay (hereafter, “the Bay”), productivity, organic matter production, and phytoplankton community composition are largely determined by stratification \citep{xuClimateForcingSalinity2012}.
-Because stratification is primarily controlled by salinity and temperature, extensive measurement networks exist to monitor these water properties (Chesapeake Bay Program Water Quality Database, \url{https://www.chesapeakebay.net/what/downloads/cbp_water_quality_database_1984_present}).
-These data are critical for many uses such as establishing point-based trends and relationships (e.g.
-river flow or sea level rise vs salinity or hypoxia), \citet{hagyHypoxiaChesapeakeBay2004}.
+Because stratification is controlled by salinity and temperature, extensive measurement networks exist to monitor these water properties (Chesapeake Bay Program Water Quality Database, \url{https://www.chesapeakebay.net/what/downloads/cbp_water_quality_database_1984_present}).
+These data are critical for establishing point-based trends and relationships between river flow or sea level rise and salinity or hypoxia \citet{hagyHypoxiaChesapeakeBay2004}.
 Of particular importance is monitoring for increased nutrient loads as result of expected increases in precipitation, tributary discharge, and river flows as a result of climate change \citep{najjarPotentialClimatechangeImpacts2010, irbyCompetingImpactsClimate2018}.
@@
 Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \citep{hoodChesapeakeBayProgram2021}.
-For example, habitat suitability calculations are typically done on a gridded basis for the entirety of a particular estuary rather than at selected point locations.
+For example, habitat suitability assessments are typically done on a gridded basis for the entirety of a particular estuary rather than at selected point locations.
 Applications such as these often turn to gridded alternatives to direct observations such as output from directly-observing satellite-based sensors \citep{foreCombinedActivePassive2016} or operational now/fore-casting products \citep{lanerolle2011second}.
@@
 The primary input to our model is remotely sensed reflectance data.
-In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in the Bay as well as in coastal Louisiana \citep{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}.
-We supplemented MODIS data with gridded estimates of tributary inflows which we used to enforce physical realism (i.e.
+In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \citep{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}.
+We supplemented MODIS data with gridded estimates of distance-weighted tributary inflows which we used to encourage physical realism (i.e.
 geographic “awareness”) especially in areas of the Bay with little to no observations.
 Our model represents a further advantage over prior efforts at data-driven modeling of salinity in the Bay (e.g.
-\citet{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
+\citet{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
+These more spatially extenstive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
 
@@
 The Bay is fairly shallow ($<$ 30m, \citet{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \citep{xuClimateForcingSalinity2012}.
+For the present study, this means that surface water properties are likely to differ substantially from water properties at deeper depths.
+It also means that the Bay receives enough freshwater inputs to produce a distinct upper-estuary to lower-estuary gradient in water properties.
 
@@
 We chose to use the Random Forest algorithm in part because \citet{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) and also because of its ease of implementation and interpretation.
-We fit RF models using the scikit-learn Python package \citep{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model.
+We fit RF models using the scikit-learn Python package \citep{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model (Figure 1).
 We subsequently “tuned” each model using k-fold cross validation to optimize the size, depth, and splitting criteria of each tree.
@@
 \begin{sloppypar}
-
 We compiled over two decades of observational data (2000-2021) on salinity (n=6k), temperature (n=10k), and turbidity (n=5k, Table S1).
@@
 The temporal range of the observations (2000-2021) includes the full range of freshwater input magnitudes to the Bay spanning both times of minimum and maximum freshwater input (\url{https://www.usgs.gov/media/images/estimated-annual-mean-streamflow-entering-chesapeake-bay}).
-
 \end{sloppypar}
@@
 Rather than straight-line distances we used “in-water” distances calculated using the scikit-image Python package \citep{littleKrigingEstuariesCrow1997,vanderwaltScikitimageImageProcessing2014}.
-We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands values for each pixel spanning from 405 to 877nm wavelength \citep{vermoteericMOD09GAMODISTerra2015}.
+Our in-water distance calculations involved creating a cost–surface whereby land pixels had a very high cost relative to water pixels.
+Then the distance from any one cell to a tributary source was the accumulated travel “cost” as given by the scikit-image \texttt{route\_through\_array} function.
+We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands of values for each pixel spanning from 405 to 877nm wavelength \citep{vermoteericMOD09GAMODISTerra2015}.
+
+All code for data processing, model fitting, and model evaluation is available at Stachelek et al. (2023). All data used in the study are available in Stachelek et al. (2023).
 
@@
 
-Our tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 1).
-Out-of-sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 1, Figure S1).
+Our tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 2).
+Out-of-sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 2, Figure S1).
 Despite excellent overall performance for the temperature and salinity models, there was some indication of poor performance at the extremes.
-They were unable to recover very low temperature values ($<$ 0 C) or very high temperature and salinity values ($>$ 30 C, $>$ 23 salinity, Figure 1).
+The models were unable to recover very low temperature values ($<$ 0 C) or very high temperature and salinity values ($>$ 30 C, $>$ 23 salinity, Figure 2).
 One explanation for these areas of poor performance is that such extremes were rare in the observational dataset (Table S2).
-The salinity model however did faithfully capture areas of low salinity (Figure 1).
+The salinity model however did faithfully capture areas of low salinity (Figure 2).
 These low salinity values can be regarded as low uncertainty (high confidence) estimates given the extent to which the upper tributaries are well represented in the observational dataset (Figure S2).
+
+The date of observation was the most important feature identified in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 3).
+Across all models, the most important MODIS band was band 8, which is the shortest wavelength band of the MODIS Aqua product at the extreme of the visible light range.
+The “size” (i.e.
+n\_estimators) of the final RF models were of a similar magnitude for each variable but the tree was much “deeper” (i.e.
+max\_depth) for temperature compared to the other variables (Table S3).
+
+Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 4).
+A realistic spatial trend can also be seen in the RF temperature predictions whereby maximum water temperatures were located in the upper tributaries while minimum water temperatures were highest at the Bay mouth (Figure 4).
+
+RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 5).
+CBOFS surface water salinity estimates are known to be saltier than observations \citep{lanerolle2011second, vogelAssessingSatelliteSea2016} partly explaining our saltier data-driven results.
+The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 4).
 
@@
 
-The date of observation was the most important feature identified in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 2).
-Across all models, the most important MODIS band was band 8, which is the shortest wavelength band of the MODIS Aqua product at the extreme of the visible light range.
-The “size” (i.e.
-n\_estimators) of the final RF models were of a similar magnitude for each variable but the tree was much “deeper” (i.e.
-max\_depth) for temperature compared to the other variables (Table S3).
-
 \begin{figure}[ht!]
@@
 \end{figure}
-
-Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 3).
-A realistic spatial trend can also be seen in the RF temperature predictions whereby maximum water temperatures were located in the upper tributaries while minimum water temperatures were highest at the Bay mouth (Figure 3).
-
-RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 4).
-CBOFS surface water salinity estimates are known to be saltier than observations \citep{lanerolle2011second, vogelAssessingSatelliteSea2016} partly explaining our saltier data-driven results.
-The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 3).
 
@@
 Model errors were generally lower than prior data-driven efforts and as good or better than physics-based models \citep{vogelAssessingSatelliteSea2016}.
-In addition to being accurate, our approach provides the ability to generate rapid bay-wide prediction surfaces in seconds and it is easily updatable insofar as saved models can be reloaded and have their training “updated”.
+In addition to improved accuracy, our approach provides the ability to generate rapid bay-wide prediction surfaces in seconds and it is easily updatable insofar as saved models can be updated as new data become available.
 Our models are lightweight with minimal internal complexity and have been extensively tested against a uniquely large observational dataset.
@@
 Although salinity and temperature models were accurate, our turbidity model showed relatively poor performance.
-One reason for this may have been because our observational data only contains data in Nephelometric Turbidity Units (NTUs) which are not standardized units.
+One reason for this may have been because our observational data were reported as Nephelometric Turbidity Units (NTUs) which are not standardized.
 As a result, they cannot be reliably compared across different instruments.
 A better turbidity model would likely result from using total suspended matter (TSM) data measured in a standardized unit such as mg/l \citep{ondrusekDevelopmentNewOptical2012}.
+Another reason the turbidity model performed poorly may be that turbidity is simply difficult to predict.
+The event-driven nature of turbidity may mean that its extreme spikes may occur over time scales which are too short to be captured by daily reflectance data.
 
@@
 
-Often it is difficult to assess the inherent spatial uncertainty of data-driven efforts such as ours.
-However, we assembled such a large observational dataset, that it makes such assessments possible in the present study.
-This is a key difference between our approach and prior efforts which have mostly been trained on limited (and relatively high-salinity) buoy data or on data from in-situ cruises located in the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013, ondrusekDevelopmentNewOptical2012}.
+Our comprehensive dataset includes greater spatial coverage than prior efforts that were  largely trained on limited (and relatively high-salinity) buoy or in-situ cruise data located in the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013, ondrusekDevelopmentNewOptical2012}.
+This coverage allows for a more comprehensive assessment of the spatial uncertainty of our model.
 Predicted areas of low salinity coming from our model likely have particularly low associated uncertainties because the upper tributaries were well-represented in our observational dataset (Figure S2).
@@
 This appears to be the collision point of the marine plume emanating from the Bay mouth with the discharge of the Potomac River.
-The hydrodynamics here seem to be complex given that the CBOFS has difficulty resolving spatial patterning in this area.
+The hydrodynamics here seem to be complex given that the CBOFS has difficulty resolving spatial patterning in this area \citep{lanerolle2011second}.
 This was an area where we found limited observational data (Figure S2).
@@
 For the temperature model, we were unable to recover low temperature observations less than 0 deg C.
-These results are consistent with prior efforts such as \citet{geigerSatellitederivedCoastalOcean2013} where winter had the highest seasonal error.
+These results are consistent with prior research showing winter had the highest seasonal error  \citet{geigerSatellitederivedCoastalOcean2013}.
 One explanation for these high winter error rates is the fact that winter in the Bay is generally characterized by strong vertical mixing due to high winds \citep{sonWaterPropertiesChesapeake2012, xuClimateForcingSalinity2012} making “true” patterns more dynamic and more difficult to resolve.
@@
 
-One shortcoming of our work which may be a fruitful area of future research is the lack of realistic small-scale features evident in the physics based output of the CBOFS (Figure 4).
+One shortcoming of our work which may be a fruitful area of future research is the lack of realistic small-scale features evident in the physics based output of the CBOFS (Figure 5).
 It may be possible to resolve small-scale features such as eddies and other circulation features with the use of convolutional neural networks (CNNs).
@@
 the “neighborhood”) surrounding observations as is the case with CNNs \citep{goodfellow2016convolutional}.
+The pipeline for training a CNN would bring some practical difficulties regarding the collection and storage of MODIS data.
+Unlike our RF approach, it would require obtaining entire multiband images for the Bay rather than simply the values that correspond with observational data.
 
@@
 We have demonstrated an approach to produce realistic estimates of water properties in Chesapeake Bay using remotely sensed data.
-Much of this realism comes from our FWI routines which enforce physical consistency in upstream-downstream spatial patterning.
+Salinity and temperature model performance was as good or better than prior efforts.
+Salinity model performance was particularly realistic in low salinity tributary areas.
+Much of this realism comes from our FWI routines which encourage physical consistency in upstream-downstream spatial patterning.
+We saw additional performance gains owing tothe fact that our training dataset had spatial coverage in the upper tributaries beyond the mainstem of the Bay.
 Estimates produced by our model have a high potential to inform calibration and evaluation of Earth Systems Models as well as in other applications such as habitat suitability and sea level rise models.
@@
 Department of Energy, Office of Science, Office of Biological and Environmental Research as part of the multi-program, collaborative Integrated Coastal Modeling (ICoM) project.
+This research used resources provided by the Darwin testbed at Los Alamos National Laboratory (LANL) which is funded by the Computational Systems and Software Environments subprogram of LANL's Advanced Simulation and Computing program (NNSA/DOE).
 

### 2022-10-05 src/fit_sine.py
more detailed sine fitting plots

@@ -3,6 +3,7 @@
 import pickle
 import numpy as np
 import pandas as pd
+import seaborn as sns
 from scipy import optimize
 import matplotlib.pyplot as plt
 
@@ -59,5 +60,15 @@ def fit_sine(data):
     plt.xlabel("DOY", fontweight="bold")
     plt.tight_layout()
     plt.savefig('figures/fitted_sine_detail.png')
+    
+    # data_fitted = pd.DataFrame({"datetime":date_doy, "fit":fitfunc(p1, date_doy)})
+    # test = data.merge(data_fitted)
+    # test["diff"] = test["SST (C)"] - test["fit"]
+    # fig, ax = plt.subplots()
+    # sns.scatterplot(x="datetime", y="SST (C)", data=test, ax=ax)
+    # sns.lineplot(x="datetime", y="fit", data=test, ax=ax)
+    # plt.show()
+    # sns.scatterplot(x="datetime", y="diff", data = test)
+    # plt.show()
 
     return fitted_sine

### 2022-10-05 figures/validation.py
create map of model errrors, closes #4

@@ -1,9 +1,13 @@
 import sys
+import math
 import pickle
 import numpy as np
 import pandas as pd
 import seaborn as sns
+import geopandas as gpd
+import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
+import h3pandas  # h3.geo_to_h3_aggregate
 from sklearn.linear_model import LinearRegression
 
 sys.path.append(".")
@@ -135,3 +139,59 @@ r2_df["variable"] = variables_str_short
 r2_df.to_csv("data/r2.csv", index=False)
 
 plt.savefig("figures/_validation.pdf")
+
+# --- map of model errors
+
+def get_rmse(i):
+    variable = variables_str_short[i]
+    imp_params = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
+    X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
+    latitude = X_test[:,int(np.where([x == "latitude" for x in imp_params])[0])]
+    longitude = X_test[:,int(np.where([x == "longitude" for x in imp_params])[0])]
+    res[i]["latitude"] = latitude
+    res[i]["longitude"] = longitude
+    res[i]["diff2"] = np.power(res[1]["predict"] - res[1]["obs"], 2)
+
+    test = res[i].groupby(["latitude", "longitude"]).mean("diff2").reset_index()
+    test["rmse"] = test["diff2"]**(1/2)
+    test = gpd.GeoDataFrame(test, geometry = gpd.points_from_xy(test["longitude"], test["latitude"]))
+    return test
+
+dt_grps = [get_rmse(i) for i in range(0, len(variables_str_short))]
+
+mins = []
+maxs = []
+gdf_aggs = []
+for i in range(0, len(dt_grps)):
+    gdf = dt_grps[i]
+    gdf_agg = gdf.h3.geo_to_h3_aggregate(resolution=6)
+    mins.append(min(gdf_agg.reset_index()["rmse"]))
+    maxs.append(max(gdf_agg.reset_index()["rmse"]))
+    gdf_aggs.append(gdf_agg)
+
+
+fig, axs = plt.subplots(
+    ncols=3,
+    nrows=1,
+    constrained_layout=True,
+    subplot_kw={"projection": ccrs.PlateCarree()},
+)
+
+for i in range(0, len(axs)):
+    ax = axs[i]
[…]

### 2022-10-05 figures/rf_stats_table.py
formatting rf stats table

@@ -5,30 +5,33 @@ from tabulate import tabulate
 sys.path.append(".")
 from src import utils
 
+# --- rf params table
 dt = pd.read_csv("data/best_params_rf.csv")
 dt = dt.drop(columns=["bootstrap", "min_samples_split"])
 variables = dt.pop("variable")
 dt.insert(0, "variable", variables)
 utils.tabulate_to_latex(
-    tabulate(
-        dt.values, headers=[x for x in dt.columns], tablefmt="latex"
-    ),
+    tabulate(dt.values, headers=[x for x in dt.columns], tablefmt="latex"),
     "figures/rf_stats_table_0_.tex",
     "Random Forest tuning results.",
     2,
 )
 
+# --- rf stats table
 dt = pd.read_csv("data/rmse_rf.csv").round(2)
-dt = dt.T.reset_index().rename(columns={"index":"variable", 0:"rmse"})
+dt = dt.T.reset_index().rename(columns={"index": "variable", 0: "rmse"})
 col_names = dt["variable"].copy()
 dt = dt.merge(pd.read_csv("data/r2.csv"), on="variable").T[1:3].reset_index()
 dt.columns = ["metric"] + [x for x in col_names]
+headers = [x for x in dt.columns]
+headers[0] = ""
+dt["metric"] = ["RMSE", "$R^2$"]
+
+tblate = tabulate(dt.values, headers=[x for x in dt.columns], tablefmt="latex_raw")
 
 utils.tabulate_to_latex(
-    tabulate(
-        dt.values, headers=[x for x in dt.columns], tablefmt="latex"
-    ),
+    tblate,
     "figures/rf_stats_table_1_.tex",
-    "Random Forest fitting results. Values shown are root mean square error.",
+    "Random Forest model fit and predictive accuracy metrics. (RMSE = root mean square error, $R^2$ = coefficient of determination)",
     0,
 )

### 2022-10-05 src/utils.py

@@ -77,6 +77,7 @@ def tabulate_to_latex(tabulate_output, fname, caption='nasdf', table_num=1):
         f.write("\\usepackage{standalone}" + "\r\n")
         f.write("\\usepackage[justification=centering]{caption}" + "\r\n")
         f.write("\\usepackage{adjustbox}" + "\r\n")
+        f.write("\\usepackage{amsmath}" + "\r\n")
         f.write("\\setcounter{table}{" + str(table_num) + "}" + "\r\n")
         f.write("\\renewcommand\\thetable{\\arabic{table}}" + "\r\n")
         f.write("\\pagestyle{empty}" + "\r\n")

### 2022-10-05 manuscript/manuscript.tex
prep pdf ms

@@
 
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/data-processing_model-architecture}
+          \caption{Diagram of data processing, model fitting, tuning, and prediction pipeline.}
+    \end{center}    
+\end{figure}
+
 \subsection{Data description}
@@
 Our tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 2).
-Out-of-sample validation performance was excellent for salinity and temperature models but poor for the turbidity model (Figure 2, Figure S1).
+Out-of-sample validation performance was good for salinity and temperature models but poor for the turbidity model (Figure 2, Figure S1).
 Despite excellent overall performance for the temperature and salinity models, there was some indication of poor performance at the extremes.
@@
 One explanation for these areas of poor performance is that such extremes were rare in the observational dataset (Table S2).
-The salinity model however did faithfully capture areas of low salinity (Figure 2).
+The salinity model however did faithfully capture areas of low salinity (Figure 2, S3).
 These low salinity values can be regarded as low uncertainty (high confidence) estimates given the extent to which the upper tributaries are well represented in the observational dataset (Figure S2).
-
-The date of observation was the most important feature identified in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 3).
-Across all models, the most important MODIS band was band 8, which is the shortest wavelength band of the MODIS Aqua product at the extreme of the visible light range.
-The “size” (i.e.
-n\_estimators) of the final RF models were of a similar magnitude for each variable but the tree was much “deeper” (i.e.
-max\_depth) for temperature compared to the other variables (Table S3).
-
-Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 4).
-A realistic spatial trend can also be seen in the RF temperature predictions whereby maximum water temperatures were located in the upper tributaries while minimum water temperatures were highest at the Bay mouth (Figure 4).
-
-RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 5).
-CBOFS surface water salinity estimates are known to be saltier than observations \citep{lanerolle2011second, vogelAssessingSatelliteSea2016} partly explaining our saltier data-driven results.
-The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 4).
 
@@
 
+The date of observation was the most important feature identified in RF training for the temperature and turbidity models while location of observation was most important in the salinity model (Figure 3).
+Across all models, the most important MODIS band was band 8, which is the shortest wavelength band of the MODIS Aqua product at the extreme of the visible light range.
+The “size” (i.e.
+n\_estimators) of the final RF models were of a similar magnitude for each variable but the tree was much “deeper” (i.e.
+max\_depth) for temperature compared to the other variables (Table S3).
+
 \begin{figure}[ht!]
@@
 
+Running the models in prediction mode produced surfaces with a realistic spatial trend whereby salinity was highest near the Bay mouth and decreased with distance from each tributary mouth (Figure 4).
+A realistic spatial trend can also be seen in the RF temperature predictions whereby maximum water temperatures were located in the upper tributaries while minimum water temperatures were highest at the Bay mouth (Figure 4).
+
 \begin{figure}[ht!]
@@
 \end{figure}
+
+RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 5).
+CBOFS surface water salinity estimates are known to be saltier than observations \citep{lanerolle2011second, vogelAssessingSatelliteSea2016} partly explaining our saltier data-driven results.
+The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 4).
 

### 2022-10-05 figures/rf_stats_table.py

@@ -19,6 +19,7 @@ utils.tabulate_to_latex(
 
 # --- rf stats table
 dt = pd.read_csv("data/rmse_rf.csv").round(2)
+dt = dt[["salinity", "temperature", "turbidity"]]
 dt = dt.T.reset_index().rename(columns={"index": "variable", 0: "rmse"})
 col_names = dt["variable"].copy()
 dt = dt.merge(pd.read_csv("data/r2.csv"), on="variable").T[1:3].reset_index()

### 2022-10-06 scripts/rf-kerastf.py
towards putting the rf model on gee

@@ -1,0 +1,48 @@
+# https://www.tensorflow.org/decision_forests/tutorials/beginner_colab#training_a_regression_model
+
+import math
+import requests
+import numpy as np
+import pandas as pd
+import tensorflow_decision_forests as tfdf
+
+def split_dataset(dataset, test_ratio=0.30):
+  """Splits a panda dataframe in two."""
+  test_indices = np.random.rand(len(dataset)) < test_ratio
+  return dataset[~test_indices], dataset[test_indices]
+
+
+url = "https://storage.googleapis.com/download.tensorflow.org/data/abalone_raw.csv"
+out_path = "abalone_raw.csv"
+response = requests.get(url)
+with open(out_path, "wb") as fd:
+    fd.write(response.content)
+
+dataset_df = pd.read_csv("abalone_raw.csv")
+print(dataset_df.head(3))
+
+# Split the dataset into a training and testing dataset.
+train_ds_pd, test_ds_pd = split_dataset(dataset_df)
+print("{} examples in training, {} examples for testing.".format(
+    len(train_ds_pd), len(test_ds_pd)))
+
+# Name of the label column.
+label = "Rings"
+
+train_ds = tfdf.keras.pd_dataframe_to_tf_dataset(train_ds_pd, label=label, task=tfdf.keras.Task.REGRESSION)
+test_ds = tfdf.keras.pd_dataframe_to_tf_dataset(test_ds_pd, label=label, task=tfdf.keras.Task.REGRESSION)
+
+# Configure the model.
+model_7 = tfdf.keras.RandomForestModel(task = tfdf.keras.Task.REGRESSION)
+
+# Train the model.
+model_7.fit(x=train_ds)
+
+model_7.compile(metrics=["mse"])
+evaluation = model_7.evaluate(test_ds, return_dict=True)
+
+print(evaluation)
+print()
+print(f"MSE: {evaluation['mse']}")
+print(f"RMSE: {math.sqrt(evaluation['mse'])}")
+

### 2022-10-06 scripts/rf-kerastf.py
towards putting the rf model on gee

@@ -1,48 +1,21 @@
-# https://www.tensorflow.org/decision_forests/tutorials/beginner_colab#training_a_regression_model
+import sys
+import pickle
 
-import math
-import requests
-import numpy as np
-import pandas as pd
-import tensorflow_decision_forests as tfdf
+sys.path.append(".")
+from src import scikit_learn_model_converter as sci2tf
 
-def split_dataset(dataset, test_ratio=0.30):
-  """Splits a panda dataframe in two."""
-  test_indices = np.random.rand(len(dataset)) < test_ratio
-  return dataset[~test_indices], dataset[test_indices]
+variable = "salinity"
 
+path = "data/rf_random_" + variable + ".pkl"
+# path = "data/rfecv_" + variable + ".pkl"
+# path = "data/ols_" + variable + ".pkl"
 
-url = "https://storage.googleapis.com/download.tensorflow.org/data/abalone_raw.csv"
-out_path = "abalone_raw.csv"
-response = requests.get(url)
-with open(out_path, "wb") as fd:
-    fd.write(response.content)
+model = pickle.load(open(path, "rb"))
 
-dataset_df = pd.read_csv("abalone_raw.csv")
-print(dataset_df.head(3))
+test = model.best_estimator_
 
-# Split the dataset into a training and testing dataset.
-train_ds_pd, test_ds_pd = split_dataset(dataset_df)
-print("{} examples in training, {} examples for testing.".format(
-    len(train_ds_pd), len(test_ds_pd)))
+tensorflow_model = sci2tf.convert(test)
 
-# Name of the label column.
-label = "Rings"
-
-train_ds = tfdf.keras.pd_dataframe_to_tf_dataset(train_ds_pd, label=label, task=tfdf.keras.Task.REGRESSION)
-test_ds = tfdf.keras.pd_dataframe_to_tf_dataset(test_ds_pd, label=label, task=tfdf.keras.Task.REGRESSION)
-
-# Configure the model.
-model_7 = tfdf.keras.RandomForestModel(task = tfdf.keras.Task.REGRESSION)
-
-# Train the model.
-model_7.fit(x=train_ds)
-
-model_7.compile(metrics=["mse"])
-evaluation = model_7.evaluate(test_ds, return_dict=True)
-
-print(evaluation)
-print()
-print(f"MSE: {evaluation['mse']}")
-print(f"RMSE: {math.sqrt(evaluation['mse'])}")
-
+type(tensorflow_model)
[…]

### 2022-10-07 scripts/rf-kerastf.py
towards putting the rf model on gee

@@ -19,3 +19,53 @@ tensorflow_model = sci2tf.convert(test)
 type(tensorflow_model)
 dir(tensorflow_model)
 tensorflow_model.save("model", save_format="tf")
+
+
+# https://www.tensorflow.org/decision_forests/tutorials/beginner_colab#training_a_regression_model
+
+import math
+import requests
+import numpy as np
+import pandas as pd
+import tensorflow_decision_forests as tfdf
+
+def split_dataset(dataset, test_ratio=0.30):
+  """Splits a panda dataframe in two."""
+  test_indices = np.random.rand(len(dataset)) < test_ratio
+  return dataset[~test_indices], dataset[test_indices]
+
+
+# url = "https://storage.googleapis.com/download.tensorflow.org/data/abalone_raw.csv"
+# out_path = "abalone_raw.csv"
+# response = requests.get(url)
+# with open(out_path, "wb") as fd:
+#     fd.write(response.content)
+
+dataset_df = pd.read_csv("abalone_raw.csv")
+print(dataset_df.head(3))
+
+# Split the dataset into a training and testing dataset.
+train_ds_pd, test_ds_pd = split_dataset(dataset_df)
+print("{} examples in training, {} examples for testing.".format(
+    len(train_ds_pd), len(test_ds_pd)))
+
+# Name of the label column.
+label = "Rings"
+
+train_ds = tfdf.keras.pd_dataframe_to_tf_dataset(train_ds_pd, label=label, task=tfdf.keras.Task.REGRESSION)
+test_ds = tfdf.keras.pd_dataframe_to_tf_dataset(test_ds_pd, label=label, task=tfdf.keras.Task.REGRESSION)
+
+# Configure the model.
+model_7 = tfdf.keras.RandomForestModel(task = tfdf.keras.Task.REGRESSION)
+
+# Train the model.
+model_7.fit(x=train_ds)
+
+model_7.compile(metrics=["mse"])
+evaluation = model_7.evaluate(test_ds, return_dict=True)
+
+print(evaluation)
+print()
+print(f"MSE: {evaluation['mse']}")
+print(f"RMSE: {math.sqrt(evaluation['mse'])}")
+

### 2022-10-10 scripts/rf-kerastf.py
cleanup

@@ -29,8 +29,7 @@ import numpy as np
 import pandas as pd
 import tensorflow_decision_forests as tfdf
 
-def split_dataset(dataset, test_ratio=0.30):
-  """Splits a panda dataframe in two."""
+def split_dataset(dataset, test_ratio=0.30):  
   test_indices = np.random.rand(len(dataset)) < test_ratio
   return dataset[~test_indices], dataset[test_indices]
 

### 2022-10-10 scripts/gee.py
towards a gee-rf pipeline

@@ -1,0 +1,31 @@
+import pickle
+import numpy as np
+import pandas as pd
+# import tensorflow_decision_forests as tfdf
+
+
+def split_dataset(dataset, test_ratio=0.30):  
+  test_indices = np.random.rand(len(dataset)) < test_ratio
+  return dataset[~test_indices], dataset[test_indices]
+
+output_bucket = "rabpro-gee-uploads"
+model_dir = "gs://" + output_bucket + "/rf_model"
+
+variable = "temperature"
+X_train = pickle.load(open("data/X_train_" + variable + ".pkl", "rb"))
+X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
+y_train = pickle.load(open("data/y_train_" + variable + ".pkl", "rb"))
+y_test = pickle.load(open("data/y_test_" + variable + ".pkl", "rb"))
+important_params = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
+
+train_ds_pd = pd.DataFrame(X_train, columns=important_params)
+train_ds_pd[variable] = y_train
+
+train_ds = tfdf.keras.pd_dataframe_to_tf_dataset(train_ds_pd, label=variable, task=tfdf.keras.Task.REGRESSION)
+
+model = tfdf.keras.RandomForestModel(task=tfdf.keras.Task.REGRESSION)
+model.compile(metrics=["mse"])
+model.fit(x=train_ds)
+
+model.save("rf_model", save_format="tf")
+# model.save(model_dir, save_format="tf")

### 2022-10-10 scripts/gee.py
towards a gee-rf pipeline

@@ -1,7 +1,7 @@
 import pickle
 import numpy as np
 import pandas as pd
-# import tensorflow_decision_forests as tfdf
+import tensorflow_decision_forests as tfdf
 
 
 def split_dataset(dataset, test_ratio=0.30):  

### 2022-10-10 scripts/rf-kerastf.py
towards gee-rf pipeline

@@ -1,5 +1,6 @@
 import sys
 import pickle
+import tensorflow as tf
 
 sys.path.append(".")
 from src import scikit_learn_model_converter as sci2tf
@@ -11,10 +12,25 @@ path = "data/rf_random_" + variable + ".pkl"
 # path = "data/ols_" + variable + ".pkl"
 
 model = pickle.load(open(path, "rb"))
+X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
+y_test = pickle.load(open("data/y_test_" + variable + ".pkl", "rb"))
+X_train = pickle.load(open("data/X_train_" + variable + ".pkl", "rb"))
+y_train = pickle.load(open("data/y_train_" + variable + ".pkl", "rb"))
+imp_params = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"))
 
-test = model.best_estimator_
+test = pd.DataFrame(X_test, columns=imp_params)
+test["y"] = y_test
+test_ds = tfdf.keras.pd_dataframe_to_tf_dataset(test, label="y", task=tfdf.keras.Task.REGRESSION)
 
-tensorflow_model = sci2tf.convert(test)
+train = pd.DataFrame(X_train, columns=imp_params)
+train["y"] = y_train
+train_ds = tfdf.keras.pd_dataframe_to_tf_dataset(train, label="y", task=tfdf.keras.Task.REGRESSION)
+
+tensorflow_model = sci2tf.convert(model.best_estimator_)
+
+tensorflow_model.compile()
+tensorflow_model.fit(x=train_ds)
+tensorflow_model.evaluate(test_ds)
 
 type(tensorflow_model)
 dir(tensorflow_model)

### 2022-10-10 scripts/gee.py
towards gee-rf pipeline

@@ -1,8 +1,11 @@
+import math
 import pickle
 import numpy as np
 import pandas as pd
+import tensorflow as tf
 import tensorflow_decision_forests as tfdf
 
+print(tf.__version__)
 
 def split_dataset(dataset, test_ratio=0.30):  
   test_indices = np.random.rand(len(dataset)) < test_ratio
@@ -20,12 +23,22 @@ important_params = pickle.load(open("data/imp_params_" + variable + ".pkl", "rb"
 
 train_ds_pd = pd.DataFrame(X_train, columns=important_params)
 train_ds_pd[variable] = y_train
-
 train_ds = tfdf.keras.pd_dataframe_to_tf_dataset(train_ds_pd, label=variable, task=tfdf.keras.Task.REGRESSION)
 
+test_ds_pd = pd.DataFrame(X_test, columns=important_params)
+test_ds_pd[variable] = y_test
+test_ds = tfdf.keras.pd_dataframe_to_tf_dataset(test_ds_pd, label=variable, task=tfdf.keras.Task.REGRESSION)
+
 model = tfdf.keras.RandomForestModel(task=tfdf.keras.Task.REGRESSION)
-model.compile(metrics=["mse"])
 model.fit(x=train_ds)
+model.compile(metrics=["mse"])
+
+evaluation = model.evaluate(test_ds, return_dict=True)
+
+print(evaluation)
+print()
+print(f"MSE: {evaluation['mse']}")
+print(f"RMSE: {math.sqrt(evaluation['mse'])}")
 
 model.save("rf_model", save_format="tf")
 # model.save(model_dir, save_format="tf")

### 2022-10-10 scripts/rf-kerastf.py

@@ -74,6 +74,7 @@ test_ds = tfdf.keras.pd_dataframe_to_tf_dataset(test_ds_pd, label=label, task=tf
 model_7 = tfdf.keras.RandomForestModel(task = tfdf.keras.Task.REGRESSION)
 
 # Train the model.
+tf.autograph.set_verbosity(3, True)
 model_7.fit(x=train_ds)
 
 model_7.compile(metrics=["mse"])

### 2022-10-10 scripts/03_get_data_predict.py
debugging

@@ -16,6 +16,7 @@ def get_modis(date):
 
     obs = ee.FeatureCollection(asset)
     ic = ee.ImageCollection(dataset)
+    breakpoint()
     ic = ic.filterDate("2018-01-01", "2018-01-02")
 
     def getBandValues(im):

### 2022-10-10 scripts/03_get_data_predict.py
complete predict pipeline w/o hardcoded date

@@ -3,7 +3,8 @@
 import ee
 import requests
 import argparse
-import pandas as pd
+from datetime import datetime
+from datetime import timedelta
 
 ee.Initialize()
 
@@ -16,8 +17,8 @@ def get_modis(date):
 
     obs = ee.FeatureCollection(asset)
     ic = ee.ImageCollection(dataset)
-    breakpoint()
-    ic = ic.filterDate("2018-01-01", "2018-01-02")
+    date_plus_one = str(datetime.strptime(date, "%Y-%m-%d") + timedelta(days=1)).split(" ")[0]
+    ic = ic.filterDate(date, date_plus_one)
 
     def getBandValues(im):
         bandValFC = ee.Algorithms.If(

### 2022-10-10 scripts/gee-predict.py
towards a gee app

@@ -1,0 +1,15 @@
+import math
+import pickle
+import numpy as np
+import pandas as pd
+import tensorflow as tf
+import tensorflow_decision_forests as tfdf
+
+print(tf.__version__)
+
+date = "2022-09-04"
+
+dataset = 'MODIS/006/MYDOCGA'
+asset_pixelcenters = "users/jstacompute/icom_pixelcenters_4326"
+
+# load static asset

### 2022-10-10 scripts/gee-app.js
towards gee app

@@ -1,0 +1,55 @@
+
+
+var theMap = ui.Map();
+ui.root.clear();
+
+colDatesPanel = function () {
+    var startDayLabel = ui.Label('Date:');
+    var startDayBox = ui.Textbox({ value: '2022-06-10' });
+    startDayBox.style().set('stretch', 'horizontal');
+
+    return ui.Panel(
+        [
+            ui.Panel(
+                [startDayLabel, startDayBox],
+                ui.Panel.Layout.Flow('horizontal'), { stretch: 'horizontal' }
+            )
+        ]
+    );
+};
+visSelectPanel = function () {
+    var indexList = ['Salinity', 'Temperature', 'Turbidity'];
+    var indexSelect = ui.Select({ items: indexList, value: 'Salinity', style: { stretch: 'horizontal' } });
+    return ui.Panel([indexSelect], null, { stretch: 'horizontal' });
+};
+
+var region = ee.Geometry.Rectangle(-77.576, 36.661, -75.605, 39.632);
+
+var startDate = '2018-01-01'
+var endDate = '2018-01-02'
+
+var collection = ee.ImageCollection('MODIS/006/MYDOCGA')
+    .filterDate(startDate, endDate)
+    .map(function (image) { return image.clip(region) });
+
+theMap.centerObject(region, 8)
+theMap.addLayer(collection)
+
+var controlPanel = ui.Panel({
+    style: { width: '350px', position: 'top-left' } //, backgroundColor: 'rgba(255, 255, 255, 0)'
+});
+
+var colDatesPanel = colDatesPanel();
+var visSelectPanel = visSelectPanel();
+
+var url = ui.Label({
+    value: 'About geowq',
+});
+url.setUrl('https://github.com/DOE-ICoM/geowq');
+
+controlPanel.add(colDatesPanel);
+controlPanel.add(visSelectPanel);
+controlPanel.add(url);
+
+ui.root.add(controlPanel);
+ui.root.add(theMap);

### 2022-10-11 scripts/gee-train.py
towards gee app

@@ -11,9 +11,6 @@ def split_dataset(dataset, test_ratio=0.30):
   test_indices = np.random.rand(len(dataset)) < test_ratio
   return dataset[~test_indices], dataset[test_indices]
 
-output_bucket = "rabpro-gee-uploads"
-model_dir = "gs://" + output_bucket + "/rf_model"
-
 variable = "temperature"
 X_train = pickle.load(open("data/X_train_" + variable + ".pkl", "rb"))
 X_test = pickle.load(open("data/X_test_" + variable + ".pkl", "rb"))
@@ -41,4 +38,3 @@ print(f"MSE: {evaluation['mse']}")
 print(f"RMSE: {math.sqrt(evaluation['mse'])}")
 
 model.save("rf_model", save_format="tf")
-# model.save(model_dir, save_format="tf")

### 2022-10-14 manuscript/manuscript.tex
rm watermark
numbered and order citations

@@
 \modulolinenumbers[2]
-\usepackage[authoryear]{natbib}
-\usepackage{draftwatermark}
-\SetWatermarkScale{5}
+% \usepackage[authoryear]{natbib}
+\usepackage[style=numeric,sorting=none]{biblatex}
+\addbibresource{manuscript/geowq.bib}
+% \usepackage{draftwatermark}
+% \SetWatermarkScale{5}
 \usepackage{float}
@@
 
-In partially mixed estuaries such as the Chesapeake Bay (hereafter, “the Bay”), productivity, organic matter production, and phytoplankton community composition are largely determined by stratification \citep{xuClimateForcingSalinity2012}.
+In partially mixed estuaries such as the Chesapeake Bay (hereafter, “the Bay”), productivity, organic matter production, and phytoplankton community composition are largely determined by stratification \cite{xuClimateForcingSalinity2012}.
 Because stratification is controlled by salinity and temperature, extensive measurement networks exist to monitor these water properties (Chesapeake Bay Program Water Quality Database, \url{https://www.chesapeakebay.net/what/downloads/cbp_water_quality_database_1984_present}).
-These data are critical for establishing point-based trends and relationships between river flow or sea level rise and salinity or hypoxia \citet{hagyHypoxiaChesapeakeBay2004}.
-Of particular importance is monitoring for increased nutrient loads as result of expected increases in precipitation, tributary discharge, and river flows as a result of climate change \citep{najjarPotentialClimatechangeImpacts2010, irbyCompetingImpactsClimate2018}.
+These data are critical for establishing point-based trends and relationships between river flow or sea level rise and salinity or hypoxia \cite{hagyHypoxiaChesapeakeBay2004}.
+Of particular importance is monitoring for increased nutrient loads as result of expected increases in precipitation, tributary discharge, and river flows as a result of climate change \cite{najjarPotentialClimatechangeImpacts2010, irbyCompetingImpactsClimate2018}.
 
-Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \citep{hoodChesapeakeBayProgram2021}.
+Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \cite{hoodChesapeakeBayProgram2021}.
 For example, habitat suitability assessments are typically done on a gridded basis for the entirety of a particular estuary rather than at selected point locations.
-Applications such as these often turn to gridded alternatives to direct observations such as output from directly-observing satellite-based sensors \citep{foreCombinedActivePassive2016} or operational now/fore-casting products \citep{lanerolle2011second}.
-Unfortunately, these too are of limited use for applications such as validation of next-generation Earth Systems Models (ESMs, \citet{golazDOEE3SMModel2022}), where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
+Applications such as these often turn to gridded alternatives to direct observations such as output from directly-observing satellite-based sensors \cite{foreCombinedActivePassive2016} or operational now/fore-casting products \cite{lanerolle2011second}.
+Unfortunately, these too are of limited use for applications such as validation of next-generation Earth Systems Models (ESMs, \cite{golazDOEE3SMModel2022}), where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
 
@@
 The primary input to our model is remotely sensed reflectance data.
-In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \citep{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}.
+In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \cite{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}.
 We supplemented MODIS data with gridded estimates of distance-weighted tributary inflows which we used to encourage physical realism (i.e.
@@
 Our model represents a further advantage over prior efforts at data-driven modeling of salinity in the Bay (e.g.
-\citet{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
+\cite{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
 These more spatially extenstive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
@@
 Chesapeake Bay is located in the Mid-Atlantic region of the United States and serves as the outlet of 9 major tributary systems.
-Of these, the Susquehanna River (at the Northern extent of the Bay) provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \citet{xuClimateForcingSalinity2012}).
-The Bay is fairly shallow ($<$ 30m, \citet{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \citep{xuClimateForcingSalinity2012}.
+Of these, the Susquehanna River (at the Northern extent of the Bay) provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \cite{xuClimateForcingSalinity2012}).
+The Bay is fairly shallow ($<$ 30m, \cite{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \cite{xuClimateForcingSalinity2012}.
 For the present study, this means that surface water properties are likely to differ substantially from water properties at deeper depths.
@@
 Prior data-driven models (e.g.
-\citet{urquhartGeospatialInterpolationMODISderived2013, vogelAssessingSatelliteSea2016}) of Bay water properties used Generalized Linear Models (GLM), Neural Networks (NN), Generalized Additive Models (GAMs), and Random Forests (RF).
-We chose to use the Random Forest algorithm in part because \citet{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) and also because of its ease of implementation and interpretation.
-We fit RF models using the scikit-learn Python package \citep{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model (Figure 1).
+\cite{urquhartGeospatialInterpolationMODISderived2013, vogelAssessingSatelliteSea2016}) of Bay water properties used Generalized Linear Models (GLM), Neural Networks (NN), Generalized Additive Models (GAMs), and Random Forests (RF).
+We chose to use the Random Forest algorithm in part because \cite{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) and also because of its ease of implementation and interpretation.
+We fit RF models using the scikit-learn Python package \cite{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model (Figure 1).
 We subsequently “tuned” each model using k-fold cross validation to optimize the size, depth, and splitting criteria of each tree.
-We evaluated the “importance” of particular variables as the average impurity decrease of tree splits based on that variable \citep{pedregosaScikitlearnMachineLearning2011}.
-
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/data-processing_model-architecture}
-          \caption{Diagram of data processing, model fitting, tuning, and prediction pipeline.}
-    \end{center}    
-\end{figure}
+We evaluated the “importance” of particular variables as the average impurity decrease of tree splits based on that variable \cite{pedregosaScikitlearnMachineLearning2011}.
 
@@
 
-Because Chesapeake Bay has a mean optical depth of approximately 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column.
+Because Chesapeake Bay has a mean optical depth of approximately 0.89m \cite{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column.
 We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours GMT).
-Prior to model fitting, we removed the annual cycle from temperature observations using least squares optimization routines \citep{virtanenSciPyFundamentalAlgorithms2020a}.
+Prior to model fitting, we removed the annual cycle from temperature observations using least squares optimization routines \cite{virtanenSciPyFundamentalAlgorithms2020a}.
 This annual cycle was added back to model estimates for evaluation purposes and in the production of prediction surfaces.
@@
 We supplemented observational data with a grid of “freshwater influence” (FWI) estimates where each grid cell represents the average of annual-flow weighted distance of that cell from the 5 major tributaries (Susquehanna, James, Pautexent, Choptank, and Potomac).
-Annual flow magnitudes were calculated from National Water Information System instantaneous data (sites: 01488110, 01491000, 01576000, 01654000, 01594440, 02037500, \citet{nwis2021usgs}).
-Rather than straight-line distances we used “in-water” distances calculated using the scikit-image Python package \citep{littleKrigingEstuariesCrow1997,vanderwaltScikitimageImageProcessing2014}.
+Annual flow magnitudes were calculated from National Water Information System instantaneous data (sites: 01488110, 01491000, 01576000, 01654000, 01594440, 02037500, \cite{nwis2021usgs}).
+Rather than straight-line distances we used “in-water” distances calculated using the scikit-image Python package \cite{littleKrigingEstuariesCrow1997,vanderwaltScikitimageImageProcessing2014}.
 Our in-water distance calculations involved creating a cost–surface whereby land pixels had a very high cost relative to water pixels.
 Then the distance from any one cell to a tributary source was the accumulated travel “cost” as given by the scikit-image \texttt{route\_through\_array} function.
-We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands of values for each pixel spanning from 405 to 877nm wavelength \citep{vermoteericMOD09GAMODISTerra2015}.
+We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands of values for each pixel spanning from 405 to 877nm wavelength \cite{vermoteericMOD09GAMODISTerra2015}.
 
-All code for data processing, model fitting, and model evaluation is available at Stachelek et al. (2023). All data used in the study are available in Stachelek et al. (2023).
+All code for data processing, model fitting, and model evaluation is available at Stachelek et al.
+(2023) the pipeline for which is shown in Figure 1.
+All data used in the study are available in Stachelek et al.
+(2023).
+
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/data-processing_model-architecture}
+          \caption{Diagram of the data processing, model fitting, tuning, and prediction pipeline.}
+    \end{center}    
+\end{figure}
 
@@
 
-We validated RF model predictions both quantitatively against a hold-out test set of observational data (one third of the total data) and qualitatively against data from the Chesapeake Bay Operational Forecast System (CBOFS, \citet{lanerolle2011second}).
+We validated RF model predictions both quantitatively against a hold-out test set of observational data (one third of the total data) and qualitatively against data from the Chesapeake Bay Operational Forecast System (CBOFS, \cite{lanerolle2011second}).
 CBOFS provides data on Bay-wide water level, temperature, and salinity every 6 hours.
-The model itself uses National Oceanic and Atmospheric Administration (NOAA) water level, salinity observations, and US Geological Survey (USGS) river discharge data to calibrate an implementation of the Rutgers University Regional Ocean Modeling System (ROMS, \citet{shchepetkinRegionalOceanicModeling2005}).
+The model itself uses National Oceanic and Atmospheric Administration (NOAA) water level, salinity observations, and US Geological Survey (USGS) river discharge data to calibrate an implementation of the Rutgers University Regional Ocean Modeling System (ROMS, \cite{shchepetkinRegionalOceanicModeling2005}).
 
@@
 RF model predictions for salinity were generally lower than corresponding values from the CBOFS in the mainstem of the Bay and higher than corresponding CBOFS values in the tributaries (Figure 5).
-CBOFS surface water salinity estimates are known to be saltier than observations \citep{lanerolle2011second, vogelAssessingSatelliteSea2016} partly explaining our saltier data-driven results.
+CBOFS surface water salinity estimates are known to be saltier than observations \cite{lanerolle2011second, vogelAssessingSatelliteSea2016} partly explaining our saltier data-driven results.
 The salinity model was able to reproduce realistic temporal dynamics whereby the Bay is generally “fresher” in the Spring season (April-June) and “saltier” in Fall and Winter seasons (Figure 4).
@@
 Our geographically aware approach was able to accurately reproduce realistic spatial and temporal patterns of Bay salinity and temperature.
-Model errors were generally lower than prior data-driven efforts and as good or better than physics-based models \citep{vogelAssessingSatelliteSea2016}.
+Model errors were generally lower than prior data-driven efforts and as good or better than physics-based models \cite{vogelAssessingSatelliteSea2016}.
 In addition to improved accuracy, our approach provides the ability to generate rapid bay-wide prediction surfaces in seconds and it is easily updatable insofar as saved models can be updated as new data become available.
@@
 As a result, they cannot be reliably compared across different instruments.
-A better turbidity model would likely result from using total suspended matter (TSM) data measured in a standardized unit such as mg/l \citep{ondrusekDevelopmentNewOptical2012}.
+A better turbidity model would likely result from using total suspended matter (TSM) data measured in a standardized unit such as mg/l \cite{ondrusekDevelopmentNewOptical2012}.
 Another reason the turbidity model performed poorly may be that turbidity is simply difficult to predict.
@@
 Temperature and salinity models produced realistic predictions insofar as they were characterized by low salinity in the upper tributaries and high minimum water temperatures at the Bay mouth.
-This indicates that the model captures both freshwater discharge from tributaries as well as the moderating influence of ocean water in areas near the Bay mouth \citep{dingSpatiotemporalPatternsWater2015}.
+This indicates that the model captures both freshwater discharge from tributaries as well as the moderating influence of ocean water in areas near the Bay mouth \cite{dingSpatiotemporalPatternsWater2015}.
 
-Our comprehensive dataset includes greater spatial coverage than prior efforts that were  largely trained on limited (and relatively high-salinity) buoy or in-situ cruise data located in the mainstem of the Bay \citep{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013, ondrusekDevelopmentNewOptical2012}.
+Our comprehensive dataset includes greater spatial coverage than prior efforts that were  largely trained on limited (and relatively high-salinity) buoy or in-situ cruise data located in the mainstem of the Bay \cite{vogelAssessingSatelliteSea2016, geigerSatellitederivedCoastalOcean2013, ondrusekDevelopmentNewOptical2012}.
 This coverage allows for a more comprehensive assessment of the spatial uncertainty of our model.
@@
 This appears to be the collision point of the marine plume emanating from the Bay mouth with the discharge of the Potomac River.
-The hydrodynamics here seem to be complex given that the CBOFS has difficulty resolving spatial patterning in this area \citep{lanerolle2011second}.
+The hydrodynamics here seem to be complex given that the CBOFS has difficulty resolving spatial patterning in this area \cite{lanerolle2011second}.
 This was an area where we found limited observational data (Figure S2).
@@
 For the temperature model, we were unable to recover low temperature observations less than 0 deg C.
-These results are consistent with prior research showing winter had the highest seasonal error  \citet{geigerSatellitederivedCoastalOcean2013}.
-One explanation for these high winter error rates is the fact that winter in the Bay is generally characterized by strong vertical mixing due to high winds \citep{sonWaterPropertiesChesapeake2012, xuClimateForcingSalinity2012} making “true” patterns more dynamic and more difficult to resolve.
+These results are consistent with prior research showing winter had the highest seasonal error  \cite{geigerSatellitederivedCoastalOcean2013}.
+One explanation for these high winter error rates is the fact that winter in the Bay is generally characterized by strong vertical mixing due to high winds \cite{sonWaterPropertiesChesapeake2012, xuClimateForcingSalinity2012} making “true” patterns more dynamic and more difficult to resolve.
 
@@
 A key difference between our approach and that of CNNs is that our approach essentially fits each observation in isolation and does not exploit the “spatial locality” (i.e.
-the “neighborhood”) surrounding observations as is the case with CNNs \citep{goodfellow2016convolutional}.
+the “neighborhood”) surrounding observations as is the case with CNNs \cite{goodfellow2016convolutional}.
 The pipeline for training a CNN would bring some practical difficulties regarding the collection and storage of MODIS data.
@@
 
-\bibliographystyle{apalike}
-\bibliography{manuscript/geowq}
+% \bibliographystyle{apalike}
+% \bibliography{manuscript/geowq}
+\printbibliography
 

### 2022-11-16 figures/validation.py
stash

@@ -1,5 +1,4 @@
 import sys
-import math
 import pickle
 import numpy as np
 import pandas as pd
@@ -49,6 +48,12 @@ variables_str_short = [utils.clean_var_name(v) for v in variables_str]
 # scatter plot of measured vs predicted
 res = [_get_predictions(variable) for variable in variables_str_short]
 
+# scatter plot of turbidity versus salinity
+# variables_str_short
+# test = pd.concat([res[0]["obs"], np.exp(res[2]["obs"])], axis=1)
+# test.columns = ["salinity", "turbidity"]
+# sns.scatterplot(x="salinity", y="turbidity", data=test)
+# plt.show()
 
 def _plot(ax,
           dt,

### 2022-11-16 manuscript/manuscript.tex
add keywords
text edits

@@
 \usepackage{lipsum}
+\usepackage{soul}
+\usepackage{color}
 \usepackage[margin=1in]{geometry}
@@
 
+Keywords: reflectance, remote sensing, salinity, temperature, turbidity, Chesapeake
+
 \linenumbers
@@
 
-Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as SLR forecasting and habitat suitability assessments \cite{hoodChesapeakeBayProgram2021}.
+Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as \hl{Sea Level Rise (SLR)} forecasting and habitat suitability assessments \cite{hoodChesapeakeBayProgram2021}.
 For example, habitat suitability assessments are typically done on a gridded basis for the entirety of a particular estuary rather than at selected point locations.
@@
 To address this validation data gap, we developed a data-driven model to produce high-resolution (< 1 $km^2$) estimates of temperature, salinity, and turbidity at daily time-steps (over decadal time scales) as required by next-generation ESMs.
-The primary input to our model is remotely sensed reflectance data.
+The primary input to our model is \hl{remote sensing} reflectance data.
 In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \cite{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}.
@@
 
-We have demonstrated an approach to produce realistic estimates of water properties in Chesapeake Bay using remotely sensed data.
+We have demonstrated an approach to produce realistic estimates of water properties in Chesapeake Bay using \hl{remote sensing} data.
 Salinity and temperature model performance was as good or better than prior efforts.
@@
 
+\section*{Biographies}
+
+\begin{description}
+    \item \textbf{Jemma Stachelek} is a Limnologist and Aquatic Earth Scientist (PhD, Msc, BSc) in the Division of Earth and Environmental Sciences at Los Alamos National Laboratory.
+Her research focuses on the applications of spatially intensive geocomputation to water and food security problems particularly those related to eutrophication and nutrient pollution.
+    \item \textbf{Sofia Avendaño} is awesome.
+    \item \textbf{Jon Schwenk} so is Jon.
+\end{description}
+
+\clearpage
+
+\section*{Figure Captions}
+
+\begin{description}
+    \item Figure 1: Diagram of the data processing, model fitting, tuning, and prediction pipeline.
+    \item Figure 2: Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.
+    \item Figure 3: Random Forest importance plot for salinity, temperature, and turbidity models.
+    \item Figure 4: Seasonal Random Forest prediction results for temperature and salinity.
+    \item Figure 5: Comparison between Random Forest salinity prediction results and a CBOFS snapshot for Sept, 4, 2022.
+\end{description}
+
 \end{document}

### 2022-11-16 manuscript/manuscript.tex
text edits

@@
 We fit RF models using the scikit-learn Python package \cite{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model (Figure 1).
-We subsequently “tuned” each model using k-fold cross validation to optimize the size, depth, and splitting criteria of each tree.
+\hl{We chose initial forest size and maximum depth values to optimize model performance and model training time respectively.} We subsequently “tuned” each model using k-fold cross validation to optimize the size, depth, and splitting criteria of each tree.
 We evaluated the “importance” of particular variables as the average impurity decrease of tree splits based on that variable \cite{pedregosaScikitlearnMachineLearning2011}.
@@
 Then the distance from any one cell to a tributary source was the accumulated travel “cost” as given by the scikit-image \texttt{route\_through\_array} function.
-We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 500m resolution MODIS Aqua product, which provides 8 bands of values for each pixel spanning from 405 to 877nm wavelength \cite{vermoteericMOD09GAMODISTerra2015}.
+We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the \hl{1000m} resolution MODIS Aqua product, which provides \hl{9} bands of values for each pixel spanning from \hl{the following wavelength ranges (405-420, 438-448, 483-493, 526-536, 546-556, 662-672, 673-683, 743-753, and 862-877nm)} \cite{vermoteericMOD09GAMODISTerra2015}.
 
@@
           \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_importance_all}
-          \caption{Random Forest importance plot for salinity, temperature, and turbidity models.}
+          \caption{Random Forest importance plot for salinity, temperature, and turbidity models.
+Note that all features may not be present in all models if they were dropped as a result of recursive feature elimination.}
     \end{center}    
@@
     \item \textbf{Sofia Avendaño} is awesome.
-    \item \textbf{Jon Schwenk} so is Jon.
+    \item \textbf{Jon Schwenk} is a River and Data Scientist (PhD) in the Division of Earth and Environmental Sciences at Los Alamos National Laboratory.
+One of his primary research themes is the use of the unprecedented wealth of water related data to drive models that address pressing water security issues.
 \end{description}
@@
     \item Figure 3: Random Forest importance plot for salinity, temperature, and turbidity models.
+\hl{Note that all features may not be present in all models if they were dropped as a result of recursive feature elimination.}
     \item Figure 4: Seasonal Random Forest prediction results for temperature and salinity.

### 2022-11-16 manuscript/manuscript.tex
text edits

@@
 \usepackage{lipsum}
+\usepackage{gensymb}
 \usepackage{soul}
@@
 
-Chesapeake Bay is located in the Mid-Atlantic region of the United States and serves as the outlet of 9 major tributary systems.
+Chesapeake Bay is located in the Mid-Atlantic region of the United States \hl{(at approximately 37.8\degree N 76.1\degree W)} and serves as the outlet of 9 major tributary systems.
 Of these, the Susquehanna River (at the Northern extent of the Bay) provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \cite{xuClimateForcingSalinity2012}).
-The Bay is fairly shallow ($<$ 30m, \cite{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \cite{xuClimateForcingSalinity2012}.
+\hl{The Bay is the largest estuary in the United States and has an area of approximately 11,000 $km^2$.} The Bay is also fairly shallow ($<$ 30m, \cite{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \cite{xuClimateForcingSalinity2012}.
 For the present study, this means that surface water properties are likely to differ substantially from water properties at deeper depths.
@@
 We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the \hl{1000m} resolution MODIS Aqua product, which provides \hl{9} bands of values for each pixel spanning from \hl{the following wavelength ranges (405-420, 438-448, 483-493, 526-536, 546-556, 662-672, 673-683, 743-753, and 862-877nm)} \cite{vermoteericMOD09GAMODISTerra2015}.
+\hl{MODIS data was pulled from Google Earth Engine (ID:MODIS/006/MYDOCGA) using the \texttt{earthengine-api} Python package.}
 
 All code for data processing, model fitting, and model evaluation is available at Stachelek et al.
-(2023) the pipeline for which is shown in Figure 1.
+(2023a) the pipeline for which is shown in Figure 1.
 All data used in the study are available in Stachelek et al.
-(2023).
+(2023b).
 

### 2022-11-16 scripts/02_rf_tune.py
back transform turb rmse

@@ -2,6 +2,7 @@ import sys
 import pickle
 import warnings
 import argparse
+import numpy as np
 from tabulate import tabulate
 
 sys.path.append(".")
@@ -37,6 +38,8 @@ def main():
 
     print('Final RMSE:')
     print(rmse)
+    if variable == "turbidity":
+        rmse = np.exp(rmse)
     with open("data/rmse_" + variable + ".md", 'w') as f:
         f.write(tabulate([[rmse]], headers=["rmse"]))
 

### 2022-11-16 figures/validation.py
fix misordered plotting and r2 calcs

@@ -127,6 +127,13 @@ plt.close()
 r2 = []
 fig, axes = plt.subplots(figsize=(8, 4.5), ncols=3)
 r2.append(_plot(axes[0], res[0], "salinity", -0.5, 32, ylab=True, text_space=2.5)[1])
+r2.append(_plot(axes[1],
+      res[1],
+      "temperature",
+      xy_min=-2,
+      xy_max=35,
+      text_anchor=(3, 30),
+      text_space=2.5)[1])
 r2.append(_plot(axes[2],
       res[2],
       "turbidity",
@@ -136,13 +143,6 @@ r2.append(_plot(axes[2],
       text_anchor=(37, 10),
       text_space=5,
       log=True)[1])
-r2.append(_plot(axes[1],
-      res[1],
-      "temperature",
-      xy_min=-2,
-      xy_max=35,
-      text_anchor=(3, 30),
-      text_space=2.5)[1])
 # plt.show()
 
 r2_df = pd.DataFrame(r2, columns=["r2"])

### 2022-11-16 manuscript/manuscript.tex
text edits, spellcheck, inline results sync

@@
 Our model fits daily MODIS Aqua reflectance data to surface observations ($<$ 1 m depth) from 2000-2021 in Chesapeake Bay, USA.
-The resulting models have similar error statistics as prior efforts of this type for salinity (RMSE: 2.2) and temperature (RMSE: 1.9 C).
+The resulting models have similar error statistics as prior efforts of this type for salinity (RMSE: 2.3) and temperature (RMSE: 1.8 C).
 However, unlike prior efforts our model is designed as a pipeline meaning that it has the advantage of producing predictions of water properties in future time periods as additional MODIS data becomes available.
@@
 \cite{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
-These more spatially extenstive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
+These more spatially extensive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
 
@@
 Of these, the Susquehanna River (at the Northern extent of the Bay) provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \cite{xuClimateForcingSalinity2012}).
-\hl{The Bay is the largest estuary in the United States and has an area of approximately 11,000 $km^2$.} The Bay is also fairly shallow ($<$ 30m, \cite{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \cite{xuClimateForcingSalinity2012}.
+\hl{The Bay is the largest estuary in the United States and has an area of approximately 11,000 $km^2$.
+As such, it is provides a wealth of natural economic benefits in the form of fisheries harvests (particularly oysters), tourism, and nature-based recreation} \cite{hoodChesapeakeBayProgram2021}.
+Despite its large size, the Bay is fairly shallow ($<$ 30m, \cite{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \cite{xuClimateForcingSalinity2012}.
 For the present study, this means that surface water properties are likely to differ substantially from water properties at deeper depths.
@@
 \begin{sloppypar}
-We compiled over two decades of observational data (2000-2021) on salinity (n=6k), temperature (n=10k), and turbidity (n=5k, Table S1).
+We compiled over two decades of observational data (2000-2021) on salinity (n=6k), temperature (n=10k, in units of \degree C), and turbidity (n=5k, in units of NTU, See Table S1).
 Although this data came from a wide variety of sources (n=20, Table S1), almost $80\%$ of the total unique observations came from three programs NDBC/CBIBS (\url{https://buoybay.noaa.gov/}), Maryland Continuous Monitoring Program (MDNR), and the Virginia Estuarine and Coastal Observing System (\url{http://vecos.vims.edu}).
@@
 
-Our tuned RF models for salinity, turbidity, and temperature had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 32.2 ($R^2$=0.07), and 1.8 ($R^2$=0.96) respectively (Table 1, Figure 2).
+Our tuned RF models for salinity, temperature, and turbidity had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 1.76 ($R^2$=0.96), and 1.82 ($R^2$=0.11) respectively (Table 1, Figure 2).
 Out-of-sample validation performance was good for salinity and temperature models but poor for the turbidity model (Figure 2, Figure S1).
@@
     \item \textbf{Jon Schwenk} is a River and Data Scientist (PhD) in the Division of Earth and Environmental Sciences at Los Alamos National Laboratory.
-One of his primary research themes is the use of the unprecedented wealth of water related data to drive models that address pressing water security issues.
+One of his primary research themes is the use of "big data" as it relates to water to drive models addressing pressing water security issues.
 \end{description}

### 2022-11-16 figures/maps.py
add coord ticks

@@ -9,6 +9,7 @@ import cartopy.crs as ccrs
 import matplotlib.pyplot as plt
 from matplotlib import gridspec
 import h3pandas  # h3.geo_to_h3_aggregate
+from cartopy.mpl.ticker import LongitudeFormatter, LatitudeFormatter
 
 sys.path.append(".")
 from src import utils
@@ -23,26 +24,41 @@ def panel_add(i,
               vmax=27,
               vmin=0,
               ticks=None,
-              height_frac=0.5):
+              height_frac=0.5,
+              lon_ticks=[],
+              lat_ticks=[],
+              shrink=0.45):
     if j is not None:
         ax = plt.subplot(axs[i, j], xlabel="", projection=ccrs.PlateCarree())
         ax.coastlines(resolution="10m", color="black", linewidth=1)
     else:
         ax = axs[i]
         ax.coastlines(resolution="10m", color="black", linewidth=1)
+
+    if len(lon_ticks) > 0 and len(lat_ticks) > 0:
+        ax.set_xticks(lon_ticks, crs=ccrs.PlateCarree())
+        ax.set_yticks(lat_ticks, crs=ccrs.PlateCarree())
+        ax.set_xticklabels([])
+        ax.set_yticklabels([])
+        lon_formatter = LongitudeFormatter(zero_direction_label=True)
+        lat_formatter = LatitudeFormatter()
+        ax.xaxis.set_major_formatter(lon_formatter)
+        ax.yaxis.set_major_formatter(lat_formatter)
+
     if diff:
         geo_grid.plot.imshow(ax=ax,
                              center=0,
                              cbar_kwargs={
-                                 "shrink": 0.5,
+                                 "shrink": shrink,
                                  'label': ''
                              })
     else:
         geo_grid.plot.imshow(ax=ax,
                              vmin=vmin,
                              vmax=vmax,
+                             add_labels=False,
                              cbar_kwargs={
-                                 "shrink": 0.5,
+                                 "shrink": shrink,
                                  'label': '',
                                  'ticks': ticks,
                              })  # , vmax=np.nanmax(img_cbofs.to_numpy()))
@@ -161,9 +177,7 @@ fig, axs = plt.subplots(
         "xlabel": ""
     },
 )
-
[…]

### 2022-11-16 figures/maps.py
formatting with yapf

@@ -177,15 +177,25 @@ fig, axs = plt.subplots(
         "xlabel": ""
     },
 )
-panel_add(0, axs, "RF Results", img_rf, height_frac=0.65, lat_ticks=[37, 38, 39], lon_ticks=[-77, -76, -75])
+panel_add(0,
+          axs,
+          "RF Results",
+          img_rf,
+          height_frac=0.65,
+          lat_ticks=[37, 38, 39],
+          lon_ticks=[-77, -76, -75])
 panel_add(1, axs, "CBOFS Snapshot", img_cbofs)
 panel_add(2, axs, "RF-CBOFS", test, diff=True, height_frac=0.7)
 
 # plt.show()
 plt.savefig("figures/_rf-vs-cbofs.pdf")
 
+
 # --- seasonality maps
-def _seasonality_map(variable, vmin=(0,0,0,0), vmax=(27,27,27,27), ticks=None):
+def _seasonality_map(variable,
+                     vmin=(0, 0, 0, 0),
+                     vmax=(27, 27, 27, 27),
+                     ticks=None):
     # variable  = "turbidity"
     # vmin=(3,3,21,21)
     # vmax=(12, 12, 29, 29)
@@ -205,10 +215,38 @@ def _seasonality_map(variable, vmin=(0,0,0,0), vmax=(27,27,27,27), ticks=None):
                             left=0.1666,
                             right=0.7)  # 0.83333
 
-    panel_add(0, axs, dates[0], imgs[0], j=0, vmin=vmin[0], vmax=vmax[0], ticks=ticks)
-    panel_add(0, axs, dates[1], imgs[1], j=1, vmin=vmin[1], vmax=vmax[1], ticks=ticks)
-    panel_add(1, axs, dates[2], imgs[2], j=0, vmin=vmin[2], vmax=vmax[2], ticks=ticks)
-    panel_add(1, axs, dates[3], imgs[3], j=1, vmin=vmin[3], vmax=vmax[3], ticks=ticks)
+    panel_add(0,
+              axs,
+              dates[0],
+              imgs[0],
+              j=0,
+              vmin=vmin[0],
+              vmax=vmax[0],
+              ticks=ticks)
+    panel_add(0,
+              axs,
+              dates[1],
+              imgs[1],
+              j=1,
+              vmin=vmin[1],
+              vmax=vmax[1],
+              ticks=ticks)
+    panel_add(1,
+              axs,
+              dates[2],
[…]

### 2022-11-17 manuscript/manuscript.tex
include study site fig in supplement

@@
 
-Chesapeake Bay is located in the Mid-Atlantic region of the United States \hl{(at approximately 37.8\degree N 76.1\degree W)} and serves as the outlet of 9 major tributary systems.
+Chesapeake Bay is located in the Mid-Atlantic region of the United States \hl{(at approximately 37.8\degree N 76.1\degree W, see Figure S1)} and serves as the outlet of 9 major tributary systems.
 Of these, the Susquehanna River (at the Northern extent of the Bay) provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \cite{xuClimateForcingSalinity2012}).
@@
 Our tuned RF models for salinity, temperature, and turbidity had a root mean square error (RMSE) of 2.34 ($R^2$=0.91), 1.76 ($R^2$=0.96), and 1.82 ($R^2$=0.11) respectively (Table 1, Figure 2).
-Out-of-sample validation performance was good for salinity and temperature models but poor for the turbidity model (Figure 2, Figure S1).
+Out-of-sample validation performance was good for salinity and temperature models but poor for the turbidity model (Figure 2, Figure S2).
 Despite excellent overall performance for the temperature and salinity models, there was some indication of poor performance at the extremes.
@@
 One explanation for these areas of poor performance is that such extremes were rare in the observational dataset (Table S2).
-The salinity model however did faithfully capture areas of low salinity (Figure 2, S3).
-These low salinity values can be regarded as low uncertainty (high confidence) estimates given the extent to which the upper tributaries are well represented in the observational dataset (Figure S2).
+The salinity model however did faithfully capture areas of low salinity (Figure 2, S4).
+These low salinity values can be regarded as low uncertainty (high confidence) estimates given the extent to which the upper tributaries are well represented in the observational dataset (Figure S3).
 
@@
 This coverage allows for a more comprehensive assessment of the spatial uncertainty of our model.
-Predicted areas of low salinity coming from our model likely have particularly low associated uncertainties because the upper tributaries were well-represented in our observational dataset (Figure S2).
+Predicted areas of low salinity coming from our model likely have particularly low associated uncertainties because the upper tributaries were well-represented in our observational dataset (Figure S3).
 Because CBOFS salinity data was fresher than our RF estimates in the upper tributaries, this suggests that CBOFS data for these areas are likely true underestimates and not a deficiency of the RF model predictions.
@@
 The hydrodynamics here seem to be complex given that the CBOFS has difficulty resolving spatial patterning in this area \cite{lanerolle2011second}.
-This was an area where we found limited observational data (Figure S2).
+This was an area where we found limited observational data (Figure S3).
 Improvements to data-driven modeling in this area may be possible with the inclusion of more observational data.

### 2022-11-17 manuscript/supplement.tex
move validation plot from supp to ms
formally include the obs_distribution table
text edits
prep pdf ms
include study site fig in supplement

@@
     \begin{center}
+          \includegraphics[width=0.85\textwidth,keepaspectratio]{figures/_study_site}
+          \caption{Map of Cheseapeake Bay study site.}
+    \end{center}    
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
           \includegraphics[width=0.85\textwidth,keepaspectratio]{figures/counts_obs_table}
@@
     \begin{center}
-          \includegraphics[width=0.7\textwidth,keepaspectratio]{figures/_validation}
-          \caption{RF predicted salinity versus test set (observed) salinity. Hatching indicates the area containing 80\% of the data points.}
+          \includegraphics[width=0.85\textwidth,keepaspectratio]{figures/obs_distribution_table}
     \end{center}    
@@
 
+\begin{figure}[h!]
+    \begin{center}
+          \includegraphics[width=1\textwidth,keepaspectratio]{figures/_seasonality_turbidity}
+          \caption{Seasonal Random Forest prediction results for turbidity.}
+    \end{center}    
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+          \includegraphics[width=1\textwidth,keepaspectratio]{figures/_freqcount_hex}
+          \caption{Locations of observational data.}
+    \end{center}    
+\end{figure}
+
+\begin{figure}[h!]
+    \begin{center}
+          \includegraphics[width=1\textwidth,keepaspectratio]{figures/_validation_map}
+          \caption{Map of model validation errors.}
+    \end{center}    
+\end{figure}
+
 \clearpage

### 2022-11-17 manuscript/manuscript.tex
text edits

@@
 Applications such as these often turn to gridded alternatives to direct observations such as output from directly-observing satellite-based sensors \cite{foreCombinedActivePassive2016} or operational now/fore-casting products \cite{lanerolle2011second}.
-Unfortunately, these too are of limited use for applications such as validation of next-generation Earth Systems Models (ESMs, \cite{golazDOEE3SMModel2022}), where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
+Unfortunately, these too are of limited use for applications such as validation of next-generation Earth Systems Models (ESMs, \cite{golazDOEE3SMModel2022, xuDevelopmentLandRiver2022a}), where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
 
@@
 We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the \hl{1000m} resolution MODIS Aqua product, which provides \hl{9} bands of values for each pixel spanning from \hl{the following wavelength ranges (405-420, 438-448, 483-493, 526-536, 546-556, 662-672, 673-683, 743-753, and 862-877nm)} \cite{vermoteericMOD09GAMODISTerra2015}.
-\hl{MODIS data was pulled from Google Earth Engine (ID:MODIS/006/MYDOCGA) using the \texttt{earthengine-api} Python package.}
+\hl{MODIS data was pulled from Google Earth Engine (Asset: \texttt{MODIS/006/MYDOCGA}) using the \texttt{earthengine-api} Python package.}
 
@@
 
+\hl{Across all models, MODIS band 8 was the most important reflectance feature.
+This result may be partly explained by the fact that Band 8 is the shortest wavelength (405-420nm) included in our study and is likely related to the concentration of Colored Dissolved Organic Matter (CDOM) the absorption of which declines exponentially with increasing wavelength under simulated and laboratory conditions} \cite{loiselleOpticalCharacterizationChromophoric2009, sathyendranath2000remote}).
+\hl{CDOM itself is strongly related to salinity} \cite{sathyendranath2000remote, vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018} \hl{and thus may be related to other properties such as temperature and turbidity which also vary spatially with the upstream-downstream environmental gradient of estuaries.
+In field conditions and in remote sensing applications, the absoption spectra of CDOM and the shape of the CDOM-salinity relationship is highly variable from region-to-region as a function of local factors such as bottom reflectance and phytoplankton community composition} \cite{sathyendranath2000remote}.
+\hl{Although band 8 was not universally the most important feature in prior remote sensing studies of estuarine water properties, the general importance of short wavelength bands is a result shared by prior studies on estuarine salinity} \cite{wangDevelopmentMODISData2018}.
+
 \subsection{Spatial patterning}

### 2022-11-17 manuscript/manuscript.tex
text edits

@@
 Unfortunately, these too are of limited use for applications such as validation of next-generation Earth Systems Models (ESMs, \cite{golazDOEE3SMModel2022, xuDevelopmentLandRiver2022a}), where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
+\hl{There is a strong need in particular for satellite-based products to validate recently developed biogeochemical modules of the E3SM Earth Systems model which aims to represent thermal, sediment, salinity, nutrient exchanges between rivers and the land on a global basis} \cite{xuDevelopmentLandRiver2022a}.
 
@@
 geographic “awareness”) especially in areas of the Bay with little to no observations.
-Our model represents a further advantage over prior efforts at data-driven modeling of salinity in the Bay (e.g.
-\cite{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
-These more spatially extensive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
+\hl{Our model was trained against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.} These more spatially extensive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
 
@@
 We chose to use the Random Forest algorithm in part because \cite{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) and also because of its ease of implementation and interpretation.
+\hl{Random Forests have a particular advantage because they are robust against interactions among variables and allow for nonlinearities} \cite{prasadNewerClassificationRegression2006}.
 We fit RF models using the scikit-learn Python package \cite{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model (Figure 1).

### 2022-11-18 manuscript/manuscript.tex
add data citation
comment out inline figures

@@
 (2023a) the pipeline for which is shown in Figure 1.
-All data used in the study are available in Stachelek et al.
-(2023b).
+All data used in the study are available in \cite{stachelekjemmachesapeake}.
 
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/data-processing_model-architecture}
-          \caption{Diagram of the data processing, model fitting, tuning, and prediction pipeline.}
-    \end{center}    
-\end{figure}
+% \begin{figure}[ht!]
+%     \begin{center}
+%           \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/data-processing_model-architecture}
+%           \caption{Diagram of the data processing, model fitting, tuning, and prediction pipeline.}
+%     \end{center}    
+% \end{figure}
 
@@
 
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
-          \caption{Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.}
-    \end{center}    
-\end{figure}
+% \begin{figure}[ht!]
+%     \begin{center}
+%           \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
+%           \caption{Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.}
+%     \end{center}
+% \end{figure}
 
@@
 
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_importance_all}
-          \caption{Random Forest importance plot for salinity, temperature, and turbidity models.
+% \begin{figure}[ht!]
+%     \begin{center}
+%           \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_importance_all}
+%           \caption{Random Forest importance plot for salinity, temperature, and turbidity models.
 Note that all features may not be present in all models if they were dropped as a result of recursive feature elimination.}
-    \end{center}    
-\end{figure}
+%     \end{center}    
+% \end{figure}
 
@@
 
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.7\textwidth,keepaspectratio]{figures/_seasonality}
-          \caption{Seasonal Random Forest prediction results for temperature and salinity.}
-    \end{center}    
-\end{figure}
+% \begin{figure}[ht!]
+%     \begin{center}
+%           \includegraphics[width=0.7\textwidth,keepaspectratio]{figures/_seasonality}
+%           \caption{Seasonal Random Forest prediction results for temperature and salinity.}
+%     \end{center}    
+% \end{figure}
 
@@
 
-\begin{figure}[ht!]
-    \begin{center}
-          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
-          \caption{Comparison between Random Forest salinity prediction results and a CBOFS snapshot for Sept, 4, 2022.}
-    \end{center}    
-\end{figure}
+% \begin{figure}[ht!]
+%     \begin{center}
+%           \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
+%           \caption{Comparison between Random Forest salinity prediction results and a CBOFS snapshot for Sept, 4, 2022.}
+%     \end{center}    
+% \end{figure}
 

### 2022-11-18 manuscript/manuscript.tex
cite zenodo

@@
 
-All code for data processing, model fitting, and model evaluation is available at Stachelek et al.
-(2023a) the pipeline for which is shown in Figure 1.
+All code for data processing, model fitting, and model evaluation for the pipeline shown in Figure 1 is available in \cite{jemma_stachelek_2022_7332559}.
 All data used in the study are available in \cite{stachelekjemmachesapeake}.

### 2022-12-09 manuscript/manuscript.tex
add sofia bio
use sofia's bio
rm highlighting

@@
 
-Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as \hl{Sea Level Rise (SLR)} forecasting and habitat suitability assessments \cite{hoodChesapeakeBayProgram2021}.
+Unfortunately, direct observations alone are not adequate for many applications that require spatially extensive data such as Sea Level Rise (SLR) forecasting and habitat suitability assessments \cite{hoodChesapeakeBayProgram2021}.
 For example, habitat suitability assessments are typically done on a gridded basis for the entirety of a particular estuary rather than at selected point locations.
@@
 Unfortunately, these too are of limited use for applications such as validation of next-generation Earth Systems Models (ESMs, \cite{golazDOEE3SMModel2022, xuDevelopmentLandRiver2022a}), where the temporal availability of operational forecasting products is too short ($<$ 1 year) and the spatial resolution of directly-observing satellite-based sensors is too coarse ($>$ 30 $km^2$).
-\hl{There is a strong need in particular for satellite-based products to validate recently developed biogeochemical modules of the E3SM Earth Systems model which aims to represent thermal, sediment, salinity, nutrient exchanges between rivers and the land on a global basis} \cite{xuDevelopmentLandRiver2022a}.
+There is a strong need in particular for satellite-based products to validate recently developed biogeochemical modules of the E3SM Earth Systems model which aims to represent thermal, sediment, salinity, nutrient exchanges between rivers and the land on a global basis \cite{xuDevelopmentLandRiver2022a}.
 
 To address this validation data gap, we developed a data-driven model to produce high-resolution (< 1 $km^2$) estimates of temperature, salinity, and turbidity at daily time-steps (over decadal time scales) as required by next-generation ESMs.
-The primary input to our model is \hl{remote sensing} reflectance data.
+The primary input to our model is remote sensing reflectance data.
 In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \cite{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}.
@@
 geographic “awareness”) especially in areas of the Bay with little to no observations.
-\hl{Our model was trained against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.} These more spatially extensive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
+Our model was trained against not only buoy observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.
+These more spatially extensive observations likely increase the accuracy and reliability of model predictions in the upper tributaries.
 
@@
 
-Chesapeake Bay is located in the Mid-Atlantic region of the United States \hl{(at approximately 37.8\degree N 76.1\degree W, see Figure S1)} and serves as the outlet of 9 major tributary systems.
+Chesapeake Bay is located in the Mid-Atlantic region of the United States (at approximately 37.8\degree N 76.1\degree W, see Figure S1) and serves as the outlet of 9 major tributary systems.
 Of these, the Susquehanna River (at the Northern extent of the Bay) provides the majority of the freshwater supply to the Bay (\texttildelow$48\%$, \cite{xuClimateForcingSalinity2012}).
-\hl{The Bay is the largest estuary in the United States and has an area of approximately 11,000 $km^2$.
-As such, it is provides a wealth of natural economic benefits in the form of fisheries harvests (particularly oysters), tourism, and nature-based recreation} \cite{hoodChesapeakeBayProgram2021}.
+The Bay is the largest estuary in the United States and has an area of approximately 11,000 $km^2$.
+As such, it is provides a wealth of natural economic benefits in the form of fisheries harvests (particularly oysters), tourism, and nature-based recreation \cite{hoodChesapeakeBayProgram2021}.
 Despite its large size, the Bay is fairly shallow ($<$ 30m, \cite{murphyComparisonSpatialInterpolation2010}) and the hydrodynamics of the Bay are such that it is considered stratified and partially mixed \cite{xuClimateForcingSalinity2012}.
@@
 We chose to use the Random Forest algorithm in part because \cite{urquhartGeospatialInterpolationMODISderived2013} found that it produced nearly equivalent results to more complex methods (particularly NN) and also because of its ease of implementation and interpretation.
-\hl{Random Forests have a particular advantage because they are robust against interactions among variables and allow for nonlinearities} \cite{prasadNewerClassificationRegression2006}.
+Random Forests have a particular advantage because they are robust against interactions among variables and allow for nonlinearities \cite{prasadNewerClassificationRegression2006}.
 We fit RF models using the scikit-learn Python package \cite{pedregosaScikitlearnMachineLearning2011} where each model was initialized with 250 trees, had a maximum depth of 20 tree “layers”, and was subjected to recursive feature elimination to produce the most parsimonious model (Figure 1).
-\hl{We chose initial forest size and maximum depth values to optimize model performance and model training time respectively.} We subsequently “tuned” each model using k-fold cross validation to optimize the size, depth, and splitting criteria of each tree.
+We chose initial forest size and maximum depth values to optimize model performance and model training time respectively.
+We subsequently “tuned” each model using k-fold cross validation to optimize the size, depth, and splitting criteria of each tree.
 We evaluated the “importance” of particular variables as the average impurity decrease of tree splits based on that variable \cite{pedregosaScikitlearnMachineLearning2011}.
@@
 Then the distance from any one cell to a tributary source was the accumulated travel “cost” as given by the scikit-image \texttt{route\_through\_array} function.
-We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the \hl{1000m} resolution MODIS Aqua product, which provides \hl{9} bands of values for each pixel spanning from \hl{the following wavelength ranges (405-420, 438-448, 483-493, 526-536, 546-556, 662-672, 673-683, 743-753, and 862-877nm)} \cite{vermoteericMOD09GAMODISTerra2015}.
-\hl{MODIS data was pulled from Google Earth Engine (Asset: \texttt{MODIS/006/MYDOCGA}) using the \texttt{earthengine-api} Python package.}
+We combined FWI estimates with the pixel-wise daily mean observations and regressed them in RF models against data from the 1000m resolution MODIS Aqua product, which provides 9 bands of values for each pixel spanning from the following wavelength ranges (405-420, 438-448, 483-493, 526-536, 546-556, 662-672, 673-683, 743-753, and 862-877nm) \cite{vermoteericMOD09GAMODISTerra2015}.
+MODIS data was pulled from Google Earth Engine (Asset: \texttt{MODIS/006/MYDOCGA}) using the \texttt{earthengine-api} Python package.
 
@@
 
-\hl{Across all models, MODIS band 8 was the most important reflectance feature.
-This result may be partly explained by the fact that Band 8 is the shortest wavelength (405-420nm) included in our study and is likely related to the concentration of Colored Dissolved Organic Matter (CDOM) the absorption of which declines exponentially with increasing wavelength under simulated and laboratory conditions} \cite{loiselleOpticalCharacterizationChromophoric2009, sathyendranath2000remote}).
-\hl{CDOM itself is strongly related to salinity} \cite{sathyendranath2000remote, vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018} \hl{and thus may be related to other properties such as temperature and turbidity which also vary spatially with the upstream-downstream environmental gradient of estuaries.
-In field conditions and in remote sensing applications, the absoption spectra of CDOM and the shape of the CDOM-salinity relationship is highly variable from region-to-region as a function of local factors such as bottom reflectance and phytoplankton community composition} \cite{sathyendranath2000remote}.
-\hl{Although band 8 was not universally the most important feature in prior remote sensing studies of estuarine water properties, the general importance of short wavelength bands is a result shared by prior studies on estuarine salinity} \cite{wangDevelopmentMODISData2018}.
+Across all models, MODIS band 8 was the most important reflectance feature.
+This result may be partly explained by the fact that Band 8 is the shortest wavelength (405-420nm) included in our study and is likely related to the concentration of Colored Dissolved Organic Matter (CDOM) the absorption of which declines exponentially with increasing wavelength under simulated and laboratory conditions \cite{loiselleOpticalCharacterizationChromophoric2009, sathyendranath2000remote}.
+CDOM itself is strongly related to salinity \cite{sathyendranath2000remote, vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018} and thus may be related to other properties such as temperature and turbidity which also vary spatially with the upstream-downstream environmental gradient of estuaries.
+In field conditions and in remote sensing applications, the absoption spectra of CDOM and the shape of the CDOM-salinity relationship is highly variable from region-to-region as a function of local factors such as bottom reflectance and phytoplankton community composition \cite{sathyendranath2000remote}.
+Although band 8 was not universally the most important feature in prior remote sensing studies of estuarine water properties, the general importance of short wavelength bands is a result shared by prior studies on estuarine salinity \cite{wangDevelopmentMODISData2018}.
 
@@
 
-We have demonstrated an approach to produce realistic estimates of water properties in Chesapeake Bay using \hl{remote sensing} data.
+We have demonstrated an approach to produce realistic estimates of water properties in Chesapeake Bay using remote sensing data.
 Salinity and temperature model performance was as good or better than prior efforts.
@@
 Her research focuses on the applications of spatially intensive geocomputation to water and food security problems particularly those related to eutrophication and nutrient pollution.
-    \item \textbf{Sofia Avendaño} is awesome.
+    \item \textbf{Sofia Avendaño} is a hydrogeologist with a Master's from New Mexico Institute of Mining and Technology.
+She has worked on a wide variety of projects including fracture flow, surface flow and hydrodynamics.
+Her scientific passion is modeling fluid flow and sediment transport.
     \item \textbf{Jon Schwenk} is a River and Data Scientist (PhD) in the Division of Earth and Environmental Sciences at Los Alamos National Laboratory.
@@
     \item Figure 3: Random Forest importance plot for salinity, temperature, and turbidity models.
-\hl{Note that all features may not be present in all models if they were dropped as a result of recursive feature elimination.}
+Note that all features may not be present in all models if they were dropped as a result of recursive feature elimination.
     \item Figure 4: Seasonal Random Forest prediction results for temperature and salinity.

### 2022-12-09 manuscript/manuscript.tex
re-enable inline figs

@@
 
-% \begin{figure}[ht!]
-%     \begin{center}
-%           \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/data-processing_model-architecture}
-%           \caption{Diagram of the data processing, model fitting, tuning, and prediction pipeline.}
-%     \end{center}    
-% \end{figure}
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/data-processing_model-architecture}
+          \caption{Diagram of the data processing, model fitting, tuning, and prediction pipeline.}
+    \end{center}    
+\end{figure}
 
@@
 
-% \begin{figure}[ht!]
-%     \begin{center}
-%           \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
-%           \caption{Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.}
-%     \end{center}
-% \end{figure}
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_validation}
+          \caption{Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.}
+    \end{center}
+\end{figure}
 
@@
 
-% \begin{figure}[ht!]
-%     \begin{center}
-%           \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_importance_all}
-%           \caption{Random Forest importance plot for salinity, temperature, and turbidity models.
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.8\textwidth,keepaspectratio]{figures/_importance_all.png}
+          \caption{Random Forest importance plot for salinity, temperature, and turbidity models.
 Note that all features may not be present in all models if they were dropped as a result of recursive feature elimination.}
-%     \end{center}    
-% \end{figure}
+    \end{center}    
+\end{figure}
 
@@
 
-% \begin{figure}[ht!]
-%     \begin{center}
-%           \includegraphics[width=0.7\textwidth,keepaspectratio]{figures/_seasonality}
-%           \caption{Seasonal Random Forest prediction results for temperature and salinity.}
-%     \end{center}    
-% \end{figure}
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.7\textwidth,keepaspectratio]{figures/_seasonality}
+          \caption{Seasonal Random Forest prediction results for temperature and salinity.}
+    \end{center}    
+\end{figure}
 
@@
 
-% \begin{figure}[ht!]
-%     \begin{center}
-%           \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
-%           \caption{Comparison between Random Forest salinity prediction results and a CBOFS snapshot for Sept, 4, 2022.}
-%     \end{center}    
-% \end{figure}
+\begin{figure}[ht!]
+    \begin{center}
+          \includegraphics[width=0.75\textwidth,keepaspectratio]{figures/_rf-vs-cbofs}
+          \caption{Comparison between Random Forest salinity prediction results and a CBOFS snapshot for Sept, 4, 2022.}
+    \end{center}    
+\end{figure}
 
@@
 
-\clearpage
-
-\section*{Figure Captions}
-
-\begin{description}
-    \item Figure 1: Diagram of the data processing, model fitting, tuning, and prediction pipeline.
-    \item Figure 2: Out of sample test set performance for salinity, turbidity, and temperature models respectively. Hatching indicates the location of 80\% of the data points.
-    \item Figure 3: Random Forest importance plot for salinity, temperature, and turbidity models.
-Note that all features may not be present in all models if they were dropped as a result of recursive feature elimination.
-    \item Figure 4: Seasonal Random Forest prediction results for temperature and salinity.
-    \item Figure 5: Comparison between Random Forest salinity prediction results and a CBOFS snapshot for Sept, 4, 2022.
-\end{description}
-
 \end{document}

### 2022-12-31 scripts/animation.py
cleanup animation

@@ -32,12 +32,6 @@ test = xr.concat(xr_list, dim="time")
 
 variable = test.band_data.sel(time=slice('2021', '2022'))
 
-fig = plt.figure()
-ax = plt.axes(projection=ccrs.PlateCarree())
-image = variable.isel(time=0).squeeze().plot.imshow(
-    ax=ax, transform=ccrs.PlateCarree(), animated=True)
-
-
 def update(t):
     # t = variable.time.values[0]
     print(t)
@@ -45,14 +39,24 @@ def update(t):
     image.set_array(variable.sel(time=t).squeeze())
     return image,
 
+fig = plt.figure(figsize=plt.figaspect(1.22))
+fig.subplots_adjust(0.1,0,1,1)
+ax = plt.axes(projection=ccrs.PlateCarree())
+image = variable.isel(time=0).squeeze().plot.imshow(
+    ax=ax, transform=ccrs.PlateCarree(), animated=True)
 
 animation = anim.FuncAnimation(fig,
                                update,
-                               interval=700,
+                               interval=400,
                                frames=variable.time.values,
                                blit=False)
-# animation.save("test.gif")
-plt.show()
+
+plt.box(False)
+ax.annotate('Salinity', xy = (0, 0), xycoords='figure fraction',
+            xytext=(0.04, 0.97), textcoords='figure fraction', fontsize=14
+            )
+animation.save("cbofs.gif")
+# plt.show()
 
 # --- simple plotting
 test = read_xr("2018-01-01")
@@ -63,4 +67,4 @@ ax = plt.axes(projection=ccrs.PlateCarree())
 image = da.squeeze().plot.imshow(ax=ax,
                                  transform=ccrs.PlateCarree(),
                                  animated=True)
-plt.show()
+plt.show(block=False)
